### 2010-06-11 Text/migration.tex
Initial import of text.
git-svn-id: http://jaga.astro.indiana.edu/paperrepo/ApJL/Migration2010@63 3b7f7548-5c66-0410-83ab-3997167635fc

%\documentclass[12pt,preprint]{aastex}

\documentclass[12pt,manuscript,authoryear]{aastex}

\usepackage{graphicx}

\begin{document}

\title{Migration of Gas Giant Planets in Gravitationally Unstable Disks}

\author{{\it Short Title: Gas Giant Migration ~~~~Article Type: Journal}}

\author{Scott Michael}
\affil{Department of Astronomy, Indiana University, Bloomington, IN 47405}
\email{scamicha@astro.indiana.edu}

\and

\author{Richard H. Durisen}
\affil{Department of Astronomy, Indiana University, Bloomington, IN 47405}
\email{durisen@astro.indiana.edu}

\and

\author{Aaron C. Boley}
\affil{Department of Astronomy; University of Florida,
211 Bryant Space Science Center,
PO Box 112055, Gainesville, FL 32611}
\email{aaron.boley@gmail.com}

\begin{abstract}

Characterization of migration in gravitationally unstable disks is necessary to understand the fate of protoplanets formed by disk instability. As part of a larger study, we are using a 3D radiative hydrodynamics code to investigate how an embedded gas giant planet interacts with a gas disk that is massive enough to experience gravitational instabilities (GIs). This Letter presents preliminary results from simulations with a Jupiter-mass planet placed in orbit at 25 AU within a 0.13 $M_{\odot}$ disk. The disk spans 5 to 40 AU around a 1 $M_{\odot}$ star and is initially only marginally stable. In one simulation, the planet is inserted prior to the initial eruption of GIs; in another, it is inserted only after the disk has settled into a quasi-steady GI-active state, where heating by GIs roughly balances radiative cooling. When the planet is present from the beginning, its own wake stimulates growth of a particular global mode with which it strongly interacts, and the planet plunges inward six AU in about 10$^3$ years. In both cases, there are times when the planet's radial motion is slow and random. At other times, when the planet interacts with a strong spiral mode near the mode's corotation radius, migration both inward and outward can be relatively rapid, covering several AUs over hundreds of years. {\bf See note in Latex file.}
%The assertion in the preceding sentence needs to be verified by analyzing both runs in more detail -- Am(t), periodograms, images zoomed in on the planet to check for arms. The migration is observed; it is the relationship to mode interactions that needs to be checked.
Planet orbit eccentricity tends to fluctuate rapidly and ranges between about 0.02 to 0.1.
\end{abstract}

\keywords{hydrodynamics --- instabilities --- planet-disk interactions --- planets and satellites: formation --- protoplanetary disks}

\label{firstpage}

\section{Introduction}

{\bf See note in Latex file.} 
%Scott: File in some references you know now. Aaron: Suggestions on references?
Migration of gas giant planets due to interactions with a circumstellar gas disk probably plays a major role in defining the architecture of planetary systems (e.g., ???). Work on migration (see review by ???  Balbus \& Terquem or Papaloizou ???) has included gravitational interaction of planets with both laminar and turbulent disks. However,  the self-gravity of the gas disk has usually been ignored, and the effects of radiative cooling and a detailed equation of state (EOS) have only recently been included (Paardekooper ???). Boss (2004 ???) and Mayer et al. (2004) examined radial migration of planet-mass fragments in gravitationally unstable disks, but their disks were violently disrupted by fragmentation under conditions (radii $<$ 40 AU, disk-to-star mass ratios $M_d/M_s \sim 0.1$, and stellar mass $M_s \sim 1\;M_{\odot}$) where this probably should not realistically have happened (Rafikov 2005, 2007; Boley et al. 2006, 2007b; Boley \& Durisen 2008; Forgan et al. 2009; Cai et al. 2010). More recently, fragmentation into clumps with gas giant or brown dwarf masses has been documented in disks that are relatively massive ($M_d/M_s \sim$ a few tenths) and spatially extended (outer radii $>$ 100 AU) (Krumholz et al. 2007; Stamatellos \& Whitworth 2007, 2009, Boley 2009; Hayfield et al. 2010), where fragmentation is expected from semi-analytic arguments (e.g., Clarke ???, Rafikov 2009, Dobson-Robinson et al. 2009). The fate of the clumps depends in part on their radial migration, which is a chaotic and messy affair in a fragmenting disk (e.g., Boley 2009, Boley et al. 2010, Vorobyov \& Basu 2010; Boley \& Durisen 2010). The occurrence of gravitational instabilities (GIs) may be episodic (Vorobyov \& Basu 2005, 2006). Clumps that survive and contract to the dimensions of young planets can later find themselves in a disk that erupts again into GI activity. As the star/disk system evolves, such a protoplanet may end up in a region of a GI-active disk where fragmentation does not occur.

To improve our understanding of how planets migrate in GI-active disks that are not fragmenting, we have begun a systematic study, using numerical 3D radiative hydrodynamics, where we investigate the effects on both the disk and the planet of inserting a planet-mass object into disks susceptible to GIs. Using techniques deployed in earlier research by our group (Pickett et al. 2003; Mej\'ia et al. 2005; Cai et al. 2006, 2008; Boley et al 2006, 2007b), we can identify the dominant spiral waves in a simulation and analyze how the waves interact with the planet's motion. Our goal is to determine both the effect of giant planets on GIs and the effect of GIs on planet migration. Because GIs are sensitive to radiative physics, we use a well-tested radiative scheme (Boley et al. 2007b) and realistic opacities (D'Alessio et al. 2001).

Section 2 below presents our numerical methods and initial conditions.  We describe and discuss the simulation results in Section 3, and summarize our main conclusions in Section 4.

\section{Computational Methodology}

\subsection{3-D Radiative Hydrodynamics}

{\bf See note in Latex file.}
%Scott and Aaron: Verify that everything I say here is correct.
The CHYMERA (Computational HYdrodynamics with MultiplE Radiation Algorithms) code (Boley et al. 2007b) is a second-order, explicit, Eulerian scheme on a 3D cylindrical grid. The code uses a realistic equation of state for H$_2$ (Boley et al. 2007a) and integrates an energy equation that includes $PdV$ work, net heating or cooling due to radiative flux divergence, and heating by artificial bulk viscosity. Calculations are done on a uniform cylindrical grid with reflection symmetry about the disk midplane and a resolution ($\varpi$,$\phi$,z) = (512,512,64). The $z$-axis is the rotation axis of the disk. The large number of azimuthal zones is necessary to resolve the planet's Hill sphere and the planet's wake. These simulations utilize the radiative cooling scheme developed and tested in Boley et al. (2007b), where the optically thick monochromatic flux in the $\varpi$ and $\phi$-directions is computed by flux-limited radiative diffusion and where the radiative transport of energy in the $z$-direction is solved using a one-ray discrete ordinate method in both optically thin and thick regions. Although the central star remains fixed at the grid center, we account for acceleration of the reference frame by the planet and by the disk via indirect potentials, as in Michael \& Durisen (2010), and, overall, the planet integration is done using methods similar to those in Nelson et al. (2000). The Rosseland mean and Planck mean opacities and molecular weights in our simulations are the same as those in Boley et al. (2006, 2007b), except that we correct the D'Alessio et al. (2001) gas mean molecular weight to $\mu$ = 2.39.

\subsection{Initial Model and Simulation Conditions}

{\bf See note in Latex file.}
%Scott: Verify the accuracy of all this.

The model disk, based on an equlibrium model in Pickett et al. (2003), orbits a 1 $M_{\odot}$ star and has a mass $M_d = 0.13 M_{\odot}$, inner and outer radii at 5 and 40 AU, and an initial surface density distribution $\Sigma \sim \varpi^{-1}$. The time unit of one ORP (= outer rotation period) is defined as the rotation period of the initial disk at $\varpi \approx$ 32 AU, or about 250 yr. The disk is initially located between radial grid zones ??? 30 ??? and 240 and is close to isentropic, which results in a Toomre-$Q$ distribution with a marginally unstable (see Durisen et al. 2007) minimum value of ??? at radial grid zone ??? (??? AU). The computational grid extends radially to 512 zones to accommodate expansion of the outer disk once GIs become nonlinear. There is an inner cylindrical hole in the grid with a radius of 13 radial zones. To seed nonaxisymmetry, the density is given an initial ??? \% random cell-to-cell perturbation.

This Letter presents three simulations. The first, which we call the fiducial run, goes from time $t = 0$ to ??? 20 ??? ORP without a planet. In the second simulation, which we call the $t = 0$ planet run, a one Jupiter-mass ($M_J$) planet with a roughly circular midplane orbit is included in the initial $t = 0$ ORP equilibrium disk at 25 AU. In the third simulation, which we call the $t = 10$ planet run, a 1 $M_J$ planet with a similar orbit is inserted into the fiducial run at $t = 10$ ORP. These three simulations are part of a much larger suite, to be described elsewhere, in which the planet mass and starting locations are varied. We estimate crude semi-major axes $a$ and eccentricities $e$ from the planet's $\varpi(t)$ by finding at the maximum and minimum radii, $\varpi_{\rm max}$ and $\varpi_{\rm min}$, over each interval that the planet's $\phi$ changes by 2$\pi$ (one ``orbit'') and then using $a = (\varpi_{\rm max}+\varpi_{\rm min})/2$ and $e = (\varpi_{\rm max}-\varpi_{\rm min})/(\varpi_{\rm max}+\varpi_{\rm min})$. These orbital elements are only intended to be indicative; there are strong orbit to orbit variations in these $a$ and $e$.

\section{Results}

{\it The Fiducial Run.} As in Mej\'ia et al (2005), the fiducial run exhibits four distinct phases: initial {\sl axisymmetric cooling}, the onset of GIs in a well-defined {\sl burst}, a {\sl transition}, and the {\sl asymptotic} establishment of quasi-steady chaotic GI-activity with an overall balance of heating and cooling. There is no fragmentation because this disk has a long cooling time (see Gammie 2001; Boley et al. 2006, 2007b).

%Scott: A number of things in this paragraph need to verified by further analysis.
{\it The $t = 0$ Planet Run.} {\bf See note in Latex file.} Figure 1 shows that inclusion of even this modest mass planet ($0.8 \%$ of $M_d$) has a dramatic effect on the burst phase. Without a planet, when all modes grow from imposed noise, it takes 4 ORPs for coherent spiral modes to organize and grow to nonlinear amplitude. The growth is centered near the $Q$-minimum at ??? AU. Modes of cos($m\phi$) symmetry with $m =$ 5 and 6 dominate. An $m = 3$ (three-armed) spiral also grows somewhat more slowly, but lags the other modes in time. In the $t = 0$ planet run, the planet develops an organized wake within about an ORP in which $m = 3$ is a dominant component ($\sim 5$ \% global density perturbation). This nonlinear seeding causes $m = 3$ to dominates the GI burst, which also occurs $\sim$ 2 ORPs earlier. Because of our initial placement of the planet inside but close to the $Q$-minimum, the corotation radius (CR) of the $m = 3$ mode is fairly close to that of the planet's orbit period. When the triggered $m = 3$ mode becomes strongly nonlinear at about 3 ORP, planet migration is significantly affected. Figure 2 shows the evolution of the planet's radial position. From 0 to 2.5 ORP, the planet is torqued primarily by its own wake and migrates inward. Beginning at 3 ORP, when the $m = 3$ mode has grown significantly and the planet is lagging one of its arms in azimuth, the planet gains angular momentum and moves outward. At about $t =$ 4 ORP, as the planet overtakes the arm, the sign of the net torque reverses and the planet plunges from 23 to 17 AU in about 4 ORP (10$^3$ yr). During this time, the planet and mode seem to interact strongly enough that the planet essentially surfs along the arm, maintaining a location within the arm as it plunges. After $t =$ 8 ORP, the main burst is over, and the disk transitions into its asymptotic state, where modes of many $m$-values are comparably strong. The planet's radial migration apparently stalls at about 16 to 17 AU. From an analysis of periodicities present in the gas disk between 8 and 12 ORP, the planet lies a few AU inside the inner Lindblad resonance of a strong $m = 2$ mode with CR at ??? AU in the GI-active disk.

%Scott: Even more work is needed to verify what is said here. Some of what I wrote is pure speculation and needs to be checked.
{\it The $t = 10$ Planet Run.} {\bf See note in Latex file.} The motion of the planet in this case is more difficult to interpret. We see intervals of fairly rapid inward or outward migration over several orbits, as well as times when radial migration appears to stall. Between $t = 15$ and 19 ORP, the pattern of outward migration for 2 ORP followed by an inward plunge over the next 2 ORP resembles the behavior in the $t = 0$ planet run during the burst, but, in this case, there is no distinct phase of GI activity. Animations of the simulation show there is a complex interaction between the planet, its spiral wake, and the global spiral arms caused by the instabilities. Analysis of images and periodicities suggests that, between 15 and 19 ORP, the planet may have passed through an arm of an $m =$ ??? pattern with CR at ??? AU. It is possible that the planet actually stimulated this particular mode, in manner similar to what occurred in the $t = 0$ planet run, because this $m =$ ??? periodicity is absent over the same time interval in the fiducial run.

{\it Eccentiricity.} Figure 3 shows the evolution of $e$ in the planet simulations. In both cases, the planets were inserted with an approximate circular velocity computed by adding the interior disk mass to the stellar point mass. The presence of GI activity in the $t = 10$ planet run immediately increases $e$ to $\sim 0.08$. In the $t = 0$ planet run, the modest initial $e$ decreases, as it should (ref ???), during the 1.5 ORP when it is migrating only due to its own wake. However, once there are strong nonlinear interactions with GI modes, $e$ jumps upward. In both runs, it appears that interaction with well-established GI activity leads to eccentricities between 0.02 and 0.1 that vary in a chaotic way between these extremes on a short time scale.

{\it Migration Relative to the Disk.} The disk surface density distribution does not vary much in the asymptotic phase. So radial motion of the planet in the $t = 10$ planet run is also radial migration relative to the gas. During the burst (see, e.g., Boley et al. 2006), however, the surface density of the disk changes dramatically. To verify that the planet is really migrating relative to the background gas disk in the $t = 0$ run, we have compared the planet's radius $\varpi_p(t)$ with the mass inside that radius, expressed in terms of $m_{\rm cyl}$, the fraction of the disk mass interior to the planet. Between $t = 3.52$ and $3.89$ ORP, when $\varpi_p$ increases modestly, $m_{\rm cyl}$ increases from about 0.50 to 0.53; between 3.89 and 8.47 ORP, when $\varpi_p$ decreases dramatically, $m_{\rm cyl}$ decreases from 0.53 to only 0.30. So, the planet is migrating significantly with respect to the disk mass even while the inner part of the gas disk moves inward during the burst. For a comparison measure of the magnitude of the gas motion, $m_{\rm cyl}$ at 23 AU increases from 0.50 to 0.60 from $t =$ 3.89 to 8.47 ORP.

\section{Discussion}

Check against estimated type I migration without GIs. {\bf Aaron: Can you do this?} Figure 2 suggests that the overall radial drift is not too different, but it is definitely not monotonic and can stall. Further work is needed to determine how long the stalls last.

Generic relationship to Laughlin's work. Just a few sentences -- structure (gaps, embedded masses) affects GI stability and the modes that grow.

Discrete mode interactions with the planet seem important -- Wind up and pitch. Surfing.

Stalling at LRs?
In both cases it is interesting to note that inward migration halts near 17 AU (coincidence???) ; this may be an artifact of the length of the simulations, clearly both simulations should be carried further. This radius is interesting because it is the location of the Inner Linblad Resonance (ILR) for the strong m = 2 mode present in the baseline simulation. Inside the ILR the mode can not effectively transfer angular momentum between the disk and the planet. Further work is necessary to adequately analyze the Fourier modes and their periodicities in the simulations with planets, but this gives a strong preliminary indication that the GI activity is driving the migration. Furthermore, where GIs are inactive planet migration is stalled, even Type I migration may be ineffective due to interference from the GIs at the planet’s OLR.

Probably not a random walk for this mass and this disk. Need to know dependence on planet mass and initial location.

Magnitude of e.

\section{Summary}

The inclusion of a 1MJup mass in the equilibrium disk initiates GIs and dramatically effects the character of the GIs.
GIs can cause both inward and outward planet migration and can cause rapid migration in the burst phase.
GIs can have a strong effect on the orbital eccentricity of the included planet.

\section{Acknowledgments}

The authors would like to thank F.C. Adams and J.C.B. Papaloizou for useful encouragement in this project. The research was conducted with the support of NASA grants from the Origins of Solar Systems  Program (NNG05GN11G and NNX08AK36G). {\bf Scott: Add necessary UITS and Teragrid acknowledgments.}

\begin{thebibliography}{99}

\bibitem[\protect\citeauthoryear{Boley}{2009}]{boley09} 
Boley,  A.C. 2009, ApJ, 695, L53

\bibitem[\protect\citeauthoryear{Boley \& Durisen}{2006}]{bd06} Boley, A.C., Durisen, R.H. 2006, ApJ, 641, 534

\bibitem[\protect\citeauthoryear{Boley \& Durisen}{2008}]{bd08} Boley, A.C., Durisen, R.H. 2008, ApJ, 685, 1193

\bibitem[\protect\citeauthoryear{Boley et al.}{2006}]{boley06} Boley, A.C., 
Mej\'ia, A.C., Durisen, R.H., Cai, K., Pickett, M.K. 2006, ApJ, 651, 517

\bibitem[\protect\citeauthoryear{Boley et al.}{2007}]{boley07} Boley, A.C., 
Durisen, R.H., Nordlund, A., Lord, J. 2007, ApJ, 665, 1254

\bibitem[\protect\citeauthoryear{Boley et al.}{2009}]{boleyetal09} Boley, A.C., 
Hayfield, T., Mayer, L., Durisen, R.H.  2010, Icarus, in press (arXiv:0909.4543)

\bibitem[\protect\citeauthoryear{Boss}{1984}]{boss84} Boss, A.P. 1984, ApJ, 277, 768

\bibitem[\protect\citeauthoryear{Boss}{1997}]{boss97} Boss, A.P. 1997, Science, 267, 1836

\bibitem[\protect\citeauthoryear{Boss}{1998}]{boss98} Boss, A.P. 1998, ApJ, 503, 923

\bibitem[\protect\citeauthoryear{Boss}{2000}]{boss00} Boss, A.P. 2000, ApJ, 536, L101

\bibitem[\protect\citeauthoryear{Boss}{2001}]{boss01} Boss, A.P. 2001, ApJ, 563, 367

\bibitem[\protect\citeauthoryear{Boss}{2002}]{boss02} Boss, A.P. 2002, ApJ, 576, 462

\bibitem[\protect\citeauthoryear{Boss}{2003}]{boss03} Boss, A.P. 2003, ApJ, 599, 577

\bibitem[\protect\citeauthoryear{Boss}{2004}]{boss04} Boss, A.P. 2004, ApJ, 610, 456

\bibitem[\protect\citeauthoryear{Boss}{2005}]{boss05} Boss, A.P. 2005, ApJ, 629, 535

\bibitem[\protect\citeauthoryear{Boss}{2007}]{boss07} Boss, A.P. 2007, ApJ, 661, L73

\bibitem[\protect\citeauthoryear{Boss}{2009}]{boss09} Boss, A.P. 2009, ApJ, 694, 107

\bibitem[\protect\citeauthoryear{Cai et al.}{2006}]{cai06} Cai, K., Durisen, R.H., Michael, S., Boley, A.C., Mej\'ia, A.C., Pickett, M.K., D'Alessio, P. 2006, ApJ, L149

\bibitem[\protect\citeauthoryear{Cai et al.}{2008}]{cai08} Cai, K., Durisen, R.H., Boley, A.C., Pickett, M.K., Mej\'ia, A.C. 2008, ApJ, 673, 1138

\bibitem[\protect\citeauthoryear{Clarke}{2009}]{clarke09} 
Clarke, C.J. 2009, MNRAS, 396, 1066

\bibitem[\protect\citeauthoryear{D'Alessio et al.}{2001}]{dalessio01} D'Alessio, P., Calvet, N., Hartmann, L. 2001, ApJ, 553, 321

\bibitem[\protect\citeauthoryear{Dodson-Robinson et al.}{2009}]{dodson09}
Dodson-Robinson, S. E., Veras, D., Ford, E. B., Beichman, C. A. 2009, ApJ, 707, 79

%\bibitem[\protect\citeauthoryear{Durisen, Mej\'ia, \& Pickett}
%{Durisen et al.}{2003}]{durisen03} Durisen, R.H., Mej\'ia, A.C.,

\bibitem[\protect\citeauthoryear{Durisen et al.}{2007}]{durisen07} Durisen R.H., Boss, A.P., Mayer, L., Nelson, A.F., Quinn, T., Rice, W.K.M. 2007, in: Protostars and Planets V, ed. B. Reipurth, D. Jewitt, \& K. Keil (Tucson:  Univ. of Arizona Press), 607

\bibitem[\protect\citeauthoryear{Durisen et al.}{2008}]{durisen08} 
Durisen R.H., Hartquis, T.W., \& Pickett M.K., Durisen R.H. 2008, ApSpSci., 317, 3

\bibitem[\protect\citeauthoryear{Forgan et al.}{2009}]{forg09} Forgan, D., Rice, K., Stamatellos, D., Whitworth, A. 2009, MNRAS 394, 882

\bibitem[\protect\citeauthoryear{Gammie}{2001}]{gammie01} Gammie, C.F. 2001, ApJ, 553, 174

\bibitem[\protect\citeauthoryear{Johnson \& Gammie}{2003}]{johnson03} 
Johnson, B.M., Gammie, C.F. 2003, ApJ, 597, 131

\bibitem[\protect\citeauthoryear{Haisch et al.}{2001}]{haisch01} 
Haisch, K.E., Jr., Lada, E.A., Lada, C.J. 2001, ApJ, 553, L153

\bibitem[\protect\citeauthoryear{Hayfield et al.}{2010}]{hayfield10}
Hayfield, T., Mayer, L., Wadsley, J., Boley, A. C. 2010, arXiv:1003.2594

\bibitem[\protect\citeauthoryear{Mayer et al.}{2002}]{mayer02} Mayer, L., Quinn, T. Wadsley, J., Stadel, J. 2002, Science, 298, 1756

\bibitem[\protect\citeauthoryear{Mayer et al.}{2004}]{mayer04} Mayer, L., Quinn, T., Wadsley, J., Stadel, J. 2004, ApJ, 609, 1045

\bibitem[\protect\citeauthoryear{Mayer et al.}{2007}]{mayer07} 
Mayer, L., Lufkin, G., Quinn, T., Wadsley, J. 2007, ApJ, L77

\bibitem[\protect\citeauthoryear{Mej\'ia et al.}{2005}]{mejia05} Mej\'ia, A.C., Durisen, R.H., Pickett, M.K., Cai, K. 2005, ApJ, 619, 1098

\bibitem[\protect\citeauthoryear{Nelson}{2006}]{nelson06} Nelson A.F. 2006, MNRAS, 373, 1039

\bibitem[\protect\citeauthoryear{Nelson et al.}{1998}]{nelson98} Nelson, A.F., Benz, W., Adams, F.C., Arnett, D. 1998, ApJ, 502, 342

\bibitem[\protect\citeauthoryear{Nelson, Benz, Ruzmaikina}{Nelson et al.}{2000}]
{nelson00} Nelson, A.F., Benz, W., Ruzmaikina, T.V. 2000, ApJ, 529, 357

\bibitem[\protect\citeauthoryear{Pickett et al.}{1998}]{pickett98} Pickett, B.K., Cassen, P., Durisen, R.H., Link, R. 1998, ApJ, 504, 468

\bibitem[\protect\citeauthoryear{Pickett et al.}{2000a}]{pickett00a} Pickett, B.K., Cassen, P., Durisen, R.H., Link, R. 2000, ApJ, 529, 1034

\bibitem[\protect\citeauthoryear{Pickett et al.}{2000b}]{pickett00b} Pickett, B.K., Durisen, R.H., Cassen, P., Durisen, R.H., Mej\'ia, A.C. 2000, ApJ, 540, L95

\bibitem[\protect\citeauthoryear{Pickett et al.}{2003}]{pickett03} Pickett, B.K., Mej\'ia, A.C., Durisen, R.H., Cassen, P., Berry, D.K., Link, R.P. 2003, ApJ, 590, 1060

\bibitem[\protect\citeauthoryear{Pickett \& Durisen}{2007}]{pickett07} 
Pickett, M.K., Durisen, R.H. 2007, ApJ, 654, L155

\bibitem[\protect\citeauthoryear{Rafikov}{2005}]{rafikov05} 
Rafikov,  R.R. 2005, ApJ, 621, L69

\bibitem[\protect\citeauthoryear{Rafikov}{2007}]{rafikov07} 
Rafikov,  R.R. 2007, ApJ, 662, 642

\bibitem[\protect\citeauthoryear{Rafikov}{2009}]{rafikov09} 
Rafikov,  R.R. 2009, ApJ, 704, 281

\bibitem[\protect\citeauthoryear{Rice et al.}{2003}]{rice03} Rice, W.K.M., Armitage, P.J., Bate, M.R., Bonnell, I.A. 2003, MNRAS, 339, 1025

\bibitem[\protect\citeauthoryear{Rice et al.}{2005}]{rice05} Rice, W.K.M., Lodato, G., Armitage, P.J. 2005, MNRAS, 364, L56

\bibitem[\protect\citeauthoryear{Stamatellos \& Whitworth}{2008}]{stam08} Stamatellous, D., \& Whitworth, A. 2008, A\& A, 480, 879

\bibitem[\protect\citeauthoryear{Stamatellos \& Whitworth}{2009}]{stam09} Stamatellous, D., \& Whitworth, A. 2009, MNRAS, 392, 413

\bibitem[\protect\citeauthoryear{Vorobyov \& Basu}{2010}]{vb10}
Vorobyov, E.I., Basu, S. 2010, ApJ, 714, L133

\bibitem[\protect\citeauthoryear{Zhu et al.}{2009}]{zhu09} Zhu, Z., Hartmann, L., Gammie, C.F.; McKinney, J.C. 2009, ApJ, 701, 620

\end{thebibliography}

\newpage

\begin{center}
{\large\bf Figure Captions}
\end{center}

{\bf Figure 1.}  Am plot.

\begin{center}
{\large\bf Figure Captions}
\end{center}

{\bf Figure 2.}  Plot of r(t) for the point mass in two simulations. The red curve shows the planet’s position when included at t = 0, the black curve for the planet inserted at t = 10 ORPs.  The diamonds track the semi-major axis for each 0 to 2π interval. Figure 2 shows the planet radius over time for both the t = 0 simulation (red) and the t = 10 ORPs simulation (black). The diamonds represent the semi-major axis for each 0 to 2π interval. As can be seen in Figure 2, for the t = 0 simulation the planet begins to drift inward, presumably due to Type 1 migration, and then moves inward rapidly once the burst of GI activity begins around 3.5 ORPs.

\begin{center}
{\large\bf Figure Captions}
\end{center}

{\bf Figure 3.}  Plot of the eccentricities for each of the 0 to 2π intervals for the t = 0 (red) simulation and t = 10 ORPs (black).

\label{lastpage}

\end{document}

### 2010-06-11 Text/migration.tex
Dick's changes 6/11/10
git-svn-id: http://jaga.astro.indiana.edu/paperrepo/ApJL/Migration2010@65 3b7f7548-5c66-0410-83ab-3997167635fc

@@
 
-Section 2 below presents our numerical methods and initial conditions.  We describe and discuss the simulation results in Section 3, and summarize our main conclusions in Section 4.
+Section 2 below presents our numerical methods and initial conditions.  We describe the simulation results in Section 3, and discussion them in Section 4.
 
@@
 %Scott: A number of things in this paragraph need to verified by further analysis.
-{\it The $t = 0$ Planet Run.} {\bf See note in Latex file.} Figure 1 shows that inclusion of even this modest mass planet ($0.8 \%$ of $M_d$) has a dramatic effect on the burst phase.
+{\it The $t = 0$ Planet Run.} {\bf See note in Latex file.} Figure \ref{fig:Am} shows that inclusion of even this modest mass planet ($0.8 \%$ of $M_d$) has a dramatic effect on the burst phase.
 Without a planet, when all modes grow from imposed noise, it takes 4 ORPs for coherent spiral modes to organize and grow to nonlinear amplitude.
@@
 When the triggered $m = 3$ mode becomes strongly nonlinear at about 3 ORP, planet migration is significantly affected.
-Figure 2 shows the evolution of the planet's radial position.
+Figure \ref{fig:a} shows the evolution of the planet's radial position.
 From 0 to 2.5 ORP, the planet is torqued primarily by its own wake and migrates inward.
@@
 
-{\it Eccentiricity.} Figure 3 shows the evolution of $e$ in the planet simulations.
+{\it Eccentiricity.} Figure \ref{fig:e} shows the evolution of $e$ in the planet simulations.
 In both cases, the planets were inserted with an approximate circular velocity computed by adding the interior disk mass to the stellar point mass.
@@
 However, once there are strong nonlinear interactions with GI modes, $e$ jumps upward.
-In both runs, it appears that interaction with well-established GI activity leads to eccentricities between 0.02 and 0.1 that vary in a chaotic way between these extremes on a short time scale.
+In both runs, it appears that interaction with well-established GI activity leads to eccentricities between 0.02 and 0.1 that vary in a chaotic way between these extremes on orbit period time scales.
 
-{\it Migration Relative to the Disk.} The disk surface density distribution does not vary much in the asymptotic phase.
-So radial motion of the planet in the $t = 10$ planet run is also radial migration relative to the gas.
-During the burst (see, e.g., Boley et al.
-2006), however, the surface density of the disk changes dramatically.
-To verify that the planet is really migrating relative to the background gas disk in the $t = 0$ run, we have compared the planet's radius $\varpi_p(t)$ with the mass inside that radius, expressed in terms of $m_{\rm cyl}$, the fraction of the disk mass interior to the planet.
-Between $t = 3.52$ and $3.89$ ORP, when $\varpi_p$ increases modestly, $m_{\rm cyl}$ increases from about 0.50 to 0.53; between 3.89 and 8.47 ORP, when $\varpi_p$ decreases dramatically, $m_{\rm cyl}$ decreases from 0.53 to only 0.30.
-So, the planet is migrating significantly with respect to the disk mass even while the inner part of the gas disk moves inward during the burst.
-For a comparison measure of the magnitude of the gas motion, $m_{\rm cyl}$ at 23 AU increases from 0.50 to 0.60 from $t =$ 3.89 to 8.47 ORP.
+{\it Migration Relative to the Disk.} The disk surface density distribution does not vary much in the asymptotic phase, so radial motion of the planet in the $t = 10$ planet run also represents radial migration relative to the gas.
+During the burst (e.g., Boley et al.
+2006), the surface density of the disk changes dramatically.
+To verify that the planet is really migrating relative to the background gas disk in the $t = 0$ run, we have compared the evolution of the planet's radial position $\varpi_p(t)$ with the mass inside that radius, expressed in terms of $m_{\rm cyl}$, the fraction of the disk mass interior to the planet.
+Between $t = 3.52$ and $3.89$ ORP, when $\varpi_p$ increases modestly, $m_{\rm cyl}$ increases from about 0.50 to 0.53; between 3.89 and 8.47 ORP, when $\varpi_p$ decreases dramatically, $m_{\rm cyl}$ also decreases dramatically from 0.53 to 0.30.
+So, the planet is migrating significantly with respect to the disk mass even while the inner part of the gas disk is moving inward during the burst.
+For a comparison measure of the magnitude of the gas motion, $m_{\rm cyl}$ at 23 AU increases from 0.50 to 0.60 from $t =$ 3.89 to 8.47 ORP, corresponding to an gas disk inflow rate of $\sim 10^{-5} M_{\odot}/{\rm yr}$.
 
@@
 
+{\bf See notes in Latex file.}
+%Scott and Aaron: This section needs a lot of help.
+
 Check against estimated type I migration without GIs. 
-{\bf Aaron: Can you do this?} Figure 2 suggests that the overall radial drift is not too different, but it is definitely not monotonic and can stall.
-Further work is needed to determine how long the stalls last.
+%Aaron: Can you do this?} 
+Figure \ref{fig:a} suggests that the overall radial drift is not too different, but it is definitely not monotonic and can stall. Further work is needed to determine how long the stalls last.
 
-Generic relationship to Laughlin's work. Just a few sentences -- structure (gaps, embedded masses) affects GI stability and the modes that grow.
+Even in non-self-gravitating disks, the presence of discrete radial structure is known to cause instabilities (Li and Lovelace articles abound 2000, 2001).
+Meschiari \& Laughlin (2008) recently demonstrated that a gap, presumably opened by an embedded planet, can cause global modes to become unstable in self-gravitating disks that are otherwise stable against the development of GIs.
+Here we are dealing with a planet that is not massive enough to open a gap, but its gravitational interaction with the disk also has a strong effect on the growth and onset of GIs.
 
-Discrete mode interactions with the planet seem important -- Wind up and pitch. Surfing.
+Discrete mode interactions with the planet seem important -- Wind up and pitch? Surfing?
 
@@
 
-Probably not a random walk for this mass and this disk. Need to know dependence on planet mass and initial location.
+Not a simple random walk for this mass and this disk. Need to know dependence on planet mass and initial location.
 
-Magnitude of e.
+Magnitude of $e$ is roughly consistent with $c_s/v_K$ and low Mach number spirals.
 
-\section{Summary}
-
-The inclusion of a 1MJup mass in the equilibrium disk initiates GIs and dramatically effects the character of the GIs.
+The inclusion of a 1$M_J$ mass in the equilibrium disk initiates GIs and dramatically effects the character of the GIs.
 GIs can cause both inward and outward planet migration and can cause rapid migration in the burst phase.
@@
 
-\begin{center}
-{\large\bf Figure Captions}
-\end{center}
+\begin{figure}[t]
+\center
+\includegraphics[width=12cm]{../Figures/planeta.pdf}
+\caption{Global amplitudes of nonaxisymmetric density perturbations $A_m$ as a function of time for individual cos($m\phi$) perturbations.
+The formula for computing $A_m$ is equation (15) of Boley et al.
+(2006).
+Here the integrals are done only over the disk outside 15 AU to suppress contributions from edge effects in the inner disk.
+The set of red and black curves correspond to the $t = 0$ and $t = 10$ planet runs, respectively.
+{\bf The radial migration plot is being used as a placeholder.}}
+\label{fig:Am}
+\end{figure}
 
-{\bf Figure 1.}  Am plot.
+\begin{figure}[t]
+\center
+\includegraphics[width=12cm]{../Figures/planeta.pdf}
+\caption{Plot of the planet's radial position as a function of time $\varpi_p(t)$ for the two planet simulations.
+The red and black indicate the curves ($\varpi(t)$), points ($a$), and horizontal scales ($t$) that correspond to the $t = 0$ and $t = 10$ planet runs, respectively.
+The diamonds are the approximate semi-major axes $a$ computed for each 0 to 2$\phi$ change in azimuth of the planet as explained in the text.}
+\label{fig:a}
+\end{figure}
 
-\begin{center}
-{\large\bf Figure Captions}
-\end{center}
+\begin{figure}[t]
+\center
+\includegraphics[width=12cm]{../Figures/planete.pdf}
+\caption{Plot of the eccentricities $e$, computed for each 0 to 2$\phi$ change in azimuth of the planet as explained in the text, as a function of time.
+The red and black indicate the points ($e$) and horizontal scales ($t$) that correspond to the $t = 0$ and $t = 10$ planet runs, respectively.}
+\label{fig:e}
+\end{figure}
 
-{\bf Figure 2.}  Plot of r(t) for the point mass in two simulations.
-The red curve shows the planet’s position when included at t = 0, the black curve for the planet inserted at t = 10 ORPs.
-The diamonds track the semi-major axis for each 0 to 2π interval.
-Figure 2 shows the planet radius over time for both the t = 0 simulation (red) and the t = 10 ORPs simulation (black).
-The diamonds represent the semi-major axis for each 0 to 2π interval.
-As can be seen in Figure 2, for the t = 0 simulation the planet begins to drift inward, presumably due to Type 1 migration, and then moves inward rapidly once the burst of GI activity begins around 3.5 ORPs.
-
-\begin{center}
-{\large\bf Figure Captions}
-\end{center}
-
-{\bf Figure 3.}  Plot of the eccentricities for each of the 0 to 2π intervals for the t = 0 (red) simulation and t = 10 ORPs (black).
+%\begin{figure}[t]
+%\includegraphics{../Figures/name_of_figure.eps}
+%\caption{here's the caption}
+%\label{fig:name}
+%\end{figure}
 

### 2010-06-14 Text/migration.tex
Dick's changes 6/14/10
git-svn-id: http://jaga.astro.indiana.edu/paperrepo/ApJL/Migration2010@66 3b7f7548-5c66-0410-83ab-3997167635fc

@@
 So, the planet is migrating significantly with respect to the disk mass even while the inner part of the gas disk is moving inward during the burst.
-For a comparison measure of the magnitude of the gas motion, $m_{\rm cyl}$ at 23 AU increases from 0.50 to 0.60 from $t =$ 3.89 to 8.47 ORP, corresponding to an gas disk inflow rate of $\sim 10^{-5} M_{\odot}/{\rm yr}$.
+For a comparison measure of the magnitude of the gas motion, $m_{\rm cyl}$ at 23 AU increases from 0.50 to 0.60 from $t =$ 3.89 to 8.47 ORP, corresponding to an gas disk inflow rate of $\sim 10^{-5} M_{\odot}$ yr$^{-1}$.
 
@@
 
-Check against estimated type I migration without GIs. 
-%Aaron: Can you do this?} 
-Figure \ref{fig:a} suggests that the overall radial drift is not too different, but it is definitely not monotonic and can stall. Further work is needed to determine how long the stalls last.
-
 Even in non-self-gravitating disks, the presence of discrete radial structure is known to cause instabilities (Li and Lovelace articles abound 2000, 2001).
@@
 
-Discrete mode interactions with the planet seem important -- Wind up and pitch? Surfing?
+%Scott and Aaron: Can you go over this part carefully and fill in some calculation for the migration rate?
+Figure \ref{fig:a} suggests that the overall radial drift during the first 10 ORP of the $t = 0$ planet run is not too different from what one would get by extrapolating the migration due solely to the planet's wake during the first 2 or 3 ORP, where the average radial migration is about 4$\times1o^{-3}$ AU yr$^{-1}$.
+Although there are no formulae that are fully applicable to our radiatively cooled disk, using equation ???
+from ???
+and parameters from the simulated disk, we estimate ???
+AU yr$^{-1}$ as an analytic prediction.
+While the interaction with a GI-active disk did not seem to result in a drastically different overall migration rate, the direction and magnitude of migration in the GI-active disk is highly variable and seems to depend strongly on interactions of the planet with discrete modes in the disk.
+In the case of the $t = 10$ planet run, the overall radial migration rate is similar but highly erratic.
+The important departures from laminar disk theory are that there are times when migration stalls and that interactions with discrete low-order modes seem to be important.
+The migration is neither a random walk nor monotonic but has a chaotic character due to nonlinear interactions.
+For all these reasons, it will be important in future studies to explore the dependence of the planet's behavior on its mass and on its placement relative to the site of GI-eruptions (for $t = 0$ type runs) and relative to the CR of low-order GI modes (for $t = 10$ type runs).
 
-Stalling at LRs?
-In both cases it is interesting to note that inward migration halts near 17 AU (coincidence???) ; this may be an artifact of the length of the simulations, clearly both simulations should be carried further.
-This radius is interesting because it is the location of the Inner Linblad Resonance (ILR) for the strong m = 2 mode present in the baseline simulation.
-Inside the ILR the mode can not effectively transfer angular momentum between the disk and the planet.
+%Scott and Aaron: This is REALLY speculative. We need analysis here! The following is a modified version of text from Scott's poster mixed with the idea from my dream.
+In both simulations, the inward migration appears to halt near 17 AU.
+This may simply be a coincidence, because the overall average migration rates are similar, the starting positions are essentially the same, and the calculations have a similar duration.
+We plan to extend simulations like these to much longer times in the future.
+However, the final positions of the planets radius happens to be the location of the Inner Lindblad Resonance (ILR) for the strong m = 2 mode present in the fiducial run.
+In fact, this ILR is associated with a surface density enhancement in the disk.
+As the planet approaches this enhancement, it experiences a steeply falling surface density gradient which,
+in a laminar disk, could cause the migration to stall (???). It could also be that, near the ILR, the mode no longer effectively transfers angular momentum from the planet. 
+%The words that follow are from Scott.
+We must DO the analyses of periods, look at the surface density, look at the Q distribution (I doubt the planets are out of the GI active region yet), and think about the physics.
 Further work is necessary to adequately analyze the Fourier modes and their periodicities in the simulations with planets, but this gives a strong preliminary indication that the GI activity is driving the migration.
 Furthermore, where GIs are inactive planet migration is stalled, even Type I migration may be ineffective due to interference from the GIs at the planet’s OLR.
-
-Not a simple random walk for this mass and this disk. Need to know dependence on planet mass and initial location.
+%Is any of this true?
+An alternative but related possibility is that the planet reaches the outer edge of a surface density enhancement, so that the local density gradient is so steep the ILR's of the planet cancel the OLR's of the planet.
+This would probably put the planet right near the ILR of a major mode, because this is where we expect pile ups of material.
+(This came to me in a dream.)
 
@@
 
+%Scott and Aaron: I am at a loss for a good punch line to end with. Help.
 The inclusion of a 1$M_J$ mass in the equilibrium disk initiates GIs and dramatically effects the character of the GIs.

### 2010-06-15 Text/migration.tex
Worked on paper.
	new file:   Text/aastex.cls
	new file:   Text/general.bib
	modified:   Text/migration.tex

git-svn-id: http://jaga.astro.indiana.edu/paperrepo/ApJL/Migration2010@67 3b7f7548-5c66-0410-83ab-3997167635fc

@@
 \usepackage{graphicx}
+\citestyle{aa}
+\usepackage{natbib}
 
@@
 \affil{Department of Astronomy, Indiana University, Bloomington, IN 47405}
-\email{scamicha@astro.indiana.edu}
+\email{scamicha@indiana.edu}
 
@@
 Balbus \& Terquem or Papaloizou ???) has included gravitational interaction of planets with both laminar and turbulent disks.
-However,  the self-gravity of the gas disk has usually been ignored, and the effects of radiative cooling and a detailed equation of state (EOS) have only recently been included (Paardekooper ???).
-Boss (2004 ???) and Mayer et al.
-(2004) examined radial migration of planet-mass fragments in gravitationally unstable disks, but their disks were violently disrupted by fragmentation under conditions (radii $<$ 40 AU, disk-to-star mass ratios $M_d/M_s \sim 0.1$, and stellar mass $M_s \sim 1\;M_{\odot}$) where this probably should not realistically have happened (Rafikov 2005, 2007; Boley et al.
-2006, 2007b; Boley \& Durisen 2008; Forgan et al.
-2009; Cai et al.
-2010).
+However,  the self-gravity of the gas disk has usually been ignored, and the effects of radiative cooling and a detailed equation of state (EOS) have only recently been included \citep{paardekooper2006}.
+\citet{boss2005} and \citet{mayer2004} examined radial migration of planet-mass fragments in gravitationally unstable disks, but their disks were violently disrupted by fragmentation under conditions (radii $<$ 40 AU, disk-to-star mass ratios $M_d/M_s \sim 0.1$, and stellar mass $M_s \sim 1\;M_{\odot}$) where this probably should not realistically have happened \citep{rafikov2005,rafikov2007,boley2006,boley2007b,boley2008,forgan2009,cai2010}.
 More recently, fragmentation into clumps with gas giant or brown dwarf masses has been documented in disks that are relatively massive ($M_d/M_s \sim$ a few tenths) and spatially extended (outer radii $>$ 100 AU) (Krumholz et al.
@@
 The research was conducted with the support of NASA grants from the Origins of Solar Systems  Program (NNG05GN11G and NNX08AK36G).
-{\bf Scott: Add necessary UITS and Teragrid acknowledgments.}
+This material is based upon work supported by the National Science Foundation under Grants No.
+ACI-0338618l, OCI-0451237, OCI-0535258, OCI-0504075, and CNS-0521433.
+Any opinions, findings and conclusions or recommendations expressed in this material are those of the authors and do not necessarily reflect the views of the National Science Foundation.
 
-\begin{thebibliography}{99}
-
-\bibitem[\protect\citeauthoryear{Boley}{2009}]{boley09} 
-Boley,  A.C. 2009, ApJ, 695, L53
-
-\bibitem[\protect\citeauthoryear{Boley \& Durisen}{2006}]{bd06} Boley, A.C., Durisen, R.H. 2006, ApJ, 641, 534
-
-\bibitem[\protect\citeauthoryear{Boley \& Durisen}{2008}]{bd08} Boley, A.C., Durisen, R.H. 2008, ApJ, 685, 1193
-
-\bibitem[\protect\citeauthoryear{Boley et al.}{2006}]{boley06} Boley, A.C., 
-Mej\'ia, A.C., Durisen, R.H., Cai, K., Pickett, M.K. 2006, ApJ, 651, 517
-
-\bibitem[\protect\citeauthoryear{Boley et al.}{2007}]{boley07} Boley, A.C., 
-Durisen, R.H., Nordlund, A., Lord, J. 2007, ApJ, 665, 1254
-
-\bibitem[\protect\citeauthoryear{Boley et al.}{2009}]{boleyetal09} Boley, A.C., 
-Hayfield, T., Mayer, L., Durisen, R.H.  2010, Icarus, in press (arXiv:0909.4543)
-
-\bibitem[\protect\citeauthoryear{Boss}{1984}]{boss84} Boss, A.P. 1984, ApJ, 277, 768
-
-\bibitem[\protect\citeauthoryear{Boss}{1997}]{boss97} Boss, A.P. 1997, Science, 267, 1836
-
-\bibitem[\protect\citeauthoryear{Boss}{1998}]{boss98} Boss, A.P. 1998, ApJ, 503, 923
-
-\bibitem[\protect\citeauthoryear{Boss}{2000}]{boss00} Boss, A.P. 2000, ApJ, 536, L101
-
-\bibitem[\protect\citeauthoryear{Boss}{2001}]{boss01} Boss, A.P. 2001, ApJ, 563, 367
-
-\bibitem[\protect\citeauthoryear{Boss}{2002}]{boss02} Boss, A.P. 2002, ApJ, 576, 462
-
-\bibitem[\protect\citeauthoryear{Boss}{2003}]{boss03} Boss, A.P. 2003, ApJ, 599, 577
-
-\bibitem[\protect\citeauthoryear{Boss}{2004}]{boss04} Boss, A.P. 2004, ApJ, 610, 456
-
-\bibitem[\protect\citeauthoryear{Boss}{2005}]{boss05} Boss, A.P. 2005, ApJ, 629, 535
-
-\bibitem[\protect\citeauthoryear{Boss}{2007}]{boss07} Boss, A.P. 2007, ApJ, 661, L73
-
-\bibitem[\protect\citeauthoryear{Boss}{2009}]{boss09} Boss, A.P. 2009, ApJ, 694, 107
-
-\bibitem[\protect\citeauthoryear{Cai et al.}{2006}]{cai06} Cai, K., Durisen, R.H., Michael, S., Boley, A.C., Mej\'ia, A.C., Pickett, M.K., D'Alessio, P. 2006, ApJ, L149
-
-\bibitem[\protect\citeauthoryear{Cai et al.}{2008}]{cai08} Cai, K., Durisen, R.H., Boley, A.C., Pickett, M.K., Mej\'ia, A.C. 2008, ApJ, 673, 1138
-
-\bibitem[\protect\citeauthoryear{Clarke}{2009}]{clarke09} 
-Clarke, C.J. 2009, MNRAS, 396, 1066
-
-\bibitem[\protect\citeauthoryear{D'Alessio et al.}{2001}]{dalessio01} D'Alessio, P., Calvet, N., Hartmann, L. 2001, ApJ, 553, 321
-
-\bibitem[\protect\citeauthoryear{Dodson-Robinson et al.}{2009}]{dodson09}
-Dodson-Robinson, S. E., Veras, D., Ford, E. B., Beichman, C. A. 2009, ApJ, 707, 79
-
-%\bibitem[\protect\citeauthoryear{Durisen, Mej\'ia, \& Pickett}
-%{Durisen et al.}{2003}]{durisen03} Durisen, R.H., Mej\'ia, A.C.,
-
-\bibitem[\protect\citeauthoryear{Durisen et al.}{2007}]{durisen07} Durisen R.H., Boss, A.P., Mayer, L., Nelson, A.F., Quinn, T., Rice, W.K.M.
-2007, in: Protostars and Planets V, ed.
-B.
-Reipurth, D.
-Jewitt, \& K.
-Keil (Tucson:  Univ.
-of Arizona Press), 607
-
-\bibitem[\protect\citeauthoryear{Durisen et al.}{2008}]{durisen08} 
-Durisen R.H., Hartquis, T.W., \& Pickett M.K., Durisen R.H. 2008, ApSpSci., 317, 3
-
-\bibitem[\protect\citeauthoryear{Forgan et al.}{2009}]{forg09} Forgan, D., Rice, K., Stamatellos, D., Whitworth, A. 2009, MNRAS 394, 882
-
-\bibitem[\protect\citeauthoryear{Gammie}{2001}]{gammie01} Gammie, C.F. 2001, ApJ, 553, 174
-
-\bibitem[\protect\citeauthoryear{Johnson \& Gammie}{2003}]{johnson03} 
-Johnson, B.M., Gammie, C.F. 2003, ApJ, 597, 131
-
-\bibitem[\protect\citeauthoryear{Haisch et al.}{2001}]{haisch01} 
-Haisch, K.E., Jr., Lada, E.A., Lada, C.J. 2001, ApJ, 553, L153
-
-\bibitem[\protect\citeauthoryear{Hayfield et al.}{2010}]{hayfield10}
-Hayfield, T., Mayer, L., Wadsley, J., Boley, A. C. 2010, arXiv:1003.2594
-
-\bibitem[\protect\citeauthoryear{Mayer et al.}{2002}]{mayer02} Mayer, L., Quinn, T. Wadsley, J., Stadel, J. 2002, Science, 298, 1756
-
-\bibitem[\protect\citeauthoryear{Mayer et al.}{2004}]{mayer04} Mayer, L., Quinn, T., Wadsley, J., Stadel, J. 2004, ApJ, 609, 1045
-
-\bibitem[\protect\citeauthoryear{Mayer et al.}{2007}]{mayer07} 
-Mayer, L., Lufkin, G., Quinn, T., Wadsley, J. 2007, ApJ, L77
-
-\bibitem[\protect\citeauthoryear{Mej\'ia et al.}{2005}]{mejia05} Mej\'ia, A.C., Durisen, R.H., Pickett, M.K., Cai, K. 2005, ApJ, 619, 1098
-
-\bibitem[\protect\citeauthoryear{Nelson}{2006}]{nelson06} Nelson A.F. 2006, MNRAS, 373, 1039
-
-\bibitem[\protect\citeauthoryear{Nelson et al.}{1998}]{nelson98} Nelson, A.F., Benz, W., Adams, F.C., Arnett, D. 1998, ApJ, 502, 342
-
-\bibitem[\protect\citeauthoryear{Nelson, Benz, Ruzmaikina}{Nelson et al.}{2000}]
-{nelson00} Nelson, A.F., Benz, W., Ruzmaikina, T.V. 2000, ApJ, 529, 357
-
-\bibitem[\protect\citeauthoryear{Pickett et al.}{1998}]{pickett98} Pickett, B.K., Cassen, P., Durisen, R.H., Link, R. 1998, ApJ, 504, 468
-
-\bibitem[\protect\citeauthoryear{Pickett et al.}{2000a}]{pickett00a} Pickett, B.K., Cassen, P., Durisen, R.H., Link, R. 2000, ApJ, 529, 1034
-
-\bibitem[\protect\citeauthoryear{Pickett et al.}{2000b}]{pickett00b} Pickett, B.K., Durisen, R.H., Cassen, P., Durisen, R.H., Mej\'ia, A.C. 2000, ApJ, 540, L95
-
-\bibitem[\protect\citeauthoryear{Pickett et al.}{2003}]{pickett03} Pickett, B.K., Mej\'ia, A.C., Durisen, R.H., Cassen, P., Berry, D.K., Link, R.P. 2003, ApJ, 590, 1060
-
-\bibitem[\protect\citeauthoryear{Pickett \& Durisen}{2007}]{pickett07} 
-Pickett, M.K., Durisen, R.H. 2007, ApJ, 654, L155
-
-\bibitem[\protect\citeauthoryear{Rafikov}{2005}]{rafikov05} 
-Rafikov,  R.R. 2005, ApJ, 621, L69
-
-\bibitem[\protect\citeauthoryear{Rafikov}{2007}]{rafikov07} 
-Rafikov,  R.R. 2007, ApJ, 662, 642
-
-\bibitem[\protect\citeauthoryear{Rafikov}{2009}]{rafikov09} 
-Rafikov,  R.R. 2009, ApJ, 704, 281
-
-\bibitem[\protect\citeauthoryear{Rice et al.}{2003}]{rice03} Rice, W.K.M., Armitage, P.J., Bate, M.R., Bonnell, I.A. 2003, MNRAS, 339, 1025
-
-\bibitem[\protect\citeauthoryear{Rice et al.}{2005}]{rice05} Rice, W.K.M., Lodato, G., Armitage, P.J. 2005, MNRAS, 364, L56
-
-\bibitem[\protect\citeauthoryear{Stamatellos \& Whitworth}{2008}]{stam08} Stamatellous, D., \& Whitworth, A. 2008, A\& A, 480, 879
-
-\bibitem[\protect\citeauthoryear{Stamatellos \& Whitworth}{2009}]{stam09} Stamatellous, D., \& Whitworth, A. 2009, MNRAS, 392, 413
-
-\bibitem[\protect\citeauthoryear{Vorobyov \& Basu}{2010}]{vb10}
-Vorobyov, E.I., Basu, S. 2010, ApJ, 714, L133
-
-\bibitem[\protect\citeauthoryear{Zhu et al.}{2009}]{zhu09} Zhu, Z., Hartmann, L., Gammie, C.F.; McKinney, J.C. 2009, ApJ, 701, 620
-
-\end{thebibliography}
+\bibliographystyle{apj}
+\bibliography{general}
 

### 2010-08-10 Text/migration.tex
First pass over Dick's draft. Need more bibliography work and text tweaking.
	new file:   .gitignore
	modified:   Text/migration.tex

git-svn-id: http://jaga.astro.indiana.edu/paperrepo/ApJL/Migration2010@68 3b7f7548-5c66-0410-83ab-3997167635fc

@@
 Characterization of migration in gravitationally unstable disks is necessary to understand the fate of protoplanets formed by disk instability.
-As part of a larger study, we are using a 3D radiative hydrodynamics code to investigate how an embedded gas giant planet interacts with a gas disk that is massive enough to experience gravitational instabilities (GIs).
-This Letter presents preliminary results from simulations with a Jupiter-mass planet placed in orbit at 25 AU within a 0.13 $M_{\odot}$ disk.
-The disk spans 5 to 40 AU around a 1 $M_{\odot}$ star and is initially only marginally stable.
+As part of a larger study, we are using a 3D radiative hydrodynamics code to investigate how an embedded gas giant planet interacts with a gas disk that is massive enough to experience {\bf any mass object will experience GIs rewrite} gravitational instabilities (GIs).
+This Letter presents preliminary results from simulations with a Jupiter-mass planet placed in orbit at 25 AU within a 0.14 $M_{\odot}$ disk.
+The disk spans 5 to 40 AU around a 1 $M_{\odot}$ star and is initially marginally stable.
 In one simulation, the planet is inserted prior to the initial eruption of GIs; in another, it is inserted only after the disk has settled into a quasi-steady GI-active state, where heating by GIs roughly balances radiative cooling.
@@
 2007a) and integrates an energy equation that includes $PdV$ work, net heating or cooling due to radiative flux divergence, and heating by artificial bulk viscosity.
-Calculations are done on a uniform cylindrical grid with reflection symmetry about the disk midplane and a resolution ($\varpi$,$\phi$,z) = (512,512,64).
+Calculations are done on a uniform cylindrical grid with reflection symmetry about the disk midplane and a resolution {\bf starting resolution is different} ($\varpi$,$\phi$,z) = (512,512,64).
 The $z$-axis is the rotation axis of the disk.
@@
 (2007b), where the optically thick monochromatic flux in the $\varpi$ and $\phi$-directions is computed by flux-limited radiative diffusion and where the radiative transport of energy in the $z$-direction is solved using a one-ray discrete ordinate method in both optically thin and thick regions.
-Although the central star remains fixed at the grid center, we account for acceleration of the reference frame by the planet and by the disk via indirect potentials, as in Michael \& Durisen (2010), and, overall, the planet integration is done using methods similar to those in Nelson et al.
+Although the central star remains fixed at the grid center, we account for acceleration of the reference frame by the planet and by the disk via indirect potentials, as in Michael \& Durisen (2010), and, overall, the planet integration is done using methods similar to those in {\bf integration is done with a verlet integrator as in hut95, yoshida90, channell90, tuckerman92.
+indirect terms are treated as in nelson} Nelson et al.
 (2000).
@@
 The model disk, based on an equlibrium model in Pickett et al.
-(2003), orbits a 1 $M_{\odot}$ star and has a mass $M_d = 0.13 M_{\odot}$, inner and outer radii at 5 and 40 AU, and an initial surface density distribution $\Sigma \sim \varpi^{-1}$.
+(2003), orbits a 1 $M_{\odot}$ star and has a mass $M_d = 0.14 M_{\odot}$, inner and outer radii at 5 and 40 AU, and an initial surface density distribution $\Sigma \sim \varpi^{-1}$.
 The time unit of one ORP (= outer rotation period) is defined as the rotation period of the initial disk at $\varpi \approx$ 32 AU, or about 250 yr.
-The disk is initially located between radial grid zones ???
-30 ???
-and 240 and is close to isentropic, which results in a Toomre-$Q$ distribution with a marginally unstable (see Durisen et al.
+The disk is initially located between radial grid zones 30 and 240 and is close to isentropic, which results in a Toomre-$Q$ distribution with a marginally unstable (see Durisen et al.
 2007) minimum value of ???
@@
 AU).
-The computational grid extends radially to 512 zones to accommodate expansion of the outer disk once GIs become nonlinear.
-There is an inner cylindrical hole in the grid with a radius of 13 radial zones.
-To seed nonaxisymmetry, the density is given an initial ???
-\% random cell-to-cell perturbation.
+The computational grid is extended radially to 512 zones to accommodate expansion of the outer disk once GIs become nonlinear.
+There is an inner cylindrical hole in the grid with a radius of 12 radial zones {\bf brief description of hole needed}.
+To seed nonaxisymmetry, the density is given an initial 0.01 \% random cell-to-cell perturbation.
 
@@
 ORP without a planet.
-In the second simulation, which we call the $t = 0$ planet run, a one Jupiter-mass ($M_J$) planet with a roughly circular midplane orbit is included in the initial $t = 0$ ORP equilibrium disk at 25 AU.
-In the third simulation, which we call the $t = 10$ planet run, a 1 $M_J$ planet with a similar orbit is inserted into the fiducial run at $t = 10$ ORP.
-These three simulations are part of a much larger suite, to be described elsewhere, in which the planet mass and starting locations are varied.
+In the second simulation, which we call the T0 planet run, a one Jupiter-mass ($M_J$) planet with a roughly circular midplane orbit is included in the initial $t = 0$ ORP equilibrium disk at 25 AU.
+In the third simulation, which we call the T10 planet run, a 1 $M_J$ planet with a similar orbit is inserted into the fiducial run at $t = 10$ ORP.
+These three simulations are part of a much larger suite, to be described elsewhere, in which the planet mass {\bf and starting locations} are varied.
 We estimate crude semi-major axes $a$ and eccentricities $e$ from the planet's $\varpi(t)$ by finding at the maximum and minimum radii, $\varpi_{\rm max}$ and $\varpi_{\rm min}$, over each interval that the planet's $\phi$ changes by 2$\pi$ (one ``orbit'') and then using $a = (\varpi_{\rm max}+\varpi_{\rm min})/2$ and $e = (\varpi_{\rm max}-\varpi_{\rm min})/(\varpi_{\rm max}+\varpi_{\rm min})$.
@@
 Here we are dealing with a planet that is not massive enough to open a gap, but its gravitational interaction with the disk also has a strong effect on the growth and onset of GIs.
+{\bf say more about gap opening criteria and whether they apply here}
 
 %Scott and Aaron: Can you go over this part carefully and fill in some calculation for the migration rate?
-Figure \ref{fig:a} suggests that the overall radial drift during the first 10 ORP of the $t = 0$ planet run is not too different from what one would get by extrapolating the migration due solely to the planet's wake during the first 2 or 3 ORP, where the average radial migration is about 4$\times1o^{-3}$ AU yr$^{-1}$.
+Figure \ref{fig:a} suggests that the overall radial drift during the first 10 ORP of the $t = 0$ planet run is not too different from what one would get by extrapolating the migration due solely to the planet's wake during the first 2 or 3 ORP, where the average radial migration is about 4$\times10^{-3}$ AU yr$^{-1}$.
 Although there are no formulae that are fully applicable to our radiatively cooled disk, using equation ???
@@
 Any opinions, findings and conclusions or recommendations expressed in this material are those of the authors and do not necessarily reflect the views of the National Science Foundation.
+{\bf need TG acks}
 

### 2010-08-25 Text/migration.tex
Dick changes 8/24/10
git-svn-id: http://jaga.astro.indiana.edu/paperrepo/ApJL/Migration2010@69 3b7f7548-5c66-0410-83ab-3997167635fc

@@
 Characterization of migration in gravitationally unstable disks is necessary to understand the fate of protoplanets formed by disk instability.
-As part of a larger study, we are using a 3D radiative hydrodynamics code to investigate how an embedded gas giant planet interacts with a gas disk that is massive enough to experience {\bf any mass object will experience GIs rewrite} gravitational instabilities (GIs).
+As part of a larger study, we are using a 3D radiative hydrodynamics code to investigate how an embedded gas giant planet interacts with a gas disk that undergoes gravitational instabilities (GIs).
 This Letter presents preliminary results from simulations with a Jupiter-mass planet placed in orbit at 25 AU within a 0.14 $M_{\odot}$ disk.
@@
 When the planet is present from the beginning, its own wake stimulates growth of a particular global mode with which it strongly interacts, and the planet plunges inward six AU in about 10$^3$ years.
-In both cases, there are times when the planet's radial motion is slow and random.
-At other times, when the planet interacts with a strong spiral mode near the mode's corotation radius, migration both inward and outward can be relatively rapid, covering several AUs over hundreds of years.
+In both cases with embedded planets, there are times when the planet's radial motion is slow and varies in direction.
+At other times, when the planet interacts with a strong spiral mode, migration both inward and outward can be relatively rapid, covering several AUs over hundreds of years.
 {\bf See note in Latex file.}
@@
 2007a) and integrates an energy equation that includes $PdV$ work, net heating or cooling due to radiative flux divergence, and heating by artificial bulk viscosity.
-Calculations are done on a uniform cylindrical grid with reflection symmetry about the disk midplane and a resolution {\bf starting resolution is different} ($\varpi$,$\phi$,z) = (512,512,64).
+Calculations are done on a uniform cylindrical grid with reflection symmetry about the disk midplane and a grid size ($\varpi$,$\phi$,z) = (512,512,64).
+{\bf See note in Latex file.} 
+%Scott: The resolution stays the same.
+The grid size varies, but this is unimportant because initially the outer 256 radial zones are empty.
+It is simplest to just give the grid size.
+So I did change "resolution" to grid size here.
 The $z$-axis is the rotation axis of the disk.
@@
 (2007b), where the optically thick monochromatic flux in the $\varpi$ and $\phi$-directions is computed by flux-limited radiative diffusion and where the radiative transport of energy in the $z$-direction is solved using a one-ray discrete ordinate method in both optically thin and thick regions.
-Although the central star remains fixed at the grid center, we account for acceleration of the reference frame by the planet and by the disk via indirect potentials, as in Michael \& Durisen (2010), and, overall, the planet integration is done using methods similar to those in {\bf integration is done with a verlet integrator as in hut95, yoshida90, channell90, tuckerman92.
-indirect terms are treated as in nelson} Nelson et al.
+Although the central star remains fixed at the grid center, we account for acceleration of the reference frame by the planet and by the disk via indirect potentials, as in Michael \& Durisen (2010).
+The planet integration is done with a Verlet integrator (e.g., Hut 1995), and the indirect potential terms are treated as in Nelson et al.
 (2000).

### 2010-09-15 Text/migration.tex
Dick's changes 9/15/10
git-svn-id: http://jaga.astro.indiana.edu/paperrepo/ApJL/Migration2010@71 3b7f7548-5c66-0410-83ab-3997167635fc

@@
 {\bf See note in Latex file.} 
-%Scott: File in some references you know now. Aaron: Suggestions on references?
-Migration of gas giant planets due to interactions with a circumstellar gas disk probably plays a major role in defining the architecture of planetary systems (e.g., ???).
-Work on migration (see review by ???
-Balbus \& Terquem or Papaloizou ???) has included gravitational interaction of planets with both laminar and turbulent disks.
-However,  the self-gravity of the gas disk has usually been ignored, and the effects of radiative cooling and a detailed equation of state (EOS) have only recently been included \citep{paardekooper2006}.
+%Scott: Fill in some references you know now. Aaron: Suggestions on references?
+Migration of gas giant planets due to interactions with a circumstellar gas disk probably plays a major role in defining the architecture of planetary systems.
+Work on migration \citep[see review by][]{papaloizou2007} has included gravitational interaction of planets with both laminar and turbulent disks.
+However,  the self-gravity of the gas disk has usually been ignored, and the effects of radiative cooling and a detailed equation of state (EOS) have only recently been included \citep[e.g.,][]{paardekooper2006,paardekooper2010,paardekooper2010b}.
 \citet{boss2005} and \citet{mayer2004} examined radial migration of planet-mass fragments in gravitationally unstable disks, but their disks were violently disrupted by fragmentation under conditions (radii $<$ 40 AU, disk-to-star mass ratios $M_d/M_s \sim 0.1$, and stellar mass $M_s \sim 1\;M_{\odot}$) where this probably should not realistically have happened \citep{rafikov2005,rafikov2007,boley2006,boley2007b,boley2008,forgan2009,cai2010}.
-More recently, fragmentation into clumps with gas giant or brown dwarf masses has been documented in disks that are relatively massive ($M_d/M_s \sim$ a few tenths) and spatially extended (outer radii $>$ 100 AU) (Krumholz et al.
-2007; Stamatellos \& Whitworth 2007, 2009, Boley 2009; Hayfield et al.
-2010), where fragmentation is expected from semi-analytic arguments (e.g., Clarke ???, Rafikov 2009, Dobson-Robinson et al.
-2009).
-The fate of the clumps depends in part on their radial migration, which is a chaotic and messy affair in a fragmenting disk (e.g., Boley 2009, Boley et al.
-2010, Vorobyov \& Basu 2010; Boley \& Durisen 2010).
-The occurrence of gravitational instabilities (GIs) may be episodic (Vorobyov \& Basu 2005, 2006).
+More recently, fragmentation into clumps with gas giant or brown dwarf masses has been documented in numerical simulations of disks that are relatively massive ($M_d/M_s \sim$ a few tenths) and spatially extended (outer radii $>$ 50 AU) \citep{krumholz2007,stamatellos2007,stamatellos2009,boley2009,boley2010}, where fragmentation is expected from semi-analytic arguments \citep[e.g.,][]{clarke2009,rafikov2009,dodson2009}.
+The fate of the clumps depends in part on their radial migration, which is a chaotic and messy affair in a fragmenting disk \citep[e.g.,][]{boley2009,boley2010,vorobyov2010,boleydurisen2010}.
+The occurrence of gravitational instabilities (GIs) may also be episodic \citep[e.g.,][]{vorobyov2006,vorobyov2010b,zhu2010}.
 Clumps that survive and contract to the dimensions of young planets can later find themselves in a disk that erupts again into GI activity.
@@
 To improve our understanding of how planets migrate in GI-active disks that are not fragmenting, we have begun a systematic study, using numerical 3D radiative hydrodynamics, where we investigate the effects on both the disk and the planet of inserting a planet-mass object into disks susceptible to GIs.
-Using techniques deployed in earlier research by our group (Pickett et al.
-2003; Mej\'ia et al.
-2005; Cai et al.
-2006, 2008; Boley et al 2006, 2007b), we can identify the dominant spiral waves in a simulation and analyze how the waves interact with the planet's motion.
+Using techniques deployed in earlier research by our group \citep{pickett2003, mejia2005,cai2006,cai2008,boley2006,boley2007b,michael2010}, we can identify the dominant spiral waves in a simulation and analyze how the waves interact with the planet's motion.
 Our goal is to determine both the effect of giant planets on GIs and the effect of GIs on planet migration.
-Because GIs are sensitive to radiative physics, we use a well-tested radiative scheme (Boley et al.
-2007b) and realistic opacities (D'Alessio et al.
-2001).
+Because GIs are sensitive to radiative physics, we use a well-tested radiative scheme \citep{boley2007b} and realistic opacities \citep{dalessio2001}.
 
-Section 2 below presents our numerical methods and initial conditions.  We describe the simulation results in Section 3, and discussion them in Section 4.
+Section 2 below presents our numerical methods and initial conditions.  We describe the simulation results in Section 3, and discuss them in Section 4.
 
@@
 %Scott and Aaron: Verify that everything I say here is correct.
-The CHYMERA (Computational HYdrodynamics with MultiplE Radiation Algorithms) code (Boley et al.
-2007b) is a second-order, explicit, Eulerian scheme on a 3D cylindrical grid.
-The code uses a realistic equation of state for H$_2$ (Boley et al.
-2007a) and integrates an energy equation that includes $PdV$ work, net heating or cooling due to radiative flux divergence, and heating by artificial bulk viscosity.
+The CHYMERA (Computational HYdrodynamics with MultiplE Radiation Algorithms) code \citep{boley2007b} is a second-order, explicit, Eulerian scheme on a 3D cylindrical grid.
+The code uses a realistic equation of state for H$_2$ \citep{boley2007} and integrates an energy equation that includes $PdV$ work, net heating or cooling due to radiative flux divergence, and heating by artificial bulk viscosity.
 Calculations are done on a uniform cylindrical grid with reflection symmetry about the disk midplane and a grid size ($\varpi$,$\phi$,z) = (512,512,64).
@@
 The grid size varies, but this is unimportant because initially the outer 256 radial zones are empty.
-It is simplest to just give the grid size.
+It is simplest to just give the eventual grid size.
 So I did change "resolution" to grid size here.
@@
 The large number of azimuthal zones is necessary to resolve the planet's Hill sphere and the planet's wake.
-These simulations utilize the radiative cooling scheme developed and tested in Boley et al.
-(2007b), where the optically thick monochromatic flux in the $\varpi$ and $\phi$-directions is computed by flux-limited radiative diffusion and where the radiative transport of energy in the $z$-direction is solved using a one-ray discrete ordinate method in both optically thin and thick regions.
-Although the central star remains fixed at the grid center, we account for acceleration of the reference frame by the planet and by the disk via indirect potentials, as in Michael \& Durisen (2010).
-The planet integration is done with a Verlet integrator (e.g., Hut 1995), and the indirect potential terms are treated as in Nelson et al.
-(2000).
-The Rosseland mean and Planck mean opacities and molecular weights in our simulations are the same as those in Boley et al.
-(2006, 2007b), except that we correct the D'Alessio et al.
-(2001) gas mean molecular weight to $\mu$ = 2.39.
+These simulations utilize the radiative cooling scheme developed and tested in \citet{boley2007b}, where the optically thick monochromatic flux in the $\varpi$ and $\phi$-directions is computed by flux-limited radiative diffusion and where the radiative transport of energy in the $z$-direction is solved using a one-ray discrete ordinate method in both optically thin and thick regions.
+Although the central star remains fixed at the grid center, we account for acceleration of the reference frame by the planet and by the disk via indirect potentials, as in \citet{michael2010}.
+The planet integration is done with a Verlet integrator \citep[e.g.,][]{hut1995}, and the indirect potential terms are treated as in \citet{nelson2000a}.
+The Rosseland mean and Planck mean opacities and molecular weights in our simulations are the same as those in \citet{boley2006,boley2007b}, except that we correct the \citet{dalessio2001} gas mean molecular weight to $\mu$ = 2.39.
 
@@
 
-The model disk, based on an equlibrium model in Pickett et al.
-(2003), orbits a 1 $M_{\odot}$ star and has a mass $M_d = 0.14 M_{\odot}$, inner and outer radii at 5 and 40 AU, and an initial surface density distribution $\Sigma \sim \varpi^{-1}$.
+The model disk, based on an equlibrium model in \citet{pickett2003}, orbits a 1 $M_{\odot}$ star and has a mass $M_d = 0.14 M_{\odot}$, inner and outer radii at 5 and 40 AU, and an initial surface density distribution $\Sigma \sim \varpi^{-1}$.
 The time unit of one ORP (= outer rotation period) is defined as the rotation period of the initial disk at $\varpi \approx$ 32 AU, or about 250 yr.
-The disk is initially located between radial grid zones 30 and 240 and is close to isentropic, which results in a Toomre-$Q$ distribution with a marginally unstable (see Durisen et al.
-2007) minimum value of ???
+The disk is initially located between radial grid zones 30 and 240 and is close to isentropic, which results in a Toomre-$Q$ distribution with a marginally unstable \citep[see ][]{durisen2007} minimum value of ???
 at radial grid zone ???
@@
 AU).
-The computational grid is extended radially to 512 zones to accommodate expansion of the outer disk once GIs become nonlinear.
-There is an inner cylindrical hole in the grid with a radius of 12 radial zones {\bf brief description of hole needed}.
+The computational grid extends radially to 512 zones to accommodate expansion of the outer disk once GIs become nonlinear.
+There is an inner cylindrical hole in the grid with a radius of 12 radial zones {\bf Scott Comment: brief description of hole needed.
+Dick's reply in Latex file}.
+%Scott: So add one!
 To seed nonaxisymmetry, the density is given an initial 0.01 \% random cell-to-cell perturbation.
@@
 This Letter presents three simulations.
-The first, which we call the fiducial run, goes from time $t = 0$ to ???
+The first, which we call the {\sl fiducial run}, goes from time $t = 0$ to ???
 20 ???
 ORP without a planet.
-In the second simulation, which we call the T0 planet run, a one Jupiter-mass ($M_J$) planet with a roughly circular midplane orbit is included in the initial $t = 0$ ORP equilibrium disk at 25 AU.
-In the third simulation, which we call the T10 planet run, a 1 $M_J$ planet with a similar orbit is inserted into the fiducial run at $t = 10$ ORP.
-These three simulations are part of a much larger suite, to be described elsewhere, in which the planet mass {\bf and starting locations} are varied.
-We estimate crude semi-major axes $a$ and eccentricities $e$ from the planet's $\varpi(t)$ by finding at the maximum and minimum radii, $\varpi_{\rm max}$ and $\varpi_{\rm min}$, over each interval that the planet's $\phi$ changes by 2$\pi$ (one ``orbit'') and then using $a = (\varpi_{\rm max}+\varpi_{\rm min})/2$ and $e = (\varpi_{\rm max}-\varpi_{\rm min})/(\varpi_{\rm max}+\varpi_{\rm min})$.
-These orbital elements are only intended to be indicative; there are strong orbit to orbit variations in these $a$ and $e$.
+In the second simulation, which we call the {\sl t = 0 planet run}, a one Jupiter-mass ($M_J$) planet with a roughly circular midplane orbit is included in the initial $t = 0$ ORP equilibrium disk at 25 AU.
+In the third simulation, which we call the {\sl t = 10 planet run}, a 1 $M_J$ planet with a similar orbit is inserted into the fiducial run at $t = 10$ ORP.
+These three simulations are part of a much larger suite, to be described elsewhere, in which the planet mass and starting locations are varied.
+We estimate crude semi-major axes $a$ and eccentricities $e$ from the planet's $\varpi(t)$ by using at the maximum and minimum radii, $\varpi_{\rm max}$ and $\varpi_{\rm min}$, attained by the planet over each interval that the planet's $\phi$-coordinate changes by 2$\pi$ (one ``orbit'').
+Then, for that orbit, $a = (\varpi_{\rm max}+\varpi_{\rm min})/2$ and $e = (\varpi_{\rm max}-\varpi_{\rm min})/(\varpi_{\rm max}+\varpi_{\rm min})$.
+These parameters are only indicative; there are strong orbit to orbit variations in $a$ and $e$.
 
@@
 
-{\it The Fiducial Run.} As in Mej\'ia et al (2005), the fiducial run exhibits four distinct phases: initial {\sl axisymmetric cooling}, the onset of GIs in a well-defined {\sl burst}, a {\sl transition}, and the {\sl asymptotic} establishment of quasi-steady chaotic GI-activity with an overall balance of heating and cooling.
-There is no fragmentation because this disk has a long cooling time (see Gammie 2001; Boley et al.
-2006, 2007b).
+{\it The Fiducial Run.} As in \citet{mejia2005}, the fiducial run exhibits four distinct phases: initial {\sl axisymmetric cooling}, the onset of GIs in a well-defined {\sl burst} in low-order modes, a {\sl transition} to more high-order nonaxisymmetric structure, and the {\sl asymptotic} establishment of quasi-steady chaotic GI-activity with an overall balance of heating and cooling.
+There is no fragmentation because this disk has a long cooling time \citep{gammie2001,boley2006,boley2007b}.
 
-%Scott: A number of things in this paragraph need to verified by further analysis.
+%Scott: A number of important assertions in this paragraph need to verified by further analysis.
 {\it The $t = 0$ Planet Run.} {\bf See note in Latex file.} Figure \ref{fig:Am} shows that inclusion of even this modest mass planet ($0.8 \%$ of $M_d$) has a dramatic effect on the burst phase.
@@
 Modes of cos($m\phi$) symmetry with $m =$ 5 and 6 dominate.
-An $m = 3$ (three-armed) spiral also grows somewhat more slowly, but lags the other modes in time.
+An $m = 3$ (three-armed) spiral also grows, but somewhat more slowly, and lags the other modes in time.
 In the $t = 0$ planet run, the planet develops an organized wake within about an ORP in which $m = 3$ is a dominant component ($\sim 5$ \% global density perturbation).
-This nonlinear seeding causes $m = 3$ to dominates the GI burst, which also occurs $\sim$ 2 ORPs earlier.
+This nonlinear seeding causes $m = 3$ to dominate the GI burst, which also occurs $\sim$ 2 ORPs earlier.
 Because of our initial placement of the planet inside but close to the $Q$-minimum, the corotation radius (CR) of the $m = 3$ mode is fairly close to that of the planet's orbit period.
@@
 From an analysis of periodicities present in the gas disk between 8 and 12 ORP, the planet lies a few AU inside the inner Lindblad resonance of a strong $m = 2$ mode with CR at ???
-AU in the GI-active disk.
+28 ???
+AU in the GI-active disk in this early part of the asymptotic phase.
 
-%Scott: Even more work is needed to verify what is said here. Some of what I wrote is pure speculation and needs to be checked.
+%Scott: Even more work is needed to verify (or refute) what is said here. Some of what I wrote is pure speculation and needs to be checked.
 {\it The $t = 10$ Planet Run.} {\bf See note in Latex file.} The motion of the planet in this case is more difficult to interpret.
@@
 Animations of the simulation show there is a complex interaction between the planet, its spiral wake, and the global spiral arms caused by the instabilities.
-Analysis of images and periodicities suggests that, between 15 and 19 ORP, the planet may have passed through an arm of an $m =$ ???
+Analysis of images and periodicities suggests that, between 15 and 19 ORP, the planet may have passed through one arm of an $m =$ ???
 pattern with CR at ???
@@
 It is possible that the planet actually stimulated this particular mode, in manner similar to what occurred in the $t = 0$ planet run, because this $m =$ ???
-periodicity is absent over the same time interval in the fiducial run.
+periodicity is absent over a similar time interval in the fiducial run.
 
-{\it Eccentiricity.} Figure \ref{fig:e} shows the evolution of $e$ in the planet simulations.
+{\it Eccentricity.} Figure \ref{fig:e} shows the evolution of $e$ in the planet simulations.
 In both cases, the planets were inserted with an approximate circular velocity computed by adding the interior disk mass to the stellar point mass.
 The presence of GI activity in the $t = 10$ planet run immediately increases $e$ to $\sim 0.08$.
-In the $t = 0$ planet run, the modest initial $e$ decreases, as it should (ref ???), during the 1.5 ORP when it is migrating only due to its own wake.
+In the $t = 0$ planet run, the modest initial $e$ decreases during the 1.5 ORP when it is migrating only due to its own wake.
 However, once there are strong nonlinear interactions with GI modes, $e$ jumps upward.
 In both runs, it appears that interaction with well-established GI activity leads to eccentricities between 0.02 and 0.1 that vary in a chaotic way between these extremes on orbit period time scales.
+The magnitude of $e$ is roughly consistent with the ratio of the sound speed to orbital speed of the disk and the modest values for the Mach numbers of the GI spirals \citep{boley2006a,boley2008}.
 
 {\it Migration Relative to the Disk.} The disk surface density distribution does not vary much in the asymptotic phase, so radial motion of the planet in the $t = 10$ planet run also represents radial migration relative to the gas.
-During the burst (e.g., Boley et al.
-2006), the surface density of the disk changes dramatically.
+During the burst \citep[e.g.,][]{boley2006}, the surface density of the disk changes dramatically.
 To verify that the planet is really migrating relative to the background gas disk in the $t = 0$ run, we have compared the evolution of the planet's radial position $\varpi_p(t)$ with the mass inside that radius, expressed in terms of $m_{\rm cyl}$, the fraction of the disk mass interior to the planet.
@@
 So, the planet is migrating significantly with respect to the disk mass even while the inner part of the gas disk is moving inward during the burst.
-For a comparison measure of the magnitude of the gas motion, $m_{\rm cyl}$ at 23 AU increases from 0.50 to 0.60 from $t =$ 3.89 to 8.47 ORP, corresponding to an gas disk inflow rate of $\sim 10^{-5} M_{\odot}$ yr$^{-1}$.
+As a measure of the gas motion, $m_{\rm cyl}$ at 23 AU increases from 0.50 to 0.60 from $t =$ 3.89 to 8.47 ORP, corresponding to an average gas disk inflow rate of $\sim 10^{-5} M_{\odot}$ yr$^{-1}$.
 
@@
 Even in non-self-gravitating disks, the presence of discrete radial structure is known to cause instabilities (Li and Lovelace articles abound 2000, 2001).
-Meschiari \& Laughlin (2008) recently demonstrated that a gap, presumably opened by an embedded planet, can cause global modes to become unstable in self-gravitating disks that are otherwise stable against the development of GIs.
+Meschiari \& Laughlin (2008) demonstrated that a gap, presumably opened by an embedded planet, can cause global modes to become unstable in self-gravitating disks that are otherwise stable against the development of GIs.
 Here we are dealing with a planet that is not massive enough to open a gap, but its gravitational interaction with the disk also has a strong effect on the growth and onset of GIs.
-{\bf say more about gap opening criteria and whether they apply here}
+{\bf Scott Comment: say more about gap opening criteria and whether they apply here.
+Dick's reply in tex file.}
+%Scott: Do it. I am not sure what you have in mind. Remember, however, that we are already at about maximum length for Letter.
 
@@
 Figure \ref{fig:a} suggests that the overall radial drift during the first 10 ORP of the $t = 0$ planet run is not too different from what one would get by extrapolating the migration due solely to the planet's wake during the first 2 or 3 ORP, where the average radial migration is about 4$\times10^{-3}$ AU yr$^{-1}$.
-Although there are no formulae that are fully applicable to our radiatively cooled disk, using equation ???
+Although there are no simple formulae that are fully applicable to our radiatively cooled disk, using equation ???
 from ???
-and parameters from the simulated disk, we estimate ???
-AU yr$^{-1}$ as an analytic prediction.
-While the interaction with a GI-active disk did not seem to result in a drastically different overall migration rate, the direction and magnitude of migration in the GI-active disk is highly variable and seems to depend strongly on interactions of the planet with discrete modes in the disk.
-In the case of the $t = 10$ planet run, the overall radial migration rate is similar but highly erratic.
+with parameters from the simulated disk, we estimate ???
+AU yr$^{-1}$ as an analytic expectation.
+This is probably a reasonable estimate for comparison purposes, because our resolution is probably insufficient to compute the corotation torque \citep{paardekooper2010}.
+%Scott and Aaron: Is this right?
+While the interaction with a GI-active disk does not seem to result in a drastically different overall migration rate, the direction and magnitude of migration in the GI-active disk is highly variable and seems to depend strongly on interactions of the planet with discrete modes in the disk.
+In the case of the $t = 10$ planet run, the overall radial migration rate is similar but even more erratic in direction.
 The important departures from laminar disk theory are that there are times when migration stalls and that interactions with discrete low-order modes seem to be important.
 The migration is neither a random walk nor monotonic but has a chaotic character due to nonlinear interactions.
-For all these reasons, it will be important in future studies to explore the dependence of the planet's behavior on its mass and on its placement relative to the site of GI-eruptions (for $t = 0$ type runs) and relative to the CR of low-order GI modes (for $t = 10$ type runs).
+For all these reasons, it will be important in future studies to explore the dependence of the planet's behavior on its mass and on its placement relative to the site of GI-eruptions (for $t = 0$ type runs) and relative to the CRs of low-order GI modes (for $t = 10$ type runs).
 
@@
 We plan to extend simulations like these to much longer times in the future.
-However, the final positions of the planets radius happens to be the location of the Inner Lindblad Resonance (ILR) for the strong m = 2 mode present in the fiducial run.
-In fact, this ILR is associated with a surface density enhancement in the disk.
-As the planet approaches this enhancement, it experiences a steeply falling surface density gradient which,
-in a laminar disk, could cause the migration to stall (???). It could also be that, near the ILR, the mode no longer effectively transfers angular momentum from the planet. 
+However, the final radial positions of both planets happens to be the location of the Inner Lindblad Resonance (ILR) for the strong m = 2 mode present in the fiducial run.
+This ILR is associated with a surface density enhancement in the disk.
+%Is this true?
+As the planet reaches the outer edge of a surface density enhancement, the local density gradient may be so steep that the ILR torques of the planet cancel the OLR torques of the planet.
+This would probably put the planet right outside the ILR of a major mode, because this is where we expect pile ups of material.
+(This came to me in a dream.) We can actually measure the surface density gradient near 17 AU and test this directly.
+As the planet approaches this enhancement, the surface density gradient may be falling steeply enough to halt migration, a mechanism can work in a laminar disk. 
 %The words that follow are from Scott.
 We must DO the analyses of periods, look at the surface density, look at the Q distribution (I doubt the planets are out of the GI active region yet), and think about the physics.
-Further work is necessary to adequately analyze the Fourier modes and their periodicities in the simulations with planets, but this gives a strong preliminary indication that the GI activity is driving the migration.
-Furthermore, where GIs are inactive planet migration is stalled, even Type I migration may be ineffective due to interference from the GIs at the planet’s OLR.
-%Is any of this true?
-An alternative but related possibility is that the planet reaches the outer edge of a surface density enhancement, so that the local density gradient is so steep the ILR's of the planet cancel the OLR's of the planet.
-This would probably put the planet right near the ILR of a major mode, because this is where we expect pile ups of material.
-(This came to me in a dream.)
+Alternatively, at 17 AU, the level of GI activity is quite different interior to the planet ($Q \approx$ ???) and exterior to the planet ($Q \approx$ ???), and this may affect the balance of the Lindblad torques exerted by the disk on the planet.
+Whether the migration really does stall and what the mechanism may be are subjects for further research.
 
-Magnitude of $e$ is roughly consistent with $c_s/v_K$ and low Mach number spirals.
-
-%Scott and Aaron: I am at a loss for a good punch line to end with. Help.
+%Scott and Aaron: I am at a loss for a good punch line to end with. Help. These are just some random remarks.
 The inclusion of a 1$M_J$ mass in the equilibrium disk initiates GIs and dramatically effects the character of the GIs.
@@
 GIs can have a strong effect on the orbital eccentricity of the included planet.
+Further work is necessary to adequately analyze these possible mechanisms, but these simulations give a preliminary indication that GI activity strongly affects planet migration.
 
@@
 \caption{Global amplitudes of nonaxisymmetric density perturbations $A_m$ as a function of time for individual cos($m\phi$) perturbations.
-The formula for computing $A_m$ is equation (15) of Boley et al.
-(2006).
+The formula for computing $A_m$ is equation (15) of \citep{boley2006}.
 Here the integrals are done only over the disk outside 15 AU to suppress contributions from edge effects in the inner disk.
-The set of red and black curves correspond to the $t = 0$ and $t = 10$ planet runs, respectively.
+The set of red and black curves correspond to the fiducial run and the $t = 0$ planet run, respectively.
 {\bf The radial migration plot is being used as a placeholder.}}

### 2010-09-16 Text/migration.tex
Dick's changes 9/16/10
git-svn-id: http://jaga.astro.indiana.edu/paperrepo/ApJL/Migration2010@73 3b7f7548-5c66-0410-83ab-3997167635fc

@@
 \begin{abstract}
-
+%Meaningless Change
 Characterization of migration in gravitationally unstable disks is necessary to understand the fate of protoplanets formed by disk instability.
@@
 The magnitude of $e$ is roughly consistent with the ratio of the sound speed to orbital speed of the disk and the modest values for the Mach numbers of the GI spirals \citep{boley2006a,boley2008}.
+In other words, the planet's orbit tends to have about as much nonaxisymmetry as reflected in the pitch angles of the GI-induced spiral \citep{cossins2009}.
 
@@
 
-%Scott and Aaron: I am at a loss for a good punch line to end with. Help. These are just some random remarks.
-The inclusion of a 1$M_J$ mass in the equilibrium disk initiates GIs and dramatically effects the character of the GIs.
-GIs can cause both inward and outward planet migration and can cause rapid migration in the burst phase.
-GIs can have a strong effect on the orbital eccentricity of the included planet.
-Further work is necessary to adequately analyze these possible mechanisms, but these simulations give a preliminary indication that GI activity strongly affects planet migration.
+{\it Summary.} The simulations in this Letter show that a planet of even only 1$M_J$ placed near the $Q$-minimum in a 0.14$M_{\odot}$ disk can hasten the onset of gravitational instabilities and change the nature of the dominant modes during the initial burst.
+Although the magnitude of radial migration is not too different in magnitude than computed from the simplest isothermal disk formula based on the Lindblad torques (a few$\times10^{-3}$ AU yr$^{-1}$), it can fluctuate in sign on time scales on the order of orbit periods and can be strongly affected by gravitational interactions of the planet with discrete GI-driven spiral modes, leading to phases of rapid radial motion.
+There is some suggestion in the simulations that planet migration may stall for long periods near the ILRs of dominant global GI-modes.
+%Scott and Aaron: I am at a loss for a good punch line to end with.
 
@@
 Here the integrals are done only over the disk outside 15 AU to suppress contributions from edge effects in the inner disk.
-The set of red and black curves correspond to the fiducial run and the $t = 0$ planet run, respectively.
+The set of red and black curves correspond to the fiducial run (no planet) and the $t = 0$ planet run, respectively.
 {\bf The radial migration plot is being used as a placeholder.}}

### 2011-05-16 Text/migration.tex
First go-over after a long dormant period, up to section 3.
	modified:   Text/migration.tex

git-svn-id: http://jaga.astro.indiana.edu/paperrepo/ApJL/Migration2010@75 3b7f7548-5c66-0410-83ab-3997167635fc

@@
 \usepackage{natbib}
+\usepackage{epstopdf}
+\usepackage{grffile}
 
@@
 \author{Aaron C. Boley}
-\affil{Department of Astronomy; University of Florida,
-211 Bryant Space Science Center,
-PO Box 112055, Gainesville, FL 32611}
+\affil{Department of Astronomy University of Florida, Gainesville, FL 32611}
 \email{aaron.boley@gmail.com}
@@
 At other times, when the planet interacts with a strong spiral mode, migration both inward and outward can be relatively rapid, covering several AUs over hundreds of years.
-{\bf See note in Latex file.}
+{\bf See note in Latex file.
+SM: confirmed from radial plots}
 %The assertion in the preceding sentence needs to be verified by analyzing both runs in more detail -- Am(t), periodograms, images zoomed in on the planet to check for arms.
@@
 %Scott: Fill in some references you know now. Aaron: Suggestions on references?
-Migration of gas giant planets due to interactions with a circumstellar gas disk probably plays a major role in defining the architecture of planetary systems.
+Migration of gas giant planets due to interactions with a circumstellar gas disk can play a major role in defining the architecture of planetary systems.
 Work on migration \citep[see review by][]{papaloizou2007} has included gravitational interaction of planets with both laminar and turbulent disks.
-However,  the self-gravity of the gas disk has usually been ignored, and the effects of radiative cooling and a detailed equation of state (EOS) have only recently been included \citep[e.g.,][]{paardekooper2006,paardekooper2010,paardekooper2010b}.
-\citet{boss2005} and \citet{mayer2004} examined radial migration of planet-mass fragments in gravitationally unstable disks, but their disks were violently disrupted by fragmentation under conditions (radii $<$ 40 AU, disk-to-star mass ratios $M_d/M_s \sim 0.1$, and stellar mass $M_s \sim 1\;M_{\odot}$) where this probably should not realistically have happened \citep{rafikov2005,rafikov2007,boley2006,boley2007b,boley2008,forgan2009,cai2010}.
+However,  the self-gravity of the gas disk has usually been ignored, and the effects of radiative cooling and a detailed equation of state (EOS) have only recently been included \citep[e.g.,][]{paardekooper2006,paardekooper2010,paardekooper2010b}{\bf NEED TO RECHECK PAARDEKOOPER FOR RADIATIVE COOLING}.
+\citet{boss2005} and \citet{mayer2004} examined radial migration of planet-mass fragments in gravitationally unstable disks, but their disks were violently disrupted by fragmentation under conditions (radii $<$ 40 AU, disk-to-star mass ratios $M_d/M_s \sim 0.1$, and stellar mass $M_s \sim 1\;M_{\odot}$) where, in reality, this probably would not have happened  \citep{rafikov2005,rafikov2007,boley2006,boley2007b,boley2008,forgan2009,cai2010}.
 More recently, fragmentation into clumps with gas giant or brown dwarf masses has been documented in numerical simulations of disks that are relatively massive ($M_d/M_s \sim$ a few tenths) and spatially extended (outer radii $>$ 50 AU) \citep{krumholz2007,stamatellos2007,stamatellos2009,boley2009,boley2010}, where fragmentation is expected from semi-analytic arguments \citep[e.g.,][]{clarke2009,rafikov2009,dodson2009}.
@@
 
-{\bf See note in Latex file.}
-%Scott and Aaron: Verify that everything I say here is correct.
 The CHYMERA (Computational HYdrodynamics with MultiplE Radiation Algorithms) code \citep{boley2007b} is a second-order, explicit, Eulerian scheme on a 3D cylindrical grid.
@@
 Calculations are done on a uniform cylindrical grid with reflection symmetry about the disk midplane and a grid size ($\varpi$,$\phi$,z) = (512,512,64).
-{\bf See note in Latex file.} 
-%Scott: The resolution stays the same.
-The grid size varies, but this is unimportant because initially the outer 256 radial zones are empty.
-It is simplest to just give the eventual grid size.
-So I did change "resolution" to grid size here.
 The $z$-axis is the rotation axis of the disk.
@@
 
-The model disk, based on an equlibrium model in \citet{pickett2003}, orbits a 1 $M_{\odot}$ star and has a mass $M_d = 0.14 M_{\odot}$, inner and outer radii at 5 and 40 AU, and an initial surface density distribution $\Sigma \sim \varpi^{-1}$.
-The time unit of one ORP (= outer rotation period) is defined as the rotation period of the initial disk at $\varpi \approx$ 32 AU, or about 250 yr.
+The model disk, based on an equlibrium model in \citet{pickett2003}, orbits a 1 $M_{\odot}$ star and has a mass $M_d = 0.14 M_{\odot}$, inner and outer radii at 5 and 40 AU, and an initial surface density distribution $\Sigma \sim \varpi^{-1/2}$.
+The time unit of one ORP (= outer rotation period) is defined as the rotation period of the initial disk at $\varpi \approx$ 32 AU, or about 180 yr.
 The disk is initially located between radial grid zones 30 and 240 and is close to isentropic, which results in a Toomre-$Q$ distribution with a marginally unstable \citep[see ][]{durisen2007} minimum value of ???
@@
 The computational grid extends radially to 512 zones to accommodate expansion of the outer disk once GIs become nonlinear.
-There is an inner cylindrical hole in the grid with a radius of 12 radial zones {\bf Scott Comment: brief description of hole needed.
-Dick's reply in Latex file}.
-%Scott: So add one!
+An outflow boundary condition is enforced at the upper vertical grid boundary, the outer radial grid boundary, and an inner radial boundary at 2 AU.
 To seed nonaxisymmetry, the density is given an initial 0.01 \% random cell-to-cell perturbation.
@@
 This Letter presents three simulations.
-The first, which we call the {\sl fiducial run}, goes from time $t = 0$ to ???
-20 ???
-ORP without a planet.
+The first, which we call the {\sl fiducial run}, goes from time $t = 0$ to 21 ORP without a planet.
 In the second simulation, which we call the {\sl t = 0 planet run}, a one Jupiter-mass ($M_J$) planet with a roughly circular midplane orbit is included in the initial $t = 0$ ORP equilibrium disk at 25 AU.
-In the third simulation, which we call the {\sl t = 10 planet run}, a 1 $M_J$ planet with a similar orbit is inserted into the fiducial run at $t = 10$ ORP.
+In the third simulation, which we call the {\sl t = 10 planet run}, a 1 $M_J$ planet with a similar orbit is inserted into the fiducial run at $t = 10.5$ ORP.
 These three simulations are part of a much larger suite, to be described elsewhere, in which the planet mass and starting locations are varied.
@@
 \center
-\includegraphics[width=12cm]{../Figures/planeta.pdf}
+\includegraphics[width=12cm]{Figures/planeta.eps}
 \caption{Global amplitudes of nonaxisymmetric density perturbations $A_m$ as a function of time for individual cos($m\phi$) perturbations.
@@
 \center
-\includegraphics[width=12cm]{../Figures/planeta.pdf}
+\includegraphics[width=12cm]{Figures/planeta.eps}
 \caption{Plot of the planet's radial position as a function of time $\varpi_p(t)$ for the two planet simulations.
@@
 \center
-\includegraphics[width=12cm]{../Figures/planete.pdf}
+\includegraphics[width=12cm]{Figures/planete.eps}
 \caption{Plot of the eccentricities $e$, computed for each 0 to 2$\phi$ change in azimuth of the planet as explained in the text, as a function of time.

### 2011-05-16 Text/migration.tex
Addressed many concerns and A_m(t) plot. Still waiting on a bit of periodogram analysis.
	new file:   Text/Figures/amt.eps
	modified:   Text/migration.tex

git-svn-id: http://jaga.astro.indiana.edu/paperrepo/ApJL/Migration2010@76 3b7f7548-5c66-0410-83ab-3997167635fc

@@
 
-{\bf See note in Latex file.}
-%Scott: Verify the accuracy of all this.
-
 The model disk, based on an equlibrium model in \citet{pickett2003}, orbits a 1 $M_{\odot}$ star and has a mass $M_d = 0.14 M_{\odot}$, inner and outer radii at 5 and 40 AU, and an initial surface density distribution $\Sigma \sim \varpi^{-1/2}$.
 The time unit of one ORP (= outer rotation period) is defined as the rotation period of the initial disk at $\varpi \approx$ 32 AU, or about 180 yr.
-The disk is initially located between radial grid zones 30 and 240 and is close to isentropic, which results in a Toomre-$Q$ distribution with a marginally unstable \citep[see ][]{durisen2007} minimum value of ???
-at radial grid zone ???
-(???
-AU).
+The disk is initially located between radial grid zones 30 and 240 and is close to isentropic, which results in a Toomre-$Q$ distribution with a marginally unstable \citep[see ][]{durisen2007} minimum value of 1.38 at radial grid zone 161 (26.7 AU).
 The computational grid extends radially to 512 zones to accommodate expansion of the outer disk once GIs become nonlinear.
@@
 Without a planet, when all modes grow from imposed noise, it takes 4 ORPs for coherent spiral modes to organize and grow to nonlinear amplitude.
-The growth is centered near the $Q$-minimum at ???
-AU.
+The growth is centered near the $Q$-minimum at $26$ 2AU.
 Modes of cos($m\phi$) symmetry with $m =$ 5 and 6 dominate.
@@
 The planet's radial migration apparently stalls at about 16 to 17 AU.
-From an analysis of periodicities present in the gas disk between 8 and 12 ORP, the planet lies a few AU inside the inner Lindblad resonance of a strong $m = 2$ mode with CR at ???
-28 ???
-AU in the GI-active disk in this early part of the asymptotic phase.
+From an analysis of periodicities present in the gas disk between 8 and 12 ORP, the planet lies a few AU inside the inner Lindblad resonance of a strong $m = 2$ mode with CR at 29 AU in the GI-active disk in this early part of the asymptotic phase.
 
@@
 We see intervals of fairly rapid inward or outward migration over several orbits, as well as times when radial migration appears to stall.
-Between $t = 15$ and 19 ORP, the pattern of outward migration for 2 ORP followed by an inward plunge over the next 2 ORP resembles the behavior in the $t = 0$ planet run during the burst, but, in this case, there is no distinct phase of GI activity.
-Animations of the simulation show there is a complex interaction between the planet, its spiral wake, and the global spiral arms caused by the instabilities.
-Analysis of images and periodicities suggests that, between 15 and 19 ORP, the planet may have passed through one arm of an $m =$ ???
-pattern with CR at ???
-AU.
-It is possible that the planet actually stimulated this particular mode, in manner similar to what occurred in the $t = 0$ planet run, because this $m =$ ???
+Between $t = 15$ and 19 ORP, the pattern of outward migration for 2 ORP followed by an inward plunge over the next 2 ORP resembles the behavior in the $t = 0$ planet run during the burst, but, in this case, there is no distinct transition between phases of GI activity.
+Animations of the simulation (available at {\bf BLAH}) show there is a complex interaction between the planet, its spiral wake, and the global spiral arms caused by the instabilities.
+Analysis of images and periodicities suggests that, between 15 and 19 ORP, the planet may be locked in a resonance with an arm of a $m = 2$ or 3 pattern both of which have CR at 27 AU.
+Over this time interval the average orbit period of the planet is 0.5 ORP, while both the $m = 2$ and 3 patterns have a pattern period of $\sim 0.74$ ORP, suggesting a 2:3 resonance.
+It is possible that the planet actually stimulated one or both of these modes, in manner similar to what occurred in the $t = 0$ planet run, because this $m =$ ???
 periodicity is absent over a similar time interval in the fiducial run.
@@
 
-Even in non-self-gravitating disks, the presence of discrete radial structure is known to cause instabilities (Li and Lovelace articles abound 2000, 2001).
-Meschiari \& Laughlin (2008) demonstrated that a gap, presumably opened by an embedded planet, can cause global modes to become unstable in self-gravitating disks that are otherwise stable against the development of GIs.
-Here we are dealing with a planet that is not massive enough to open a gap, but its gravitational interaction with the disk also has a strong effect on the growth and onset of GIs.
-{\bf Scott Comment: say more about gap opening criteria and whether they apply here.
-Dick's reply in tex file.}
-%Scott: Do it. I am not sure what you have in mind. Remember, however, that we are already at about maximum length for Letter.
+Even in non-self-gravitating disks, the presence of discrete radial structure is known to cause instabilities \citep{li1999,li2000}.
+\citet{meschiari2008} demonstrated that a gap, presumably opened by an embedded planet, can cause global modes to become unstable in self-gravitating disks that are otherwise stable against the development of GIs.
+Here we are dealing with a planet that is not quite massive enough to open a gap, but its gravitational interaction with the disk also has a strong effect on the growth and onset of GIs.
 
@@
 Figure \ref{fig:a} suggests that the overall radial drift during the first 10 ORP of the $t = 0$ planet run is not too different from what one would get by extrapolating the migration due solely to the planet's wake during the first 2 or 3 ORP, where the average radial migration is about 4$\times10^{-3}$ AU yr$^{-1}$.
-Although there are no simple formulae that are fully applicable to our radiatively cooled disk, using equation ???
-from ???
-with parameters from the simulated disk, we estimate ???
-AU yr$^{-1}$ as an analytic expectation.
+Although there are no simple formulae that are fully applicable to our radiatively cooled disk, using equation 70 from \citet{tanaka2002} with parameters from the simulated disk, we estimate 5.8$\times10^{-3}$ AU yr$^{-1}$ as an analytic expectation.
 This is probably a reasonable estimate for comparison purposes, because our resolution is probably insufficient to compute the corotation torque \citep{paardekooper2010}.
@@
 We plan to extend simulations like these to much longer times in the future.
-However, the final radial positions of both planets happens to be the location of the Inner Lindblad Resonance (ILR) for the strong m = 2 mode present in the fiducial run.
+However, the final radial positions of both planets happens to be the location of the Inner Lindblad Resonance (ILR) for the strong m = 2 mode present in both runs.
 This ILR is associated with a surface density enhancement in the disk.
@@
 We must DO the analyses of periods, look at the surface density, look at the Q distribution (I doubt the planets are out of the GI active region yet), and think about the physics.
-Alternatively, at 17 AU, the level of GI activity is quite different interior to the planet ($Q \approx$ ???) and exterior to the planet ($Q \approx$ ???), and this may affect the balance of the Lindblad torques exerted by the disk on the planet.
+Alternatively, at 17 AU, the level of GI activity is quite different interior to the planet ($Q \gtrsim 2$) and exterior to the planet ($Q \lesssim 2$), and this may affect the balance of the Lindblad torques exerted by the disk on the planet.
 Whether the migration really does stall and what the mechanism may be are subjects for further research.
@@
 \center
-\includegraphics[width=12cm]{Figures/planeta.eps}
+\includegraphics[width=12cm]{Figures/amt.eps}
 \caption{Global amplitudes of nonaxisymmetric density perturbations $A_m$ as a function of time for individual cos($m\phi$) perturbations.

### 2011-05-16 Text/migration.tex
Minor tweaks regarding the periodogram for the GI-active run.
	modified:   Text/migration.tex

git-svn-id: http://jaga.astro.indiana.edu/paperrepo/ApJL/Migration2010@77 3b7f7548-5c66-0410-83ab-3997167635fc

@@
 Over this time interval the average orbit period of the planet is 0.5 ORP, while both the $m = 2$ and 3 patterns have a pattern period of $\sim 0.74$ ORP, suggesting a 2:3 resonance.
-It is possible that the planet actually stimulated one or both of these modes, in manner similar to what occurred in the $t = 0$ planet run, because this $m =$ ???
-periodicity is absent over a similar time interval in the fiducial run.
+Both of these modes are present, at nearly the same periodicities, in the fiducial run so it is possible that the planet is migrating into a resonance orbit.
 

### 2011-05-23 Text/migration.tex
Dick's May 23, 2011 revisions
git-svn-id: http://jaga.astro.indiana.edu/paperrepo/ApJL/Migration2010@79 3b7f7548-5c66-0410-83ab-3997167635fc

@@
 \begin{abstract}
-%Meaningless Change
 Characterization of migration in gravitationally unstable disks is necessary to understand the fate of protoplanets formed by disk instability.
@@
 This Letter presents preliminary results from simulations with a Jupiter-mass planet placed in orbit at 25 AU within a 0.14 $M_{\odot}$ disk.
-The disk spans 5 to 40 AU around a 1 $M_{\odot}$ star and is initially marginally stable.
-In one simulation, the planet is inserted prior to the initial eruption of GIs; in another, it is inserted only after the disk has settled into a quasi-steady GI-active state, where heating by GIs roughly balances radiative cooling.
+The disk spans 5 to 40 AU around a 1 $M_{\odot}$ star and is initially close to marginally instability.
+In one simulation, the planet is inserted prior to the eruption of GIs; in another, it is inserted only after the disk has settled into a quasi-steady GI-active state, where heating by GIs roughly balances radiative cooling.
 When the planet is present from the beginning, its own wake stimulates growth of a particular global mode with which it strongly interacts, and the planet plunges inward six AU in about 10$^3$ years.
 In both cases with embedded planets, there are times when the planet's radial motion is slow and varies in direction.
-At other times, when the planet interacts with a strong spiral mode, migration both inward and outward can be relatively rapid, covering several AUs over hundreds of years.
-{\bf See note in Latex file.
-SM: confirmed from radial plots}
-%The assertion in the preceding sentence needs to be verified by analyzing both runs in more detail -- Am(t), periodograms, images zoomed in on the planet to check for arms.
-The migration is observed; it is the relationship to mode interactions that needs to be checked.
-Planet orbit eccentricity tends to fluctuate rapidly and ranges between about 0.02 to 0.1.
+At other times, when the planet appears to be interacting with strong spiral modes, migration both inward and outward can be relatively rapid, covering several AUs over hundreds of years.
+Migration in both cases appears to stall near the inner Lindblad resonance of a dominant low-order mode.
+Planet orbit eccentricities fluctuate rapidly between about 0.02 to 0.1 throughout the GI-active phases of the simulations.
 \end{abstract}
@@
 
-{\bf See note in Latex file.} 
-%Scott: Fill in some references you know now. Aaron: Suggestions on references?
 Migration of gas giant planets due to interactions with a circumstellar gas disk can play a major role in defining the architecture of planetary systems.
 Work on migration \citep[see review by][]{papaloizou2007} has included gravitational interaction of planets with both laminar and turbulent disks.
-However,  the self-gravity of the gas disk has usually been ignored, and the effects of radiative cooling and a detailed equation of state (EOS) have only recently been included \citep[e.g.,][]{paardekooper2006,paardekooper2010,paardekooper2010b}{\bf NEED TO RECHECK PAARDEKOOPER FOR RADIATIVE COOLING}.
-\citet{boss2005} and \citet{mayer2004} examined radial migration of planet-mass fragments in gravitationally unstable disks, but their disks were violently disrupted by fragmentation under conditions (radii $<$ 40 AU, disk-to-star mass ratios $M_d/M_s \sim 0.1$, and stellar mass $M_s \sim 1\;M_{\odot}$) where, in reality, this probably would not have happened  \citep{rafikov2005,rafikov2007,boley2006,boley2007b,boley2008,forgan2009,cai2010}.
+However,  the self-gravity of the gas disk has usually been ignored, and the effects of radiative cooling and a detailed equation of state (EOS) have only recently been included \citep[e.g.,][]{paardekooper2006,paardekooper2010,paardekooper2010b}.
+%***DICK: Scott, you need to do this: {\bf NEED TO RECHECK PAARDEKOOPER FOR RADIATIVE COOLING} 
+\citet{boss2005} and \citet{mayer2004} examined radial migration of planet-mass fragments in gravitationally unstable disks, but their disks were violently disrupted by fragmentation under conditions (radii $<$ 40 AU, disk-to-star mass ratios $M_d/M_s \sim 0.1$, and stellar mass $M_s \sim 1\;M_{\odot}$) where fragmentation may not actually occur  \citep{rafikov2005,rafikov2007,boley2006,boley2007b,boley2008,forgan2009,cai2010}.
 More recently, fragmentation into clumps with gas giant or brown dwarf masses has been documented in numerical simulations of disks that are relatively massive ($M_d/M_s \sim$ a few tenths) and spatially extended (outer radii $>$ 50 AU) \citep{krumholz2007,stamatellos2007,stamatellos2009,boley2009,boley2010}, where fragmentation is expected from semi-analytic arguments \citep[e.g.,][]{clarke2009,rafikov2009,dodson2009}.
-The fate of the clumps depends in part on their radial migration, which is a chaotic and messy affair in a fragmenting disk \citep[e.g.,][]{boley2009,boley2010,vorobyov2010,boleydurisen2010}.
+The fate of the clumps then depends in part on their radial migration, which is a chaotic and messy affair in a fragmenting disk \citep[e.g.,][]{boley2009,boley2010,vorobyov2010,boleydurisen2010}.
 The occurrence of gravitational instabilities (GIs) may also be episodic \citep[e.g.,][]{vorobyov2006,vorobyov2010b,zhu2010}.
@@
 The planet integration is done with a Verlet integrator \citep[e.g.,][]{hut1995}, and the indirect potential terms are treated as in \citet{nelson2000a}.
-The Rosseland mean and Planck mean opacities and molecular weights in our simulations are the same as those in \citet{boley2006,boley2007b}, except that we correct the \citet{dalessio2001} gas mean molecular weight to $\mu$ = 2.39.
+The Rosseland mean and Planck mean opacities and molecular weights in our simulations are the same as those in \citet{boley2006,boley2007b}, except that we correct the \citet{dalessio2001} gas mean molecular weight to $\mu$ = 2.33.
 
@@
 In the third simulation, which we call the {\sl t = 10 planet run}, a 1 $M_J$ planet with a similar orbit is inserted into the fiducial run at $t = 10.5$ ORP.
-These three simulations are part of a much larger suite, to be described elsewhere, in which the planet mass and starting locations are varied.
-We estimate crude semi-major axes $a$ and eccentricities $e$ from the planet's $\varpi(t)$ by using at the maximum and minimum radii, $\varpi_{\rm max}$ and $\varpi_{\rm min}$, attained by the planet over each interval that the planet's $\phi$-coordinate changes by 2$\pi$ (one ``orbit'').
+These three simulations are part of a much larger suite, to be described elsewhere, in which the planet mass is varied.
+We estimate crude semi-major axes $a$ and eccentricities $e$ from the planet's $\varpi(t)$ by using the maximum and minimum radii, $\varpi_{\rm max}$ and $\varpi_{\rm min}$, attained by the planet over each interval that the planet's $\phi$-coordinate changes by 2$\pi$ (one ``orbit'').
 Then, for that orbit, $a = (\varpi_{\rm max}+\varpi_{\rm min})/2$ and $e = (\varpi_{\rm max}-\varpi_{\rm min})/(\varpi_{\rm max}+\varpi_{\rm min})$.
@@
 
-{\it The Fiducial Run.} As in \citet{mejia2005}, the fiducial run exhibits four distinct phases: initial {\sl axisymmetric cooling}, the onset of GIs in a well-defined {\sl burst} in low-order modes, a {\sl transition} to more high-order nonaxisymmetric structure, and the {\sl asymptotic} establishment of quasi-steady chaotic GI-activity with an overall balance of heating and cooling.
+{\it The Fiducial Run.} As in \citet{mejia2005}, the fiducial run exhibits four distinct phases: initial {\sl axisymmetric cooling}, the onset of GIs in a well-defined {\sl burst} of a few discrete low-order modes, a {\sl transition} to more complex nonaxisymmetric structure, and the {\sl asymptotic} establishment of quasi-steady GI activity with an overall balance of heating and cooling.
 There is no fragmentation because this disk has a long cooling time \citep{gammie2001,boley2006,boley2007b}.
 
-%Scott: A number of important assertions in this paragraph need to verified by further analysis.
-{\it The $t = 0$ Planet Run.} {\bf See note in Latex file.} Figure \ref{fig:Am} shows that inclusion of even this modest mass planet ($0.8 \%$ of $M_d$) has a dramatic effect on the burst phase.
+{\it The $t = 0$ Planet Run.} Figure \ref{fig:Am} shows that inclusion of even this modest mass planet ($0.7 \%$ of $M_d$) has a dramatic effect on the burst phase.
 Without a planet, when all modes grow from imposed noise, it takes 4 ORPs for coherent spiral modes to organize and grow to nonlinear amplitude.
-The growth is centered near the $Q$-minimum at $26$ 2AU.
-Modes of cos($m\phi$) symmetry with $m =$ 5 and 6 dominate.
+The growth is centered near the $Q$-minimum at $26$ AU.
+Modes of cos($m\phi$) symmetry with $m =$ 4 and 5 dominate.
 An $m = 3$ (three-armed) spiral also grows, but somewhat more slowly, and lags the other modes in time.
 In the $t = 0$ planet run, the planet develops an organized wake within about an ORP in which $m = 3$ is a dominant component ($\sim 5$ \% global density perturbation).
-This nonlinear seeding causes $m = 3$ to dominate the GI burst, which also occurs $\sim$ 2 ORPs earlier.
+This nonlinear seeding causes $m = 3$ to dominate the GI burst, which also occurs $\sim$ 2 ORP earlier.
 Because of our initial placement of the planet inside but close to the $Q$-minimum, the corotation radius (CR) of the $m = 3$ mode is fairly close to that of the planet's orbit period.
@@
 From 0 to 2.5 ORP, the planet is torqued primarily by its own wake and migrates inward.
-Beginning at 3 ORP, when the $m = 3$ mode has grown significantly and the planet is lagging one of its arms in azimuth, the planet gains angular momentum and moves outward.
+%DICK: New words.
+Beginning at about 3 ORP, the planet interacts with the now nonlinear $m$ = 3 GI mode.
+At first, the planet gains angular momentum and moves outward, but, from 4 to 8 ORP, a time interval of  about 720 yr, the plant experiences a negative torque and plunges from 23 to 17 AU.
+%***DICK: As we agreed, the wording below could not be verified. I leave it here in case you want to do more analysis to check it out. I have replaced these words with something more vague.
+%Beginning at 3 ORP, when the $m = 3$ mode has grown significantly and the planet is lagging one of its arms in azimuth, the planet gains angular momentum and moves outward.
 At about $t =$ 4 ORP, as the planet overtakes the arm, the sign of the net torque reverses and the planet plunges from 23 to 17 AU in about 4 ORP (10$^3$ yr).
 During this time, the planet and mode seem to interact strongly enough that the planet essentially surfs along the arm, maintaining a location within the arm as it plunges.
-After $t =$ 8 ORP, the main burst is over, and the disk transitions into its asymptotic state, where modes of many $m$-values are comparably strong.
+After $t =$ 8 ORP, the main burst is over, and the disk transitions into its asymptotic state, where modes of many $m$-values become comparably strong (see Fig.
+\ref{fig:Am}).
 The planet's radial migration apparently stalls at about 16 to 17 AU.
-From an analysis of periodicities present in the gas disk between 8 and 12 ORP, the planet lies a few AU inside the inner Lindblad resonance of a strong $m = 2$ mode with CR at 29 AU in the GI-active disk in this early part of the asymptotic phase.
+From an analysis of periodicities present in the gas disk between 8 and 12 ORP, the planet lies a few AU inside the inner Lindblad resonance of a strong $m = 2$ mode with CR at 29 AU in this early part of the asymptotic phase.
 
-%Scott: Even more work is needed to verify (or refute) what is said here. Some of what I wrote is pure speculation and needs to be checked.
-{\it The $t = 10$ Planet Run.} {\bf See note in Latex file.} The motion of the planet in this case is more difficult to interpret.
+{\it The $t = 10$ Planet Run.} The motion of the planet in this case is more difficult to interpret.
 We see intervals of fairly rapid inward or outward migration over several orbits, as well as times when radial migration appears to stall.
 Between $t = 15$ and 19 ORP, the pattern of outward migration for 2 ORP followed by an inward plunge over the next 2 ORP resembles the behavior in the $t = 0$ planet run during the burst, but, in this case, there is no distinct transition between phases of GI activity.
-Animations of the simulation (available at {\bf BLAH}) show there is a complex interaction between the planet, its spiral wake, and the global spiral arms caused by the instabilities.
-Analysis of images and periodicities suggests that, between 15 and 19 ORP, the planet may be locked in a resonance with an arm of a $m = 2$ or 3 pattern both of which have CR at 27 AU.
-Over this time interval the average orbit period of the planet is 0.5 ORP, while both the $m = 2$ and 3 patterns have a pattern period of $\sim 0.74$ ORP, suggesting a 2:3 resonance.
-Both of these modes are present, at nearly the same periodicities, in the fiducial run so it is possible that the planet is migrating into a resonance orbit.
+Animations of the simulation (available at ???URL???)
+%***DICK: Fill in ???URL???
+show there is a complex interaction between the planet, its spiral wake, and the global spiral arms caused by the instabilities.
+Analysis of images and periodicities suggests that, between 15 and 19 ORP, the planet may be in a mean motion resonance with $m = 2$ and 3 patterns, both of which have CR at 27 AU.
+Over this time interval, the average orbit period of the planet is $\sim$ 0.50 ORP, while both the $m = 2$ and 3 spirals have a pattern period of $\sim 0.74$ ORP, suggesting a 3:2 resonance.
+Both of these modes are present, at nearly the same periodicities, in the fiducial run.
+So, in contrast with the $t = 0$ planet run, this planet's rapid phases of radial migration are caused by interaction with strong GI modes that exist independent of the planet.
 
@@
 The presence of GI activity in the $t = 10$ planet run immediately increases $e$ to $\sim 0.08$.
-In the $t = 0$ planet run, the modest initial $e$ decreases during the 1.5 ORP when it is migrating only due to its own wake.
+In the $t = 0$ planet run, the modest initial $e$ decreases during the 1.5 ORP when it is migrating only due to its own wake, as expected.
+%***DICK: Scott, we need a reference for this "expectation".
 However, once there are strong nonlinear interactions with GI modes, $e$ jumps upward.
@@
 The magnitude of $e$ is roughly consistent with the ratio of the sound speed to orbital speed of the disk and the modest values for the Mach numbers of the GI spirals \citep{boley2006a,boley2008}.
-In other words, the planet's orbit tends to have about as much nonaxisymmetry as reflected in the pitch angles of the GI-induced spiral \citep{cossins2009}.
+In other words, the planet's orbit tends to have about as much nonaxisymmetry as reflected in the pitch angles of the GI-induced spirals \citep{cossins2009}.
 
@@
 
-{\bf See notes in Latex file.}
-%Scott and Aaron: This section needs a lot of help.
-
 Even in non-self-gravitating disks, the presence of discrete radial structure is known to cause instabilities \citep{li1999,li2000}.
@@
 
-%Scott and Aaron: Can you go over this part carefully and fill in some calculation for the migration rate?
 Figure \ref{fig:a} suggests that the overall radial drift during the first 10 ORP of the $t = 0$ planet run is not too different from what one would get by extrapolating the migration due solely to the planet's wake during the first 2 or 3 ORP, where the average radial migration is about 4$\times10^{-3}$ AU yr$^{-1}$.
-Although there are no simple formulae that are fully applicable to our radiatively cooled disk, using equation 70 from \citet{tanaka2002} with parameters from the simulated disk, we estimate 5.8$\times10^{-3}$ AU yr$^{-1}$ as an analytic expectation.
+Although there are no simple formulae that are fully applicable to our radiatively cooled disk, using eq.
+(70) from \citet{tanaka2002} with parameters from the simulated disk, we estimate 6$\times10^{-3}$ AU yr$^{-1}$ as an analytic expectation.
 This is probably a reasonable estimate for comparison purposes, because our resolution is probably insufficient to compute the corotation torque \citep{paardekooper2010}.
-%Scott and Aaron: Is this right?
-While the interaction with a GI-active disk does not seem to result in a drastically different overall migration rate, the direction and magnitude of migration in the GI-active disk is highly variable and seems to depend strongly on interactions of the planet with discrete modes in the disk.
-In the case of the $t = 10$ planet run, the overall radial migration rate is similar but even more erratic in direction.
-The important departures from laminar disk theory are that there are times when migration stalls and that interactions with discrete low-order modes seem to be important.
+While the interaction with a GI-active disk does not seem to result in a drastically different overall migration rate, the direction and magnitude of migration in both simulations is quite variable and seems to depend strongly on interactions of the planet with discrete modes in the disk.
+In $t = 10$ planet run, the overall radial migration rate is similar but is even more erratic in direction than in the $t = 0$ planet run.
+The important departures from laminar disk theory are the more erratic nature of migration, the importance of interactions with discrete low-order modes, and stalling of migration near the inner edge of the strongest GI activity.
 The migration is neither a random walk nor monotonic but has a chaotic character due to nonlinear interactions.
@@
 
-%Scott and Aaron: This is REALLY speculative. We need analysis here! The following is a modified version of text from Scott's poster mixed with the idea from my dream.
 In both simulations, the inward migration appears to halt near 17 AU.
@@
 We plan to extend simulations like these to much longer times in the future.
-However, the final radial positions of both planets happens to be the location of the Inner Lindblad Resonance (ILR) for the strong m = 2 mode present in both runs.
+However, the final radial positions of both planets happen to be the location of the Inner Lindblad resonance (ILR) for a strong m = 2 mode present in both runs.
 This ILR is associated with a surface density enhancement in the disk.
-%Is this true?
-As the planet reaches the outer edge of a surface density enhancement, the local density gradient may be so steep that the ILR torques of the planet cancel the OLR torques of the planet.
-This would probably put the planet right outside the ILR of a major mode, because this is where we expect pile ups of material.
-(This came to me in a dream.) We can actually measure the surface density gradient near 17 AU and test this directly.
-As the planet approaches this enhancement, the surface density gradient may be falling steeply enough to halt migration, a mechanism can work in a laminar disk. 
-%The words that follow are from Scott.
-We must DO the analyses of periods, look at the surface density, look at the Q distribution (I doubt the planets are out of the GI active region yet), and think about the physics.
+Near this enhancement, the surface density gradient may be falling steeply enough to halt migration, a mechanism that can work in a laminar disk \citep[note the $\alpha$ dependence in eq.
+(70) of][]{tanaka2002}.
+%***DICK: Is my reference here OK?.
 Alternatively, at 17 AU, the level of GI activity is quite different interior to the planet ($Q \gtrsim 2$) and exterior to the planet ($Q \lesssim 2$), and this may affect the balance of the Lindblad torques exerted by the disk on the planet.
@@
 {\it Summary.} The simulations in this Letter show that a planet of even only 1$M_J$ placed near the $Q$-minimum in a 0.14$M_{\odot}$ disk can hasten the onset of gravitational instabilities and change the nature of the dominant modes during the initial burst.
-Although the magnitude of radial migration is not too different in magnitude than computed from the simplest isothermal disk formula based on the Lindblad torques (a few$\times10^{-3}$ AU yr$^{-1}$), it can fluctuate in sign on time scales on the order of orbit periods and can be strongly affected by gravitational interactions of the planet with discrete GI-driven spiral modes, leading to phases of rapid radial motion.
-There is some suggestion in the simulations that planet migration may stall for long periods near the ILRs of dominant global GI-modes.
-%Scott and Aaron: I am at a loss for a good punch line to end with.
+Although the magnitude of radial migration (a few$\times10^{-3}$ AU yr$^{-1}$) is not too different from that computed using the \citet{tanaka2002}  isothermal disk formula based on the Lindblad torques, it can fluctuate in sign on time scales on the order of the orbit period and can be strongly affected by gravitational interactions of the planet with discrete GI-driven spiral modes, leading to phases of rapid radial motion.
+Planet orbit eccentricities fluctuate between about 0.02 and 0.10.
+The simulations also suggest that planet migration may stall for long periods near the ILRs of dominant global GI-modes.
 
@@
 The research was conducted with the support of NASA grants from the Origins of Solar Systems  Program (NNG05GN11G and NNX08AK36G).
+S.M.
+was supported during portions of this research by an NESSP Graduate Fellowship and by Indiana Space Grant Consortium Graduate Fellowships; A.C.B.
+is currently a Sagan Postdoctoral Fellow at U.
+Florida.
 This material is based upon work supported by the National Science Foundation under Grants No.
@@
 Any opinions, findings and conclusions or recommendations expressed in this material are those of the authors and do not necessarily reflect the views of the National Science Foundation.
-{\bf need TG acks}
+%***DICK: Any other acknowledgments needed? Scott, you had this in here: {\bf need TG acks}
 
@@
 
+%***DICK: Scott, remember that figures need some improvements.
+
 \begin{figure}[t]
 \center
-\includegraphics[width=12cm]{Figures/amt.eps}
+\includegraphics[width=11cm]{Figures/amt.eps}
 \caption{Global amplitudes of nonaxisymmetric density perturbations $A_m$ as a function of time for individual cos($m\phi$) perturbations.
@@
 Here the integrals are done only over the disk outside 15 AU to suppress contributions from edge effects in the inner disk.
-The set of red and black curves correspond to the fiducial run (no planet) and the $t = 0$ planet run, respectively.
-{\bf The radial migration plot is being used as a placeholder.}}
+The set of red and black curves correspond to the fiducial run (no planet) and the $t = 0$ planet run, respectively.}
 \label{fig:Am}

### 2011-05-26 Text/migration.tex
Merged in Dick's changes and fixed the last of the issues. Still needs figure tweaking.
	modified:   Text/migration.tex

git-svn-id: http://jaga.astro.indiana.edu/paperrepo/ApJL/Migration2010@80 3b7f7548-5c66-0410-83ab-3997167635fc

@@
 Work on migration \citep[see review by][]{papaloizou2007} has included gravitational interaction of planets with both laminar and turbulent disks.
-However,  the self-gravity of the gas disk has usually been ignored, and the effects of radiative cooling and a detailed equation of state (EOS) have only recently been included \citep[e.g.,][]{paardekooper2006,paardekooper2010,paardekooper2010b}.
-%***DICK: Scott, you need to do this: {\bf NEED TO RECHECK PAARDEKOOPER FOR RADIATIVE COOLING} 
+However, radiative transport, detailed equations of state (EOS), and the self-gravity of the gas disk have usually been ignored, the effects of a non-isothermal EOS have only recently been included \citep[e.g.,][]{paardekooper2006,paardekooper2010,paardekooper2011}.
 \citet{boss2005} and \citet{mayer2004} examined radial migration of planet-mass fragments in gravitationally unstable disks, but their disks were violently disrupted by fragmentation under conditions (radii $<$ 40 AU, disk-to-star mass ratios $M_d/M_s \sim 0.1$, and stellar mass $M_s \sim 1\;M_{\odot}$) where fragmentation may not actually occur  \citep{rafikov2005,rafikov2007,boley2006,boley2007b,boley2008,forgan2009,cai2010}.
 More recently, fragmentation into clumps with gas giant or brown dwarf masses has been documented in numerical simulations of disks that are relatively massive ($M_d/M_s \sim$ a few tenths) and spatially extended (outer radii $>$ 50 AU) \citep{krumholz2007,stamatellos2007,stamatellos2009,boley2009,boley2010}, where fragmentation is expected from semi-analytic arguments \citep[e.g.,][]{clarke2009,rafikov2009,dodson2009}.
-The fate of the clumps then depends in part on their radial migration, which is a chaotic and messy affair in a fragmenting disk \citep[e.g.,][]{boley2009,boley2010,vorobyov2010,boleydurisen2010}.
+The fate of the clumps then depends in part on their radial migration, which is a chaotic and messy affair in a fragmenting disk \citep[e.g.,][]{boley2009,boley2010,vorobyov2010,boley2010b}.
 The occurrence of gravitational instabilities (GIs) may also be episodic \citep[e.g.,][]{vorobyov2006,vorobyov2010b,zhu2010}.
@@
 From 0 to 2.5 ORP, the planet is torqued primarily by its own wake and migrates inward.
-%DICK: New words.
-Beginning at about 3 ORP, the planet interacts with the now nonlinear $m$ = 3 GI mode.
+Beginning at about 3 ORP, the planet interacts with the now nonlinear $m = 3$ GI mode.
 At first, the planet gains angular momentum and moves outward, but, from 4 to 8 ORP, a time interval of  about 720 yr, the plant experiences a negative torque and plunges from 23 to 17 AU.
-%***DICK: As we agreed, the wording below could not be verified. I leave it here in case you want to do more analysis to check it out. I have replaced these words with something more vague.
-%Beginning at 3 ORP, when the $m = 3$ mode has grown significantly and the planet is lagging one of its arms in azimuth, the planet gains angular momentum and moves outward.
-At about $t =$ 4 ORP, as the planet overtakes the arm, the sign of the net torque reverses and the planet plunges from 23 to 17 AU in about 4 ORP (10$^3$ yr).
-During this time, the planet and mode seem to interact strongly enough that the planet essentially surfs along the arm, maintaining a location within the arm as it plunges.
+
 After $t =$ 8 ORP, the main burst is over, and the disk transitions into its asymptotic state, where modes of many $m$-values become comparably strong (see Fig.
@@
 The presence of GI activity in the $t = 10$ planet run immediately increases $e$ to $\sim 0.08$.
-In the $t = 0$ planet run, the modest initial $e$ decreases during the 1.5 ORP when it is migrating only due to its own wake, as expected.
-%***DICK: Scott, we need a reference for this "expectation".
+In the $t = 0$ planet run, the modest initial $e$ decreases during the 1.5 ORP when it is migrating only due to its own wake, as expected \citep{ward1998,ward2003,goldreich2004}.
 However, once there are strong nonlinear interactions with GI modes, $e$ jumps upward.
@@
 This ILR is associated with a surface density enhancement in the disk.
-Near this enhancement, the surface density gradient may be falling steeply enough to halt migration, a mechanism that can work in a laminar disk \citep[note the $\alpha$ dependence in eq.
-(70) of][]{tanaka2002}.
-%***DICK: Is my reference here OK?.
+Near this enhancement, the surface density gradient may be falling steeply enough to halt migration, a mechanism that can work in a laminar disk \citep{paardekooper2009}.
 Alternatively, at 17 AU, the level of GI activity is quite different interior to the planet ($Q \gtrsim 2$) and exterior to the planet ($Q \lesssim 2$), and this may affect the balance of the Lindblad torques exerted by the disk on the planet.
@@
 This material is based upon work supported by the National Science Foundation under Grants No.
-ACI-0338618l, OCI-0451237, OCI-0535258, OCI-0504075, and CNS-0521433.
+ACI-0338618l, OCI-0451237, OCI-0535258, OCI-0504075, and CNS-0521433, and through TeraGrid resources provided by Indiana University, PSC, SDSC, and NCSA.
+TeraGrid systems are hosted by Indiana University, LONI, NCAR, NCSA, NICS, ORNL, PSC, Purdue University, SDSC, TACC, and UC/ANL..
 Any opinions, findings and conclusions or recommendations expressed in this material are those of the authors and do not necessarily reflect the views of the National Science Foundation.
-%***DICK: Any other acknowledgments needed? Scott, you had this in here: {\bf need TG acks}
 

### 2011-05-26 Text/migration.tex
Added URL
	modified:   Text/migration.tex

git-svn-id: http://jaga.astro.indiana.edu/paperrepo/ApJL/Migration2010@83 3b7f7548-5c66-0410-83ab-3997167635fc

@@
 This Letter presents preliminary results from simulations with a Jupiter-mass planet placed in orbit at 25 AU within a 0.14 $M_{\odot}$ disk.
-The disk spans 5 to 40 AU around a 1 $M_{\odot}$ star and is initially close to marginally instability.
+The disk spans 5 to 40 AU around a 1 $M_{\odot}$ star and is initially close to marginal stability.
 In one simulation, the planet is inserted prior to the eruption of GIs; in another, it is inserted only after the disk has settled into a quasi-steady GI-active state, where heating by GIs roughly balances radiative cooling.
@@
 The growth is centered near the $Q$-minimum at $26$ AU.
-Modes of cos($m\phi$) symmetry with $m =$ 4 and 5 dominate.
+Modes of cos($m\phi$) symmetry with $m = 4$ and 5 dominate.
 An $m = 3$ (three-armed) spiral also grows, but somewhat more slowly, and lags the other modes in time.
@@
 When the triggered $m = 3$ mode becomes strongly nonlinear at about 3 ORP, planet migration is significantly affected.
-Figure \ref{fig:a} shows the evolution of the planet's radial position.
-From 0 to 2.5 ORP, the planet is torqued primarily by its own wake and migrates inward.
+
+Figure \ref{fig:a} shows the evolution of the planet's radial position. From 0 to 2.5 ORP, the planet is torqued primarily by its own wake and migrates inward. 
 Beginning at about 3 ORP, the planet interacts with the now nonlinear $m = 3$ GI mode.
 At first, the planet gains angular momentum and moves outward, but, from 4 to 8 ORP, a time interval of  about 720 yr, the plant experiences a negative torque and plunges from 23 to 17 AU.
-
-After $t =$ 8 ORP, the main burst is over, and the disk transitions into its asymptotic state, where modes of many $m$-values become comparably strong (see Fig.
+After $t = 8$ ORP, the main burst is over, and the disk transitions into its asymptotic state, where modes of many $m$-values become comparably strong (see Fig.
 \ref{fig:Am}).
@@
 Between $t = 15$ and 19 ORP, the pattern of outward migration for 2 ORP followed by an inward plunge over the next 2 ORP resembles the behavior in the $t = 0$ planet run during the burst, but, in this case, there is no distinct transition between phases of GI activity.
-Animations of the simulation (available at ???URL???)
-%***DICK: Fill in ???URL???
-show there is a complex interaction between the planet, its spiral wake, and the global spiral arms caused by the instabilities.
+Animations of the simulation (available at http://hdl.handle.net/2022/13304) show there is a complex interaction between the planet, its spiral wake, and the global spiral arms caused by the instabilities.
 Analysis of images and periodicities suggests that, between 15 and 19 ORP, the planet may be in a mean motion resonance with $m = 2$ and 3 patterns, both of which have CR at 27 AU.

### 2011-05-30 Text/migration.tex
Dick's Final Changes May 30, 2011
git-svn-id: http://jaga.astro.indiana.edu/paperrepo/ApJL/Migration2010@85 3b7f7548-5c66-0410-83ab-3997167635fc

@@
 This Letter presents preliminary results from simulations with a Jupiter-mass planet placed in orbit at 25 AU within a 0.14 $M_{\odot}$ disk.
-The disk spans 5 to 40 AU around a 1 $M_{\odot}$ star and is initially close to marginal stability.
+The disk spans 5 to 40 AU around a 1 $M_{\odot}$ star and is initially marginally unstable.
 In one simulation, the planet is inserted prior to the eruption of GIs; in another, it is inserted only after the disk has settled into a quasi-steady GI-active state, where heating by GIs roughly balances radiative cooling.
@@
 Work on migration \citep[see review by][]{papaloizou2007} has included gravitational interaction of planets with both laminar and turbulent disks.
-However, radiative transport, detailed equations of state (EOS), and the self-gravity of the gas disk have usually been ignored, the effects of a non-isothermal EOS have only recently been included \citep[e.g.,][]{paardekooper2006,paardekooper2010,paardekooper2011}.
+However, radiative transport, detailed equations of state (EOS), and the self-gravity of the gas disk have usually been ignored; the effects of a non-isothermal EOS have only recently been included \citep[e.g.,][]{paardekooper2006,paardekooper2010,paardekooper2011}.
 \citet{boss2005} and \citet{mayer2004} examined radial migration of planet-mass fragments in gravitationally unstable disks, but their disks were violently disrupted by fragmentation under conditions (radii $<$ 40 AU, disk-to-star mass ratios $M_d/M_s \sim 0.1$, and stellar mass $M_s \sim 1\;M_{\odot}$) where fragmentation may not actually occur  \citep{rafikov2005,rafikov2007,boley2006,boley2007b,boley2008,forgan2009,cai2010}.
@@
 The fate of the clumps then depends in part on their radial migration, which is a chaotic and messy affair in a fragmenting disk \citep[e.g.,][]{boley2009,boley2010,vorobyov2010,boley2010b}.
-The occurrence of gravitational instabilities (GIs) may also be episodic \citep[e.g.,][]{vorobyov2006,vorobyov2010b,zhu2010}.
+The occurrence of gravitational instabilities (GIs) may be episodic \citep[e.g.,][]{vorobyov2006,vorobyov2010b,zhu2010}.
 Clumps that survive and contract to the dimensions of young planets can later find themselves in a disk that erupts again into GI activity.
@@
 In the third simulation, which we call the {\sl t = 10 planet run}, a 1 $M_J$ planet with a similar orbit is inserted into the fiducial run at $t = 10.5$ ORP.
-These three simulations are part of a much larger suite, to be described elsewhere, in which the planet mass is varied.
+These three simulations are part of a larger suite, to be described elsewhere, in which the planet mass is varied.
 We estimate crude semi-major axes $a$ and eccentricities $e$ from the planet's $\varpi(t)$ by using the maximum and minimum radii, $\varpi_{\rm max}$ and $\varpi_{\rm min}$, attained by the planet over each interval that the planet's $\phi$-coordinate changes by 2$\pi$ (one ``orbit'').
@@
 This nonlinear seeding causes $m = 3$ to dominate the GI burst, which also occurs $\sim$ 2 ORP earlier.
-Because of our initial placement of the planet inside but close to the $Q$-minimum, the corotation radius (CR) of the $m = 3$ mode is fairly close to that of the planet's orbit period.
+Because of our initial placement of the planet inside but close to the $Q$-minimum, the corotation radius (CR) of the $m = 3$ mode is fairly close to that of the planet's orbit radius.
 When the triggered $m = 3$ mode becomes strongly nonlinear at about 3 ORP, planet migration is significantly affected.
@@
 
-Even in non-self-gravitating disks, the presence of discrete radial structure is known to cause instabilities \citep{li1999,li2000}.
+Even in non-self-gravitating disks, the presence of discrete radial structure is known to cause instabilities \citep{li2000,li2001}.
 \citet{meschiari2008} demonstrated that a gap, presumably opened by an embedded planet, can cause global modes to become unstable in self-gravitating disks that are otherwise stable against the development of GIs.
@@
 (70) from \citet{tanaka2002} with parameters from the simulated disk, we estimate 6$\times10^{-3}$ AU yr$^{-1}$ as an analytic expectation.
-This is probably a reasonable estimate for comparison purposes, because our resolution is probably insufficient to compute the corotation torque \citep{paardekooper2010}.
+This is a reasonable estimate for comparison purposes, because our resolution is probably insufficient to compute the corotation torque \citep{paardekooper2010}.
 While the interaction with a GI-active disk does not seem to result in a drastically different overall migration rate, the direction and magnitude of migration in both simulations is quite variable and seems to depend strongly on interactions of the planet with discrete modes in the disk.
 In $t = 10$ planet run, the overall radial migration rate is similar but is even more erratic in direction than in the $t = 0$ planet run.
-The important departures from laminar disk theory are the more erratic nature of migration, the importance of interactions with discrete low-order modes, and stalling of migration near the inner edge of the strongest GI activity.
+The important departures from laminar disk theory are threefold: the more erratic nature of migration, the importance of interactions with discrete low-order modes, and stalling of migration near the inner edge of the strongest GI activity.
 The migration is neither a random walk nor monotonic but has a chaotic character due to nonlinear interactions.
@@
 We plan to extend simulations like these to much longer times in the future.
-However, the final radial positions of both planets happen to be the location of the Inner Lindblad resonance (ILR) for a strong m = 2 mode present in both runs.
+However, the final radial positions of both planets happen to be the location of the inner Lindblad resonance (ILR) for a strong m = 2 mode present in both runs.
 This ILR is associated with a surface density enhancement in the disk.

### 2011-06-07 Text/migration.tex
Final submitted version of the migration letter.
	modified:   Text/Figures/amt.eps
	new file:   Text/Figures/amt_bw.eps
	modified:   Text/Figures/planeta.eps
	deleted:    Text/Figures/planeta.pdf
	new file:   Text/Figures/planeta_bw.eps
	modified:   Text/Figures/planete.eps
	deleted:    Text/Figures/planete.pdf
	new file:   Text/Figures/planete_bw.eps
	modified:   Text/general.bib
	modified:   Text/migration.tex
	copied:     Text/migration.tex -> Text/migration_apjlsubmission.tex

git-svn-id: http://jaga.astro.indiana.edu/paperrepo/ApJL/Migration2010@86 3b7f7548-5c66-0410-83ab-3997167635fc

@@
 As part of a larger study, we are using a 3D radiative hydrodynamics code to investigate how an embedded gas giant planet interacts with a gas disk that undergoes gravitational instabilities (GIs).
-This Letter presents preliminary results from simulations with a Jupiter-mass planet placed in orbit at 25 AU within a 0.14 $M_{\odot}$ disk.
+This Letter presents results from simulations with a Jupiter-mass planet placed in orbit at 25 AU within a 0.14 $M_{\odot}$ disk.
 The disk spans 5 to 40 AU around a 1 $M_{\odot}$ star and is initially marginally unstable.
@@
 However, radiative transport, detailed equations of state (EOS), and the self-gravity of the gas disk have usually been ignored; the effects of a non-isothermal EOS have only recently been included \citep[e.g.,][]{paardekooper2006,paardekooper2010,paardekooper2011}.
+Emerging studies, such as \citet{baruteau2011} are just beginning to address some of these issues.
 \citet{boss2005} and \citet{mayer2004} examined radial migration of planet-mass fragments in gravitationally unstable disks, but their disks were violently disrupted by fragmentation under conditions (radii $<$ 40 AU, disk-to-star mass ratios $M_d/M_s \sim 0.1$, and stellar mass $M_s \sim 1\;M_{\odot}$) where fragmentation may not actually occur  \citep{rafikov2005,rafikov2007,boley2006,boley2007b,boley2008,forgan2009,cai2010}.
@@
 To improve our understanding of how planets migrate in GI-active disks that are not fragmenting, we have begun a systematic study, using numerical 3D radiative hydrodynamics, where we investigate the effects on both the disk and the planet of inserting a planet-mass object into disks susceptible to GIs.
-Using techniques deployed in earlier research by our group \citep{pickett2003, mejia2005,cai2006,cai2008,boley2006,boley2007b,michael2010}, we can identify the dominant spiral waves in a simulation and analyze how the waves interact with the planet's motion.
+Using techniques developed in earlier research by our group \citep{pickett2003, mejia2005,cai2006,cai2008,boley2006,boley2007b,michael2010}, we can identify the dominant spiral waves in a simulation and analyze how the waves interact with the planet's motion.
 Our goal is to determine both the effect of giant planets on GIs and the effect of GIs on planet migration.
@@
 
-Section 2 below presents our numerical methods and initial conditions.  We describe the simulation results in Section 3, and discuss them in Section 4.
+Section 2 below presents our numerical methods and initial conditions.  We describe the simulation results in \S3, and discuss them in \S4.
 
@@
 
-The model disk, based on an equlibrium model in \citet{pickett2003}, orbits a 1 $M_{\odot}$ star and has a mass $M_d = 0.14 M_{\odot}$, inner and outer radii at 5 and 40 AU, and an initial surface density distribution $\Sigma \sim \varpi^{-1/2}$.
+The model disk, based on an equlibrium model from \citet{pickett2003}, orbits a 1 $M_{\odot}$ star and has a mass $M_d = 0.14 M_{\odot}$, inner and outer radii at 5 and 40 AU, and an initial surface density distribution $\Sigma \sim \varpi^{-1/2}$.
 The time unit of one ORP (= outer rotation period) is defined as the rotation period of the initial disk at $\varpi \approx$ 32 AU, or about 180 yr.
@@
 An outflow boundary condition is enforced at the upper vertical grid boundary, the outer radial grid boundary, and an inner radial boundary at 2 AU.
-To seed nonaxisymmetry, the density is given an initial 0.01 \% random cell-to-cell perturbation.
+To seed nonaxisymmetry, the density distribution is given an initial 0.01 \% random cell-to-cell perturbation.
 
 This Letter presents three simulations.
-The first, which we call the {\sl fiducial run}, goes from time $t = 0$ to 21 ORP without a planet.
+The first, which we call the {\sl fiducial run},  is simulated from time $t = 0$ to 21 ORP without a planet.
 In the second simulation, which we call the {\sl t = 0 planet run}, a one Jupiter-mass ($M_J$) planet with a roughly circular midplane orbit is included in the initial $t = 0$ ORP equilibrium disk at 25 AU.
@@
 These three simulations are part of a larger suite, to be described elsewhere, in which the planet mass is varied.
-We estimate crude semi-major axes $a$ and eccentricities $e$ from the planet's $\varpi(t)$ by using the maximum and minimum radii, $\varpi_{\rm max}$ and $\varpi_{\rm min}$, attained by the planet over each interval that the planet's $\phi$-coordinate changes by 2$\pi$ (one ``orbit'').
-Then, for that orbit, $a = (\varpi_{\rm max}+\varpi_{\rm min})/2$ and $e = (\varpi_{\rm max}-\varpi_{\rm min})/(\varpi_{\rm max}+\varpi_{\rm min})$.
-These parameters are only indicative; there are strong orbit to orbit variations in $a$ and $e$.
+We estimate semi-major axes $a$ and eccentricities $e$ from the planet's $\varpi(t)$ by using the maximum and minimum radii, $\varpi_{\rm max}$ and $\varpi_{\rm min}$, over each complete orbit of the planet, where one orbit is $\Delta\phi=2\pi$.
+Specifically, $a = (\varpi_{\rm max}+\varpi_{\rm min})/2$ and $e = (\varpi_{\rm max}-\varpi_{\rm min})/(\varpi_{\rm max}+\varpi_{\rm min})$.
+These parameters are intended to highlight the strong orbit-to-orbit variations in the motion of the planet.
+A more refined estimate of the orbital elements is unnecessary because the potential of the disk, even if were axisymmetric, will cause departures from Keplerian dynamics.
 
@@
 Between $t = 15$ and 19 ORP, the pattern of outward migration for 2 ORP followed by an inward plunge over the next 2 ORP resembles the behavior in the $t = 0$ planet run during the burst, but, in this case, there is no distinct transition between phases of GI activity.
-Animations of the simulation (available at http://hdl.handle.net/2022/13304) show there is a complex interaction between the planet, its spiral wake, and the global spiral arms caused by the instabilities.
+Animations of the evolution of the midplane density (available at http://hdl.handle.net/2022/13304) show there is a complex interaction between the planet, its spiral wake, and the global spiral arms caused by the instabilities.
 Analysis of images and periodicities suggests that, between 15 and 19 ORP, the planet may be in a mean motion resonance with $m = 2$ and 3 patterns, both of which have CR at 27 AU.
@@
 
-{\it Summary.} The simulations in this Letter show that a planet of even only 1$M_J$ placed near the $Q$-minimum in a 0.14$M_{\odot}$ disk can hasten the onset of gravitational instabilities and change the nature of the dominant modes during the initial burst.
+{\it Summary.} The simulations in this Letter show that a planet of only 1$M_J$ placed near the $Q$-minimum in a 0.14$M_{\odot}$ disk can hasten the onset of gravitational instabilities and change the nature of the dominant modes during the initial burst.
 Although the magnitude of radial migration (a few$\times10^{-3}$ AU yr$^{-1}$) is not too different from that computed using the \citet{tanaka2002}  isothermal disk formula based on the Lindblad torques, it can fluctuate in sign on time scales on the order of the orbit period and can be strongly affected by gravitational interactions of the planet with discrete GI-driven spiral modes, leading to phases of rapid radial motion.
@@
 ACI-0338618l, OCI-0451237, OCI-0535258, OCI-0504075, and CNS-0521433, and through TeraGrid resources provided by Indiana University, PSC, SDSC, and NCSA.
-TeraGrid systems are hosted by Indiana University, LONI, NCAR, NCSA, NICS, ORNL, PSC, Purdue University, SDSC, TACC, and UC/ANL..
+TeraGrid systems are hosted by Indiana University, LONI, NCAR, NCSA, NICS, ORNL, PSC, Purdue University, SDSC, TACC, and UC/ANL.
 Any opinions, findings and conclusions or recommendations expressed in this material are those of the authors and do not necessarily reflect the views of the National Science Foundation.
@@
 \includegraphics[width=11cm]{Figures/amt.eps}
-\caption{Global amplitudes of nonaxisymmetric density perturbations $A_m$ as a function of time for individual cos($m\phi$) perturbations.
+\caption{(Color figure available in the online version) Global amplitudes of nonaxisymmetric density perturbations $A_m$ as a function of time for individual cos($m\phi$) perturbations.
 The formula for computing $A_m$ is equation (15) of \citep{boley2006}.
 Here the integrals are done only over the disk outside 15 AU to suppress contributions from edge effects in the inner disk.
-The set of red and black curves correspond to the fiducial run (no planet) and the $t = 0$ planet run, respectively.}
+The top and bottom panels correspond to the fiducial run (no planet) and the $t = 0$ planet run, respectively.}
 \label{fig:Am}
@@
 \caption{Plot of the planet's radial position as a function of time $\varpi_p(t)$ for the two planet simulations.
-The red and black indicate the curves ($\varpi(t)$), points ($a$), and horizontal scales ($t$) that correspond to the $t = 0$ and $t = 10$ planet runs, respectively.
-The diamonds are the approximate semi-major axes $a$ computed for each 0 to 2$\phi$ change in azimuth of the planet as explained in the text.}
+The dashed-diamond and solid-star lines (colored red and black in the online version) indicate the curves ($\varpi(t)$), points ($a$), and horizontal scales ($t$) that correspond to the $t = 0$ and $t = 10$ planet runs, respectively.
+The symbols are the approximate semi-major axes $a$ computed for each 0 to 2$\phi$ change in azimuth of the planet as explained in the text.}
 \label{fig:a}
@@
 \caption{Plot of the eccentricities $e$, computed for each 0 to 2$\phi$ change in azimuth of the planet as explained in the text, as a function of time.
-The red and black indicate the points ($e$) and horizontal scales ($t$) that correspond to the $t = 0$ and $t = 10$ planet runs, respectively.}
+The diamonds and stars (colored red and black in the online version) indicate the points ($e$) and horizontal scales ($t$) that correspond to the $t = 0$ and $t = 10$ planet runs, respectively.}
 \label{fig:e}

### 2011-06-07 Text/migration_apjlsubmission.tex
Final submitted version of the migration letter.
	modified:   Text/Figures/amt.eps
	new file:   Text/Figures/amt_bw.eps
	modified:   Text/Figures/planeta.eps
	deleted:    Text/Figures/planeta.pdf
	new file:   Text/Figures/planeta_bw.eps
	modified:   Text/Figures/planete.eps
	deleted:    Text/Figures/planete.pdf
	new file:   Text/Figures/planete_bw.eps
	modified:   Text/general.bib
	modified:   Text/migration.tex
	copied:     Text/migration.tex -> Text/migration_apjlsubmission.tex

git-svn-id: http://jaga.astro.indiana.edu/paperrepo/ApJL/Migration2010@86 3b7f7548-5c66-0410-83ab-3997167635fc

--- 2011-06-07 Text/migration.tex
@@
 
-\bibliographystyle{apj}
-\bibliography{general}
+\begin{thebibliography}{48}
+\expandafter\ifx\csname natexlab\endcsname\relax\def\natexlab#1{#1}\fi
+
+\bibitem[{{Baruteau} {et~al.}(2011){Baruteau}, {Meru}, \&
+  {Paardekooper}}]{baruteau2011}
+{Baruteau}, C., {Meru}, F., \& {Paardekooper}, S.-J. 2011, ArXiv e-prints
+
+\bibitem[{{Boley}(2009)}]{boley2009}
+{Boley}, A.~C. 2009, ApJ, 695, L53
+
+\bibitem[{{Boley} \& {Durisen}(2006)}]{boley2006a}
+{Boley}, A.~C. \& {Durisen}, R.~H. 2006, ApJ, 641, 534
+
+\bibitem[{{Boley} \& {Durisen}(2008)}]{boley2008}
+---. 2008, ApJ, 685, 1193
+
+\bibitem[{{Boley} \& {Durisen}(2010)}]{boley2010b}
+---. 2010, \apj, 724, 618
+
+\bibitem[{{Boley} {et~al.}(2007{\natexlab{a}}){Boley}, {Durisen}, {Nordlund},
+  \& {Lord}}]{boley2007b}
+{Boley}, A.~C., {Durisen}, R.~H., {Nordlund}, {\AA}., \& {Lord}, J.
+  2007{\natexlab{a}}, ApJ, 665, 1254
+
+\bibitem[{{Boley} {et~al.}(2007{\natexlab{b}}){Boley}, {Hartquist}, {Durisen},
+  \& {Michael}}]{boley2007}
+{Boley}, A.~C., {Hartquist}, T.~W., {Durisen}, R.~H., \& {Michael}, S.
+  2007{\natexlab{b}}, ApJL, 656, L89
+
+\bibitem[{{Boley} {et~al.}(2010){Boley}, {Hayfield}, {Mayer}, \&
+  {Durisen}}]{boley2010}
+{Boley}, A.~C., {Hayfield}, T., {Mayer}, L., \& {Durisen}, R.~H. 2010, Icarus,
+  207, 509
+
+\bibitem[{{Boley} {et~al.}(2006){Boley}, {Mej{\'{\i}}a}, {Durisen}, {Cai},
+  {Pickett}, \& {D'Alessio}}]{boley2006}
+{Boley}, A.~C., {Mej{\'{\i}}a}, A.~C., {Durisen}, R.~H., {Cai}, K., {Pickett},
+  M.~K., \& {D'Alessio}, P. 2006, ApJ, 651, 517
+
+\bibitem[{{Boss}(2005)}]{boss2005}
+{Boss}, A.~P. 2005, ApJ, 629, 535
+
+\bibitem[{{Cai} {et~al.}(2008){Cai}, {Durisen}, {Boley}, {Pickett}, \&
+  {Mej{\'{\i}}a}}]{cai2008}
+{Cai}, K., {Durisen}, R.~H., {Boley}, A.~C., {Pickett}, M.~K., \&
+  {Mej{\'{\i}}a}, A.~C. 2008, ApJ, 673, 1138
+
+\bibitem[{{Cai} {et~al.}(2006){Cai}, {Durisen}, {Michael}, {Boley},
+  {Mej{\'{\i}}a}, {Pickett}, \& {D'Alessio}}]{cai2006}
+{Cai}, K., {Durisen}, R.~H., {Michael}, S., {Boley}, A.~C., {Mej{\'{\i}}a},
+  A.~C., {Pickett}, M.~K., \& {D'Alessio}, P. 2006, ApJL, 636, L149
+
+\bibitem[{{Cai} {et~al.}(2010){Cai}, {Pickett}, {Durisen}, \&
+  {Milne}}]{cai2010}
+{Cai}, K., {Pickett}, M.~K., {Durisen}, R.~H., \& {Milne}, A.~M. 2010, \apjl,
+  716, L176
+
+\bibitem[{{Clarke}(2009)}]{clarke2009}
+{Clarke}, C.~J. 2009, MNRAS, 612
+
+\bibitem[{{Cossins} {et~al.}(2009){Cossins}, {Lodato}, \&
+  {Clarke}}]{cossins2009}
+{Cossins}, P., {Lodato}, G., \& {Clarke}, C.~J. 2009, \mnras, 393, 1157
+
+\bibitem[{{D'Alessio} {et~al.}(2001){D'Alessio}, {Calvet}, \&
+  {Hartmann}}]{dalessio2001}
+{D'Alessio}, P., {Calvet}, N., \& {Hartmann}, L. 2001, ApJ, 553, 321
+
+\bibitem[{{Dodson-Robinson} {et~al.}(2009){Dodson-Robinson}, {Veras}, {Ford},
+  \& {Beichman}}]{dodson2009}
+{Dodson-Robinson}, S.~E., {Veras}, D., {Ford}, E.~B., \& {Beichman}, C.~A.
+  2009, ApJ, 707, 79
+
+\bibitem[{{Durisen} {et~al.}(2007){Durisen}, {Boss}, {Mayer}, {Nelson},
+  {Quinn}, \& {Rice}}]{durisen2007}
+{Durisen}, R.~H., {Boss}, A.~P., {Mayer}, L., {Nelson}, A.~F., {Quinn}, T., \&
+  {Rice}, W.~K.~M. 2007, Protostars and Planets V, 607
+
+\bibitem[{{Forgan} {et~al.}(2009){Forgan}, {Rice}, {Stamatellos}, \&
+  {Whitworth}}]{forgan2009}
+{Forgan}, D., {Rice}, K., {Stamatellos}, D., \& {Whitworth}, A. 2009, MNRAS,
+  394, 882
+
+\bibitem[{{Gammie}(2001)}]{gammie2001}
+{Gammie}, C.~F. 2001, ApJ, 553, 174
+
+\bibitem[{{Goldreich} {et~al.}(2004){Goldreich}, {Lithwick}, \&
+  {Sari}}]{goldreich2004}
+{Goldreich}, P., {Lithwick}, Y., \& {Sari}, R. 2004, ApJ, 614, 497
+
+\bibitem[{{Hut} {et~al.}(1995){Hut}, {Makino}, \& {McMillan}}]{hut1995}
+{Hut}, P., {Makino}, J., \& {McMillan}, S. 1995, ApJL, 443, L93
+
+\bibitem[{{Krumholz} {et~al.}(2007){Krumholz}, {Klein}, \&
+  {McKee}}]{krumholz2007}
+{Krumholz}, M.~R., {Klein}, R.~I., \& {McKee}, C.~F. 2007, \apj, 656, 959
+
+\bibitem[{{Li} {et~al.}(2001){Li}, {Colgate}, {Wendroff}, \& {Liska}}]{li2001}
+{Li}, H., {Colgate}, S.~A., {Wendroff}, B., \& {Liska}, R. 2001, \apj, 551, 874
+
+\bibitem[{{Li} {et~al.}(2000){Li}, {Finn}, {Lovelace}, \& {Colgate}}]{li2000}
+{Li}, H., {Finn}, J.~M., {Lovelace}, R.~V.~E., \& {Colgate}, S.~A. 2000, \apj,
+  533, 1023
+
+\bibitem[{{Mayer} {et~al.}(2004){Mayer}, {Quinn}, {Wadsley}, \&
+  {Stadel}}]{mayer2004}
+{Mayer}, L., {Quinn}, T., {Wadsley}, J., \& {Stadel}, J. 2004, ApJ, 609, 1045
+
+\bibitem[{{Mej{\'{\i}}a} {et~al.}(2005){Mej{\'{\i}}a}, {Durisen}, {Pickett}, \&
+  {Cai}}]{mejia2005}
+{Mej{\'{\i}}a}, A.~C., {Durisen}, R.~H., {Pickett}, M.~K., \& {Cai}, K. 2005,
+  ApJ, 619, 1098
+
+\bibitem[{{Meschiari} \& {Laughlin}(2008)}]{meschiari2008}
+{Meschiari}, S. \& {Laughlin}, G. 2008, \apjl, 679, L135
+
+\bibitem[{{Michael} \& {Durisen}(2010)}]{michael2010}
+{Michael}, S. \& {Durisen}, R.~H. 2010, \mnras, 406, 279
+
+\bibitem[{{Nelson} {et~al.}(2000){Nelson}, {Benz}, \&
+  {Ruzmaikina}}]{nelson2000a}
+{Nelson}, A.~F., {Benz}, W., \& {Ruzmaikina}, T.~V. 2000, ApJ, 529, 357
+
+\bibitem[{{Paardekooper} {et~al.}(2010){Paardekooper}, {Baruteau}, {Crida}, \&
+  {Kley}}]{paardekooper2010}
+{Paardekooper}, S., {Baruteau}, C., {Crida}, A., \& {Kley}, W. 2010, \mnras,
+  401, 1950
+
+\bibitem[{{Paardekooper} \& {Mellema}(2006)}]{paardekooper2006}
+{Paardekooper}, S. \& {Mellema}, G. 2006, A\&A, 459, L17
+
+\bibitem[{{Paardekooper} \& {Papaloizou}(2009)}]{paardekooper2009}
+{Paardekooper}, S. \& {Papaloizou}, J.~C.~B. 2009, MNRAS, 394, 2283
+
+\bibitem[{{Paardekooper} {et~al.}(2011){Paardekooper}, {Baruteau}, \&
+  {Kley}}]{paardekooper2011}
+{Paardekooper}, S.-J., {Baruteau}, C., \& {Kley}, W. 2011, \mnras, 410, 293
+
+\bibitem[{{Papaloizou} {et~al.}(2007){Papaloizou}, {Nelson}, {Kley}, {Masset},
+  \& {Artymowicz}}]{papaloizou2007}
+{Papaloizou}, J.~C.~B., {Nelson}, R.~P., {Kley}, W., {Masset}, F.~S., \&
+  {Artymowicz}, P. 2007, Protostars and Planets V, 655
+
+\bibitem[{{Pickett} {et~al.}(2003){Pickett}, {Mej{\'{\i}}a}, {Durisen},
+  {Cassen}, {Berry}, \& {Link}}]{pickett2003}
+{Pickett}, B.~K., {Mej{\'{\i}}a}, A.~C., {Durisen}, R.~H., {Cassen}, P.~M.,
+  {Berry}, D.~K., \& {Link}, R.~P. 2003, ApJ, 590, 1060
+
+\bibitem[{{Rafikov}(2005)}]{rafikov2005}
+{Rafikov}, R.~R. 2005, ApJL, 621, L69
+
+\bibitem[{{Rafikov}(2007)}]{rafikov2007}
+---. 2007, ApJ, 662, 642
+
+\bibitem[{{Rafikov}(2009)}]{rafikov2009}
+---. 2009, \apj, 704, 281
+
+\bibitem[{{Stamatellos} {et~al.}(2007){Stamatellos}, {Hubber}, \&
+  {Whitworth}}]{stamatellos2007}
+{Stamatellos}, D., {Hubber}, D.~A., \& {Whitworth}, A.~P. 2007, \mnras, 382,
+  L30
+
+\bibitem[{{Stamatellos} \& {Whitworth}(2009)}]{stamatellos2009}
+{Stamatellos}, D. \& {Whitworth}, A.~P. 2009, MNRAS, 392, 413
+
+\bibitem[{{Tanaka} {et~al.}(2002){Tanaka}, {Takeuchi}, \& {Ward}}]{tanaka2002}
+{Tanaka}, H., {Takeuchi}, T., \& {Ward}, W.~R. 2002, \apj, 565, 1257
+
+\bibitem[{{Vorobyov} \& {Basu}(2006)}]{vorobyov2006}
+{Vorobyov}, E.~I. \& {Basu}, S. 2006, ApJ, 650, 956
+
+\bibitem[{{Vorobyov} \& {Basu}(2010{\natexlab{a}})}]{vorobyov2010}
+---. 2010{\natexlab{a}}, \apjl, 714, L133
+
+\bibitem[{{Vorobyov} \& {Basu}(2010{\natexlab{b}})}]{vorobyov2010b}
+---. 2010{\natexlab{b}}, \apj, 719, 1896
+
+\bibitem[{{Ward} \& {Hahn}(1998)}]{ward1998}
+{Ward}, W.~R. \& {Hahn}, J.~M. 1998, AJ, 116, 489
+
+\bibitem[{{Ward} \& {Hahn}(2003)}]{ward2003}
+---. 2003, AJ, 125, 3389
+
+\bibitem[{{Zhu} {et~al.}(2010){Zhu}, {Hartmann}, \& {Gammie}}]{zhu2010}
+{Zhu}, Z., {Hartmann}, L., \& {Gammie}, C. 2010, ApJ, 713, 1143
+
+\end{thebibliography}
 
@@
 
-%***DICK: Scott, remember that figures need some improvements.
-
 \begin{figure}[t]
 \center
-\includegraphics[width=11cm]{Figures/amt.eps}
+\includegraphics[width=11cm]{amt.eps}
 \caption{(Color figure available in the online version) Global amplitudes of nonaxisymmetric density perturbations $A_m$ as a function of time for individual cos($m\phi$) perturbations.
@@
 \center
-\includegraphics[width=12cm]{Figures/planeta.eps}
+\includegraphics[width=12cm]{planeta.eps}
 \caption{Plot of the planet's radial position as a function of time $\varpi_p(t)$ for the two planet simulations.
@@
 \center
-\includegraphics[width=12cm]{Figures/planete.eps}
+\includegraphics[width=12cm]{planete.eps}
 \caption{Plot of the eccentricities $e$, computed for each 0 to 2$\phi$ change in azimuth of the planet as explained in the text, as a function of time.

### 2011-06-23 Text/migration.tex
Final, Final, Final edits. Added Dick's comment and submitted 06/07/2011
	modified:   Text/migration.tex
	modified:   Text/migration_apjlsubmission.tex

git-svn-id: http://jaga.astro.indiana.edu/paperrepo/ApJL/Migration2010@87 3b7f7548-5c66-0410-83ab-3997167635fc
Changes prompted by first review by referee
Added paragraph in discussion section; Added h/r aspect ratio plot for all runs at 2 different times.

@@
 \begin{abstract} 
-Characterization of migration in gravitationally unstable disks is necessary to understand the fate of protoplanets formed by disk instability.
-As part of a larger study, we are using a 3D radiative hydrodynamics code to investigate how an embedded gas giant planet interacts with a gas disk that undergoes gravitational instabilities (GIs).
-This Letter presents results from simulations with a Jupiter-mass planet placed in orbit at 25 AU within a 0.14 $M_{\odot}$ disk.
-The disk spans 5 to 40 AU around a 1 $M_{\odot}$ star and is initially marginally unstable.
-In one simulation, the planet is inserted prior to the eruption of GIs; in another, it is inserted only after the disk has settled into a quasi-steady GI-active state, where heating by GIs roughly balances radiative cooling.
-When the planet is present from the beginning, its own wake stimulates growth of a particular global mode with which it strongly interacts, and the planet plunges inward six AU in about 10$^3$ years.
-In both cases with embedded planets, there are times when the planet's radial motion is slow and varies in direction.
-At other times, when the planet appears to be interacting with strong spiral modes, migration both inward and outward can be relatively rapid, covering several AUs over hundreds of years.
-Migration in both cases appears to stall near the inner Lindblad resonance of a dominant low-order mode.
-Planet orbit eccentricities fluctuate rapidly between about 0.02 to 0.1 throughout the GI-active phases of the simulations.
+  Characterization of migration in gravitationally unstable disks is necessary to understand the fate of protoplanets
+  formed by disk instability. As part of a larger study, we are using a 3D radiative hydrodynamics code to investigate
+  how an embedded gas giant planet interacts with a gas disk that undergoes gravitational instabilities (GIs). This
+  Letter presents results from simulations with a Jupiter-mass planet placed in orbit at 25 AU within a 0.14 $M_{\odot}$
+  disk. The disk spans 5 to 40 AU around a 1 $M_{\odot}$ star and is initially marginally unstable. In one simulation,
+  the planet is inserted prior to the eruption of GIs; in another, it is inserted only after the disk has settled into a
+  quasi-steady GI-active state, where heating by GIs roughly balances radiative cooling. When the planet is present from
+  the beginning, its own wake stimulates growth of a particular global mode with which it strongly interacts, and the
+  planet plunges inward six AU in about 10$^3$ years. In both cases with embedded planets, there are times when the
+  planet's radial motion is slow and varies in direction. At other times, when the planet appears to be interacting with
+  strong spiral modes, migration both inward and outward can be relatively rapid, covering several AUs over hundreds of
+  years. Migration in both cases appears to stall near the inner Lindblad resonance of a dominant low-order mode. Planet
+  orbit eccentricities fluctuate rapidly between about 0.02 to 0.1 throughout the GI-active phases of the simulations.
 \end{abstract}
@@
 
-Migration of gas giant planets due to interactions with a circumstellar gas disk can play a major role in defining the architecture of planetary systems.
-Work on migration \citep[see review by][]{papaloizou2007} has included gravitational interaction of planets with both laminar and turbulent disks.
-However, radiative transport, detailed equations of state (EOS), and the self-gravity of the gas disk have usually been ignored; the effects of a non-isothermal EOS have only recently been included \citep[e.g.,][]{paardekooper2006,paardekooper2010,paardekooper2011}.
-Emerging studies, such as \citet{baruteau2011} are just beginning to address some of these issues.
-\citet{boss2005} and \citet{mayer2004} examined radial migration of planet-mass fragments in gravitationally unstable disks, but their disks were violently disrupted by fragmentation under conditions (radii $<$ 40 AU, disk-to-star mass ratios $M_d/M_s \sim 0.1$, and stellar mass $M_s \sim 1\;M_{\odot}$) where fragmentation may not actually occur  \citep{rafikov2005,rafikov2007,boley2006,boley2007b,boley2008,forgan2009,cai2010}.
-More recently, fragmentation into clumps with gas giant or brown dwarf masses has been documented in numerical simulations of disks that are relatively massive ($M_d/M_s \sim$ a few tenths) and spatially extended (outer radii $>$ 50 AU) \citep{krumholz2007,stamatellos2007,stamatellos2009,boley2009,boley2010}, where fragmentation is expected from semi-analytic arguments \citep[e.g.,][]{clarke2009,rafikov2009,dodson2009}.
-The fate of the clumps then depends in part on their radial migration, which is a chaotic and messy affair in a fragmenting disk \citep[e.g.,][]{boley2009,boley2010,vorobyov2010,boley2010b}.
-The occurrence of gravitational instabilities (GIs) may be episodic \citep[e.g.,][]{vorobyov2006,vorobyov2010b,zhu2010}.
-Clumps that survive and contract to the dimensions of young planets can later find themselves in a disk that erupts again into GI activity.
-As the star/disk system evolves, such a protoplanet may end up in a region of a GI-active disk where fragmentation does not occur.
+Migration of gas giant planets due to interactions with a circumstellar gas disk can play a major role in defining the
+architecture of planetary systems. Work on migration \citep[see review by][]{papaloizou2007} has included gravitational
+interaction of planets with both laminar and turbulent disks. However, radiative transport, detailed equations of state
+(EOS), and the self-gravity of the gas disk have usually been ignored; the effects of a non-isothermal EOS have only
+recently been included \citep[e.g.,][]{paardekooper2006,paardekooper2010,paardekooper2011}. Emerging studies, such as
+\citet{baruteau2011} and the one reported here, are beginning to address some of these issues.  \citet{boss2005} and
+\citet{mayer2004} examined radial migration of planet-mass fragments in gravitationally unstable disks, but their disks
+were violently disrupted by fragmentation under conditions (radii $<$ 40 AU, disk-to-star mass ratios $M_d/M_s \sim
+0.1$, and stellar mass $M_s \sim 1\;M_{\odot}$) where fragmentation may not actually occur
+\citep{rafikov2005,rafikov2007,boley2006,boley2007b,boley2008,forgan2009,cai2010}. More recently, fragmentation into
+clumps with gas giant or brown dwarf masses has been documented in numerical simulations of disks that are relatively
+massive ($M_d/M_s \sim$ a few tenths) and spatially extended (outer radii $>$ 50 AU)
+\citep{krumholz2007,stamatellos2007,stamatellos2009,boley2009,boley2010}, where fragmentation is expected from
+semi-analytic arguments \citep[e.g.,][]{clarke2009,rafikov2009,dodson2009}. The fate of the clumps then depends in part
+on their radial migration, which is a chaotic and messy affair in a fragmenting disk
+\citep[e.g.,][]{boley2009,boley2010,vorobyov2010,boley2010b}. The occurrence of gravitational instabilities (GIs) may be
+episodic \citep[e.g.,][]{vorobyov2006,vorobyov2010b,zhu2010}. Clumps that survive and contract to the dimensions of
+young planets can later find themselves in a disk that erupts again into GI activity. As the star/disk system evolves,
+such a protoplanet may end up in a region of a GI-active disk where fragmentation does not occur.
 
-To improve our understanding of how planets migrate in GI-active disks that are not fragmenting, we have begun a systematic study, using numerical 3D radiative hydrodynamics, where we investigate the effects on both the disk and the planet of inserting a planet-mass object into disks susceptible to GIs.
-Using techniques developed in earlier research by our group \citep{pickett2003, mejia2005,cai2006,cai2008,boley2006,boley2007b,michael2010}, we can identify the dominant spiral waves in a simulation and analyze how the waves interact with the planet's motion.
-Our goal is to determine both the effect of giant planets on GIs and the effect of GIs on planet migration.
-Because GIs are sensitive to radiative physics, we use a well-tested radiative scheme \citep{boley2007b} and realistic opacities \citep{dalessio2001}.
+To improve our understanding of how planets migrate in GI-active disks that are not fragmenting, we have begun a
+systematic study, using numerical 3D radiative hydrodynamics, where we investigate the effects on both the disk and the
+planet of inserting a planet-mass object into disks susceptible to GIs. Using techniques developed in earlier research
+by our group \citep{pickett2003, mejia2005,cai2006,cai2008,boley2006,boley2007b,michael2010}, we can identify the
+dominant spiral waves in a simulation and analyze how the waves interact with the planet's motion. Our goal is to
+determine both the effect of giant planets on GIs and the effect of GIs on planet migration. Because GIs are sensitive
+to radiative physics, we use a well-tested radiative scheme \citep{boley2007b} and realistic opacities
+\citep{dalessio2001}.
 
-Section 2 below presents our numerical methods and initial conditions.  We describe the simulation results in \S3, and discuss them in \S4.
+Section 2 below presents our numerical methods and initial conditions.  We describe the simulation results in \S3, and
+discuss them in \S4.
 
@@
 
-The CHYMERA (Computational HYdrodynamics with MultiplE Radiation Algorithms) code \citep{boley2007b} is a second-order, explicit, Eulerian scheme on a 3D cylindrical grid.
-The code uses a realistic equation of state for H$_2$ \citep{boley2007} and integrates an energy equation that includes $PdV$ work, net heating or cooling due to radiative flux divergence, and heating by artificial bulk viscosity.
-Calculations are done on a uniform cylindrical grid with reflection symmetry about the disk midplane and a grid size ($\varpi$,$\phi$,z) = (512,512,64).
-The $z$-axis is the rotation axis of the disk.
-The large number of azimuthal zones is necessary to resolve the planet's Hill sphere and the planet's wake.
-These simulations utilize the radiative cooling scheme developed and tested in \citet{boley2007b}, where the optically thick monochromatic flux in the $\varpi$ and $\phi$-directions is computed by flux-limited radiative diffusion and where the radiative transport of energy in the $z$-direction is solved using a one-ray discrete ordinate method in both optically thin and thick regions.
-Although the central star remains fixed at the grid center, we account for acceleration of the reference frame by the planet and by the disk via indirect potentials, as in \citet{michael2010}.
-The planet integration is done with a Verlet integrator \citep[e.g.,][]{hut1995}, and the indirect potential terms are treated as in \citet{nelson2000a}.
-The Rosseland mean and Planck mean opacities and molecular weights in our simulations are the same as those in \citet{boley2006,boley2007b}, except that we correct the \citet{dalessio2001} gas mean molecular weight to $\mu$ = 2.33.
+The CHYMERA (Computational HYdrodynamics with MultiplE Radiation Algorithms) code \citep{boley2007b} is a second-order,
+explicit, Eulerian scheme on a 3D cylindrical grid. The code uses a realistic equation of state for H$_2$
+\citep{boley2007} and integrates an energy equation that includes $PdV$ work, net heating or cooling due to radiative
+flux divergence, and heating by artificial bulk viscosity. Calculations are done on a uniform cylindrical grid with
+reflection symmetry about the disk midplane and a grid size ($\varpi$,$\phi$,z) = (512,512,64).  The $z$-axis is the
+rotation axis of the disk. The large number of azimuthal zones is necessary to resolve the planet's Hill sphere and the
+planet's wake. These simulations utilize the radiative cooling scheme developed and tested in \citet{boley2007b}, where
+the optically thick monochromatic flux in the $\varpi$ and $\phi$-directions is computed by flux-limited radiative
+diffusion and where the radiative transport of energy in the $z$-direction is solved using a one-ray discrete ordinate
+method in both optically thin and thick regions. Although the central star remains fixed at the grid center, we account
+for acceleration of the reference frame by the planet and by the disk via indirect potentials, as in
+\citet{michael2010}. The planet integration is done with a Verlet integrator \citep[e.g.,][]{hut1995}, and the indirect
+potential terms are treated as in \citet{nelson2000a}. The Rosseland mean and Planck mean opacities and molecular
+weights in our simulations are the same as those in \citet{boley2006,boley2007b}, except that we correct the
+\citet{dalessio2001} gas mean molecular weight to $\mu$ = 2.33.
 
@@
 
-The model disk, based on an equlibrium model from \citet{pickett2003}, orbits a 1 $M_{\odot}$ star and has a mass $M_d = 0.14 M_{\odot}$, inner and outer radii at 5 and 40 AU, and an initial surface density distribution $\Sigma \sim \varpi^{-1/2}$.
-The time unit of one ORP (= outer rotation period) is defined as the rotation period of the initial disk at $\varpi \approx$ 32 AU, or about 180 yr.
-The disk is initially located between radial grid zones 30 and 240 and is close to isentropic, which results in a Toomre-$Q$ distribution with a marginally unstable \citep[see ][]{durisen2007} minimum value of 1.38 at radial grid zone 161 (26.7 AU).
-The computational grid extends radially to 512 zones to accommodate expansion of the outer disk once GIs become nonlinear.
-An outflow boundary condition is enforced at the upper vertical grid boundary, the outer radial grid boundary, and an inner radial boundary at 2 AU.
-To seed nonaxisymmetry, the density distribution is given an initial 0.01 \% random cell-to-cell perturbation.
+The model disk, based on an equilibrium model from \citet{pickett2003}, orbits a 1 $M_{\odot}$ star and has a mass $M_d =
+0.14 M_{\odot}$, inner and outer radii at 5 and 40 AU, and an initial surface density distribution $\Sigma \sim
+\varpi^{-1/2}$. The time unit of one ORP (= outer rotation period) is defined as the rotation period of the initial disk
+at $\varpi \approx$ 32 AU, or about 180 yr. The disk is initially located between radial grid zones 30 and 240 and is
+close to isentropic, which results in a Toomre-$Q$ distribution with a marginally unstable \citep[see ][]{durisen2007}
+minimum value of 1.38 at radial grid zone 161 (26.7 AU). The computational grid extends radially to 512 zones to
+accommodate expansion of the outer disk once GIs become nonlinear. An outflow boundary condition is enforced at the
+upper vertical grid boundary, the outer radial grid boundary, and an inner radial boundary at 2 AU. To seed
+nonaxisymmetry, the density distribution is given an initial 0.01 \% random cell-to-cell perturbation.
 
-This Letter presents three simulations.
-The first, which we call the {\sl fiducial run},  is simulated from time $t = 0$ to 21 ORP without a planet.
-In the second simulation, which we call the {\sl t = 0 planet run}, a one Jupiter-mass ($M_J$) planet with a roughly circular midplane orbit is included in the initial $t = 0$ ORP equilibrium disk at 25 AU.
-In the third simulation, which we call the {\sl t = 10 planet run}, a 1 $M_J$ planet with a similar orbit is inserted into the fiducial run at $t = 10.5$ ORP.
-These three simulations are part of a larger suite, to be described elsewhere, in which the planet mass is varied.
-We estimate semi-major axes $a$ and eccentricities $e$ from the planet's $\varpi(t)$ by using the maximum and minimum radii, $\varpi_{\rm max}$ and $\varpi_{\rm min}$, over each complete orbit of the planet, where one orbit is $\Delta\phi=2\pi$.
-Specifically, $a = (\varpi_{\rm max}+\varpi_{\rm min})/2$ and $e = (\varpi_{\rm max}-\varpi_{\rm min})/(\varpi_{\rm max}+\varpi_{\rm min})$.
-These parameters are intended to highlight the strong orbit-to-orbit variations in the motion of the planet.
-A more refined estimate of the orbital elements is unnecessary because the potential of the disk, even if were axisymmetric, will cause departures from Keplerian dynamics.
+This Letter presents three simulations. The first, which we call the {\sl fiducial run}, is simulated from time $t = 0$
+to 21 ORP without a planet. In the second simulation, which we call the {\sl t = 0 planet run}, a one Jupiter-mass
+($M_J$) planet with a roughly circular midplane orbit is included in the initial $t = 0$ ORP equilibrium disk at 25
+AU. In the third simulation, which we call the {\sl t = 10 planet run}, a 1 $M_J$ planet with a similar orbit is
+inserted into the fiducial run at $t = 10.5$ ORP. These three simulations are part of a larger suite, to be described
+elsewhere, in which the planet mass is varied. We estimate semi-major axes $a$ and eccentricities $e$ from the planet's
+$\varpi(t)$ by using the maximum and minimum radii, $\varpi_{\rm max}$ and $\varpi_{\rm min}$, over each complete orbit
+of the planet, where one orbit is $\Delta\phi=2\pi$. Specifically, $a = (\varpi_{\rm max}+\varpi_{\rm min})/2$ and $e =
+(\varpi_{\rm max}-\varpi_{\rm min})/(\varpi_{\rm max}+\varpi_{\rm min})$. These parameters are intended to highlight the
+strong orbit-to-orbit variations in the motion of the planet.  A more refined estimate of the orbital elements is
+unnecessary because the potential of the disk, even if were axisymmetric, will cause departures from Keplerian dynamics.
 
@@
 
-{\it The Fiducial Run.} As in \citet{mejia2005}, the fiducial run exhibits four distinct phases: initial {\sl axisymmetric cooling}, the onset of GIs in a well-defined {\sl burst} of a few discrete low-order modes, a {\sl transition} to more complex nonaxisymmetric structure, and the {\sl asymptotic} establishment of quasi-steady GI activity with an overall balance of heating and cooling.
-There is no fragmentation because this disk has a long cooling time \citep{gammie2001,boley2006,boley2007b}.
+{\it The Fiducial Run.} As in \citet{mejia2005}, the fiducial run exhibits four distinct phases: initial {\sl
+  axisymmetric cooling}, the onset of GIs in a well-defined {\sl burst} of a few discrete low-order modes, a {\sl
+  transition} to more complex nonaxisymmetric structure, and the {\sl asymptotic} establishment of quasi-steady GI
+activity with an overall balance of heating and cooling. There is no fragmentation because this disk has a long cooling
+time \citep{gammie2001,boley2006,boley2007b}.
 
-{\it The $t = 0$ Planet Run.} Figure \ref{fig:Am} shows that inclusion of even this modest mass planet ($0.7 \%$ of $M_d$) has a dramatic effect on the burst phase.
-Without a planet, when all modes grow from imposed noise, it takes 4 ORPs for coherent spiral modes to organize and grow to nonlinear amplitude.
-The growth is centered near the $Q$-minimum at $26$ AU.
-Modes of cos($m\phi$) symmetry with $m = 4$ and 5 dominate.
-An $m = 3$ (three-armed) spiral also grows, but somewhat more slowly, and lags the other modes in time.
-In the $t = 0$ planet run, the planet develops an organized wake within about an ORP in which $m = 3$ is a dominant component ($\sim 5$ \% global density perturbation).
-This nonlinear seeding causes $m = 3$ to dominate the GI burst, which also occurs $\sim$ 2 ORP earlier.
-Because of our initial placement of the planet inside but close to the $Q$-minimum, the corotation radius (CR) of the $m = 3$ mode is fairly close to that of the planet's orbit radius.
-When the triggered $m = 3$ mode becomes strongly nonlinear at about 3 ORP, planet migration is significantly affected.
+{\it The $t = 0$ Planet Run.} Figure \ref{fig:Am} shows that inclusion of even this modest mass planet ($0.7 \%$ of
+$M_d$) has a dramatic effect on the burst phase. Without a planet, when all modes grow from imposed noise, it takes 4
+ORPs for coherent spiral modes to organize and grow to nonlinear amplitude. The growth is centered near the $Q$-minimum
+at $26$ AU. Modes of cos($m\phi$) symmetry with $m = 4$ and 5 dominate. An $m = 3$ (three-armed) spiral also grows, but
+somewhat more slowly, and lags the other modes in time. In the $t = 0$ planet run, the planet develops an organized wake
+within about an ORP in which $m = 3$ is a dominant component ($\sim 5$ \% global density perturbation). This nonlinear
+seeding causes $m = 3$ to dominate the GI burst, which also occurs $\sim$ 2 ORP earlier. Because of our initial
+placement of the planet inside but close to the $Q$-minimum, the corotation radius (CR) of the $m = 3$ mode is fairly
+close to that of the planet's orbit radius. When the triggered $m = 3$ mode becomes strongly nonlinear at about 3 ORP,
+planet migration is significantly affected.
 
-Figure \ref{fig:a} shows the evolution of the planet's radial position. From 0 to 2.5 ORP, the planet is torqued primarily by its own wake and migrates inward. 
-Beginning at about 3 ORP, the planet interacts with the now nonlinear $m = 3$ GI mode.
-At first, the planet gains angular momentum and moves outward, but, from 4 to 8 ORP, a time interval of  about 720 yr, the plant experiences a negative torque and plunges from 23 to 17 AU.
-After $t = 8$ ORP, the main burst is over, and the disk transitions into its asymptotic state, where modes of many $m$-values become comparably strong (see Fig.
-\ref{fig:Am}).
-The planet's radial migration apparently stalls at about 16 to 17 AU.
-From an analysis of periodicities present in the gas disk between 8 and 12 ORP, the planet lies a few AU inside the inner Lindblad resonance of a strong $m = 2$ mode with CR at 29 AU in this early part of the asymptotic phase.
+Figure \ref{fig:a} shows the evolution of the planet's radial position. From 0 to 2.5 ORP, the planet is torqued
+primarily by its own wake and migrates inward.  Beginning at about 3 ORP, the planet interacts with the now nonlinear $m
+= 3$ GI mode. At first, the planet gains angular momentum and moves outward, but, from 4 to 8 ORP, a time interval of
+about 720 yr, the plant experiences a negative torque and plunges from 23 to 17 AU.  After $t = 8$ ORP, the main burst
+is over, and the disk transitions into its asymptotic state, where modes of many $m$-values become comparably strong
+(see Fig. \ref{fig:Am}). The planet's radial migration apparently stalls at about 16 to 17 AU. From an analysis of
+periodicities present in the gas disk between 8 and 12 ORP, the planet lies a few AU inside the inner Lindblad resonance
+of a strong $m = 2$ mode with CR at 29 AU in this early part of the asymptotic phase.
 
-{\it The $t = 10$ Planet Run.} The motion of the planet in this case is more difficult to interpret.
-We see intervals of fairly rapid inward or outward migration over several orbits, as well as times when radial migration appears to stall.
-Between $t = 15$ and 19 ORP, the pattern of outward migration for 2 ORP followed by an inward plunge over the next 2 ORP resembles the behavior in the $t = 0$ planet run during the burst, but, in this case, there is no distinct transition between phases of GI activity.
-Animations of the evolution of the midplane density (available at http://hdl.handle.net/2022/13304) show there is a complex interaction between the planet, its spiral wake, and the global spiral arms caused by the instabilities.
-Analysis of images and periodicities suggests that, between 15 and 19 ORP, the planet may be in a mean motion resonance with $m = 2$ and 3 patterns, both of which have CR at 27 AU.
-Over this time interval, the average orbit period of the planet is $\sim$ 0.50 ORP, while both the $m = 2$ and 3 spirals have a pattern period of $\sim 0.74$ ORP, suggesting a 3:2 resonance.
-Both of these modes are present, at nearly the same periodicities, in the fiducial run.
-So, in contrast with the $t = 0$ planet run, this planet's rapid phases of radial migration are caused by interaction with strong GI modes that exist independent of the planet.
+{\it The $t = 10$ Planet Run.} The motion of the planet in this case is more difficult to interpret. We see intervals of
+fairly rapid inward or outward migration over several orbits, as well as times when radial migration appears to
+stall. Between $t = 15$ and 19 ORP, the pattern of outward migration for 2 ORP followed by an inward plunge over the
+next 2 ORP resembles the behavior in the $t = 0$ planet run during the burst, but, in this case, there is no distinct
+transition between phases of GI activity. Animations of the evolution of the midplane density (available at
+http://hdl.handle.net/2022/13304) show there is a complex interaction between the planet, its spiral wake, and the
+global spiral arms caused by the instabilities. Analysis of images and periodicities suggests that, between 15 and 19
+ORP, the planet may be in a mean motion resonance with $m = 2$ and 3 patterns, both of which have CR at 27 AU. Over this
+time interval, the average orbit period of the planet is $\sim$ 0.50 ORP, while both the $m = 2$ and 3 spirals have a
+pattern period of $\sim 0.74$ ORP, suggesting a 3:2 resonance. Both of these modes are present, at nearly the same
+periodicities, in the fiducial run. So, in contrast with the $t = 0$ planet run, this planet's rapid phases of radial
+migration are caused by interaction with strong GI modes that exist independent of the planet.
 
-{\it Eccentricity.} Figure \ref{fig:e} shows the evolution of $e$ in the planet simulations.
-In both cases, the planets were inserted with an approximate circular velocity computed by adding the interior disk mass to the stellar point mass.
-The presence of GI activity in the $t = 10$ planet run immediately increases $e$ to $\sim 0.08$.
-In the $t = 0$ planet run, the modest initial $e$ decreases during the 1.5 ORP when it is migrating only due to its own wake, as expected \citep{ward1998,ward2003,goldreich2004}.
-However, once there are strong nonlinear interactions with GI modes, $e$ jumps upward.
-In both runs, it appears that interaction with well-established GI activity leads to eccentricities between 0.02 and 0.1 that vary in a chaotic way between these extremes on orbit period time scales.
-The magnitude of $e$ is roughly consistent with the ratio of the sound speed to orbital speed of the disk and the modest values for the Mach numbers of the GI spirals \citep{boley2006a,boley2008}.
-In other words, the planet's orbit tends to have about as much nonaxisymmetry as reflected in the pitch angles of the GI-induced spirals \citep{cossins2009}.
+{\it Eccentricity.} Figure \ref{fig:e} shows the evolution of $e$ in the planet simulations. In both cases, the planets
+were inserted with an approximate circular velocity computed by adding the interior disk mass to the stellar point
+mass. The presence of GI activity in the $t = 10$ planet run immediately increases $e$ to $\sim 0.08$. In the $t = 0$
+planet run, the modest initial $e$ decreases during the 1.5 ORP when it is migrating only due to its own wake, as
+expected \citep{ward1998,ward2003,goldreich2004}.  However, once there are strong nonlinear interactions with GI modes,
+$e$ jumps upward. In both runs, it appears that interaction with well-established GI activity leads to eccentricities
+between 0.02 and 0.1 that vary in a chaotic way between these extremes on orbit period time scales. The magnitude of $e$
+is roughly consistent with the ratio of the sound speed to orbital speed of the disk and the modest values for the Mach
+numbers of the GI spirals \citep{boley2006a,boley2008}. In other words, the planet's orbit tends to have about as much
+nonaxisymmetry as reflected in the pitch angles of the GI-induced spirals \citep{cossins2009}.
 
-{\it Migration Relative to the Disk.} The disk surface density distribution does not vary much in the asymptotic phase, so radial motion of the planet in the $t = 10$ planet run also represents radial migration relative to the gas.
-During the burst \citep[e.g.,][]{boley2006}, the surface density of the disk changes dramatically.
-To verify that the planet is really migrating relative to the background gas disk in the $t = 0$ run, we have compared the evolution of the planet's radial position $\varpi_p(t)$ with the mass inside that radius, expressed in terms of $m_{\rm cyl}$, the fraction of the disk mass interior to the planet.
-Between $t = 3.52$ and $3.89$ ORP, when $\varpi_p$ increases modestly, $m_{\rm cyl}$ increases from about 0.50 to 0.53; between 3.89 and 8.47 ORP, when $\varpi_p$ decreases dramatically, $m_{\rm cyl}$ also decreases dramatically from 0.53 to 0.30.
-So, the planet is migrating significantly with respect to the disk mass even while the inner part of the gas disk is moving inward during the burst.
-As a measure of the gas motion, $m_{\rm cyl}$ at 23 AU increases from 0.50 to 0.60 from $t =$ 3.89 to 8.47 ORP, corresponding to an average gas disk inflow rate of $\sim 10^{-5} M_{\odot}$ yr$^{-1}$.
+{\it Migration Relative to the Disk.} The disk surface density distribution does not vary much in the asymptotic phase,
+so radial motion of the planet in the $t = 10$ planet run also represents radial migration relative to the gas. During
+the burst \citep[e.g.,][]{boley2006}, the surface density of the disk changes dramatically. To verify that the planet is
+really migrating relative to the background gas disk in the $t = 0$ run, we have compared the evolution of the planet's
+radial position $\varpi_p(t)$ with the mass inside that radius, expressed in terms of $m_{\rm cyl}$, the fraction of the
+disk mass interior to the planet. Between $t = 3.52$ and $3.89$ ORP, when $\varpi_p$ increases modestly, $m_{\rm cyl}$
+increases from about 0.50 to 0.53; between 3.89 and 8.47 ORP, when $\varpi_p$ decreases dramatically, $m_{\rm cyl}$ also
+decreases dramatically from 0.53 to 0.30. So, the planet is migrating significantly with respect to the disk mass even
+while the inner part of the gas disk is moving inward during the burst. As a measure of the gas motion, $m_{\rm cyl}$ at
+23 AU increases from 0.50 to 0.60 from $t =$ 3.89 to 8.47 ORP, corresponding to an average gas disk inflow rate of $\sim
+10^{-5} M_{\odot}$ yr$^{-1}$.
 
@@
 
-Even in non-self-gravitating disks, the presence of discrete radial structure is known to cause instabilities \citep{li2000,li2001}.
-\citet{meschiari2008} demonstrated that a gap, presumably opened by an embedded planet, can cause global modes to become unstable in self-gravitating disks that are otherwise stable against the development of GIs.
-Here we are dealing with a planet that is not quite massive enough to open a gap, but its gravitational interaction with the disk also has a strong effect on the growth and onset of GIs.
+Even in non-self-gravitating disks, the presence of discrete radial structure is known to cause instabilities
+\citep{li2000,li2001}. \citet{meschiari2008} demonstrated that a gap, presumably opened by an embedded planet, can cause
+global modes to become unstable in self-gravitating disks that are otherwise stable against the development of GIs. Here
+we are dealing with a planet that is not quite massive enough to open a gap, but its gravitational interaction with the
+disk also has a strong effect on the growth and onset of GIs.
 
-Figure \ref{fig:a} suggests that the overall radial drift during the first 10 ORP of the $t = 0$ planet run is not too different from what one would get by extrapolating the migration due solely to the planet's wake during the first 2 or 3 ORP, where the average radial migration is about 4$\times10^{-3}$ AU yr$^{-1}$.
-Although there are no simple formulae that are fully applicable to our radiatively cooled disk, using eq.
-(70) from \citet{tanaka2002} with parameters from the simulated disk, we estimate 6$\times10^{-3}$ AU yr$^{-1}$ as an analytic expectation.
-This is a reasonable estimate for comparison purposes, because our resolution is probably insufficient to compute the corotation torque \citep{paardekooper2010}.
-While the interaction with a GI-active disk does not seem to result in a drastically different overall migration rate, the direction and magnitude of migration in both simulations is quite variable and seems to depend strongly on interactions of the planet with discrete modes in the disk.
-In $t = 10$ planet run, the overall radial migration rate is similar but is even more erratic in direction than in the $t = 0$ planet run.
-The important departures from laminar disk theory are threefold: the more erratic nature of migration, the importance of interactions with discrete low-order modes, and stalling of migration near the inner edge of the strongest GI activity.
-The migration is neither a random walk nor monotonic but has a chaotic character due to nonlinear interactions.
-For all these reasons, it will be important in future studies to explore the dependence of the planet's behavior on its mass and on its placement relative to the site of GI-eruptions (for $t = 0$ type runs) and relative to the CRs of low-order GI modes (for $t = 10$ type runs).
+Figure \ref{fig:a} suggests that the overall radial drift during the first 10 ORP of the $t = 0$ planet run is not too
+different from what one would get by extrapolating the migration due solely to the planet's wake during the first 2 or 3
+ORP, where the average radial migration is about 4$\times10^{-3}$ AU yr$^{-1}$. Although there are no simple formulae
+that are fully applicable to our radiatively cooled disk, using eq. (70) from \citet{tanaka2002} with parameters from
+the simulated disk, we estimate 6$\times10^{-3}$ AU yr$^{-1}$ as an analytic expectation. While the interaction with a
+GI-active disk does not seem to result in a drastically different overall migration rate, the direction and magnitude of
+migration in both simulations is quite variable and seems to depend strongly on interactions of the planet with discrete
+modes in the disk. In $t = 10$ planet run, the overall radial migration rate is similar but is even more erratic in
+direction than in the $t = 0$ planet run. The important departures from laminar disk theory are threefold: the more
+erratic nature of migration, the importance of interactions with discrete low-order modes, and stalling of migration
+near the inner edge of the strongest GI activity. The migration is neither a random walk nor monotonic but has a chaotic
+character due to nonlinear interactions. For all these reasons, it will be important in future studies to explore the
+dependence of the planet's behavior on its mass and on its placement relative to the site of GI-eruptions (for $t = 0$
+type runs) and relative to the CRs of low-order GI modes (for $t = 10$ type runs).
 
-In both simulations, the inward migration appears to halt near 17 AU.
-This may simply be a coincidence, because the overall average migration rates are similar, the starting positions are essentially the same, and the calculations have a similar duration.
-We plan to extend simulations like these to much longer times in the future.
-However, the final radial positions of both planets happen to be the location of the inner Lindblad resonance (ILR) for a strong m = 2 mode present in both runs.
-This ILR is associated with a surface density enhancement in the disk.
-Near this enhancement, the surface density gradient may be falling steeply enough to halt migration, a mechanism that can work in a laminar disk \citep{paardekooper2009}.
-Alternatively, at 17 AU, the level of GI activity is quite different interior to the planet ($Q \gtrsim 2$) and exterior to the planet ($Q \lesssim 2$), and this may affect the balance of the Lindblad torques exerted by the disk on the planet.
-Whether the migration really does stall and what the mechanism may be are subjects for further research.
+In both simulations, the inward migration appears to halt near 17 AU. This may simply be a coincidence, because the
+overall average migration rates are similar, the starting positions are essentially the same, and the calculations have
+a similar duration. We plan to extend simulations like these to much longer times in the future. However, the final
+radial positions of both planets happen to be the location of the inner Lindblad resonance (ILR) for a strong m = 2 mode
+present in both runs. This ILR is associated with a surface density enhancement in the disk. Near this enhancement, the
+surface density gradient may be falling steeply enough to halt migration, a mechanism that can work in a laminar disk
+\citep{paardekooper2009}.  Alternatively, at 17 AU, the level of GI activity is quite different interior to the planet
+($Q \gtrsim 2$) and exterior to the planet ($Q \lesssim 2$), and this may affect the balance of the Lindblad torques
+exerted by the disk on the planet. Whether the migration really does stall and what the mechanism may be are subjects
+for further research.
 
-{\it Summary.} The simulations in this Letter show that a planet of only 1$M_J$ placed near the $Q$-minimum in a 0.14$M_{\odot}$ disk can hasten the onset of gravitational instabilities and change the nature of the dominant modes during the initial burst.
-Although the magnitude of radial migration (a few$\times10^{-3}$ AU yr$^{-1}$) is not too different from that computed using the \citet{tanaka2002}  isothermal disk formula based on the Lindblad torques, it can fluctuate in sign on time scales on the order of the orbit period and can be strongly affected by gravitational interactions of the planet with discrete GI-driven spiral modes, leading to phases of rapid radial motion.
-Planet orbit eccentricities fluctuate between about 0.02 and 0.10.
-The simulations also suggest that planet migration may stall for long periods near the ILRs of dominant global GI-modes.
+Due to the global nature of the GIs and their strong interaction with the embedded planet, it is important that the gas
+phase physics are as accurate as possible and detailed analysis of the interaction between the planet and global modes
+is required. Of particular importance is the treatment of cooling in the disk. Although these simulations show some
+similarities to \citet{baruteau2011}, their simulations employ an idealized cooling rate governed by the local dynamical
+timescale while the simulations in this work use the more realistic radiative physics of \citet{boley2007b}. {\bf
+  citations on global vs. local needed} Another primary difference between this work and that of \citet{baruteau2011} is
+the simulations presented here are three dimensional. Many studies have shown that three dimensional simulations are
+necessary to accurately measure the effects of GIs {\bf need some more citations here}. Further in depth analysis of the
+GIs and the interaction between the GIs and the planet will be presented in a future work.
+
+{\bf Say something about possible physical scenarios for t=0 run}
+
+{\it Summary.} The simulations in this Letter show that a planet of only 1$M_J$ placed near the $Q$-minimum in a
+0.14$M_{\odot}$ disk can hasten the onset of gravitational instabilities and change the nature of the dominant modes
+during the initial burst. Although the magnitude of radial migration (a few$\times10^{-3}$ AU yr$^{-1}$) is not too
+different from that computed using the \citet{tanaka2002} isothermal disk formula based on the Lindblad torques, it can
+fluctuate in sign on time scales on the order of the orbit period and can be strongly affected by gravitational
+interactions of the planet with discrete GI-driven spiral modes, leading to phases of rapid radial motion. Planet orbit
+eccentricities fluctuate between about 0.02 and 0.10. The simulations also suggest that planet migration may stall for
+long periods near the ILRs of dominant global GI-modes.
 
@@
 
-The authors would like to thank F.C.
-Adams and J.C.B.
-Papaloizou for useful encouragement in this project.
-The research was conducted with the support of NASA grants from the Origins of Solar Systems  Program (NNG05GN11G and NNX08AK36G).
-S.M.
-was supported during portions of this research by an NESSP Graduate Fellowship and by Indiana Space Grant Consortium Graduate Fellowships; A.C.B.
-is currently a Sagan Postdoctoral Fellow at U.
-Florida.
-This material is based upon work supported by the National Science Foundation under Grants No.
-ACI-0338618l, OCI-0451237, OCI-0535258, OCI-0504075, and CNS-0521433, and through TeraGrid resources provided by Indiana University, PSC, SDSC, and NCSA.
-TeraGrid systems are hosted by Indiana University, LONI, NCAR, NCSA, NICS, ORNL, PSC, Purdue University, SDSC, TACC, and UC/ANL.
-Any opinions, findings and conclusions or recommendations expressed in this material are those of the authors and do not necessarily reflect the views of the National Science Foundation.
+The authors would like to thank F.C. Adams and J.C.B. Papaloizou for useful encouragement in this project. The research
+was conducted with the support of NASA grants from the Origins of Solar Systems Program (NNG05GN11G and
+NNX08AK36G). S.M. was supported during portions of this research by an NESSP Graduate Fellowship and by Indiana Space
+Grant Consortium Graduate Fellowships; A.C.B. is currently a Sagan Postdoctoral Fellow at U. Florida. This material is
+based upon work supported by the National Science Foundation under Grants No. ACI-0338618l, OCI-0451237, OCI-0535258,
+OCI-0504075, and CNS-0521433, and through TeraGrid resources provided by Indiana University, PSC, SDSC, and
+NCSA. TeraGrid systems are hosted by Indiana University, LONI, NCAR, NCSA, NICS, ORNL, PSC, Purdue University, SDSC,
+TACC, and UC/ANL. Any opinions, findings and conclusions or recommendations expressed in this material are those of the
+authors and do not necessarily reflect the views of the National Science Foundation.
 
@@
 
-%***DICK: Scott, remember that figures need some improvements.
+\begin{figure}[t]
+\center
+\includegraphics[width=12cm]{Figures/hr_migration.eps}
+\caption{(Color figure available in the online version) Plots of the disk aspect ratio $h/r$ as determined by a decrease
+  from midplane density by a factor of $e^2$ for the fiducial and $t=0$ runs (black) at $\approx 10$ ORP and the fiducial and
+  $t=10$ runs (red) at $\approx 20$ ORP.}
+\label{fig:hr}
+\end{figure}
 
@@
 \center
-\includegraphics[width=11cm]{Figures/amt.eps}
-\caption{(Color figure available in the online version) Global amplitudes of nonaxisymmetric density perturbations $A_m$ as a function of time for individual cos($m\phi$) perturbations.
-The formula for computing $A_m$ is equation (15) of \citep{boley2006}.
-Here the integrals are done only over the disk outside 15 AU to suppress contributions from edge effects in the inner disk.
-The top and bottom panels correspond to the fiducial run (no planet) and the $t = 0$ planet run, respectively.}
+\includegraphics[width=12cm]{Figures/amt.eps}
+\caption{(Color figure available in the online version) Global amplitudes of nonaxisymmetric density perturbations $A_m$
+  as a function of time for individual cos($m\phi$) perturbations. The formula for computing $A_m$ is equation (15) of
+  \citep{boley2006}. Here the integrals are done only over the disk outside 15 AU to suppress contributions from edge
+  effects in the inner disk. The top and bottom panels correspond to the fiducial run (no planet) and the $t = 0$ planet
+  run, respectively.}
 \label{fig:Am}
@@
 \includegraphics[width=12cm]{Figures/planeta.eps}
-\caption{Plot of the planet's radial position as a function of time $\varpi_p(t)$ for the two planet simulations.
-The dashed-diamond and solid-star lines (colored red and black in the online version) indicate the curves ($\varpi(t)$), points ($a$), and horizontal scales ($t$) that correspond to the $t = 0$ and $t = 10$ planet runs, respectively.
-The symbols are the approximate semi-major axes $a$ computed for each 0 to 2$\phi$ change in azimuth of the planet as explained in the text.}
+\caption{Plot of the planet's radial position as a function of time $\varpi_p(t)$ for the two planet simulations. The
+  dashed-diamond and solid-star lines (colored red and black in the online version) indicate the curves ($\varpi(t)$),
+  points ($a$), and horizontal scales ($t$) that correspond to the $t = 0$ and $t = 10$ planet runs, respectively. The
+  symbols are the approximate semi-major axes $a$ computed for each 0 to 2$\phi$ change in azimuth of the planet as
+  explained in the text.}
 \label{fig:a}
@@
 \includegraphics[width=12cm]{Figures/planete.eps}
-\caption{Plot of the eccentricities $e$, computed for each 0 to 2$\phi$ change in azimuth of the planet as explained in the text, as a function of time.
-The diamonds and stars (colored red and black in the online version) indicate the points ($e$) and horizontal scales ($t$) that correspond to the $t = 0$ and $t = 10$ planet runs, respectively.}
+\caption{Plot of the eccentricities $e$, computed for each 0 to 2$\phi$ change in azimuth of the planet as explained in
+  the text, as a function of time. The diamonds and stars (colored red and black in the online version) indicate the
+  points ($e$) and horizontal scales ($t$) that correspond to the $t = 0$ and $t = 10$ planet runs, respectively.}
 \label{fig:e}

### 2011-06-24 Text/migration.tex
Manuscript prepared for resubmission to ApJL
Added h/r plot and text to address several of the referee's concerns.
	modified:   ../README.md
	modified:   migration.tex
	modified:   migration_apjlsubmission.tex

@@
 to 21 ORP without a planet. In the second simulation, which we call the {\sl t = 0 planet run}, a one Jupiter-mass
-($M_J$) planet with a roughly circular midplane orbit is included in the initial $t = 0$ ORP equilibrium disk at 25
-AU. In the third simulation, which we call the {\sl t = 10 planet run}, a 1 $M_J$ planet with a similar orbit is
-inserted into the fiducial run at $t = 10.5$ ORP. These three simulations are part of a larger suite, to be described
-elsewhere, in which the planet mass is varied. We estimate semi-major axes $a$ and eccentricities $e$ from the planet's
-$\varpi(t)$ by using the maximum and minimum radii, $\varpi_{\rm max}$ and $\varpi_{\rm min}$, over each complete orbit
-of the planet, where one orbit is $\Delta\phi=2\pi$. Specifically, $a = (\varpi_{\rm max}+\varpi_{\rm min})/2$ and $e =
-(\varpi_{\rm max}-\varpi_{\rm min})/(\varpi_{\rm max}+\varpi_{\rm min})$. These parameters are intended to highlight the
-strong orbit-to-orbit variations in the motion of the planet.  A more refined estimate of the orbital elements is
-unnecessary because the potential of the disk, even if were axisymmetric, will cause departures from Keplerian dynamics.
+($M_J$) planet with a roughly circular midplane orbit at 25 AU is included in the initial $t = 0$ ORP equilibrium
+disk. This run investigates the effects of an existing planet on the growth of GIs, which could occur if disk the disk
+undergoes episodic or localized instabilities. In the third simulation, which we call the {\sl t = 10 planet run}, a 1
+$M_J$ planet with a similar orbit is inserted into the fiducial run at $t = 10.5$ ORP. We estimate semi-major axes $a$
+and eccentricities $e$ from the planet's $\varpi(t)$ by using the maximum and minimum radii, $\varpi_{\rm max}$ and
+$\varpi_{\rm min}$, over each complete orbit of the planet, where one orbit is $\Delta\phi=2\pi$. Specifically, $a =
+(\varpi_{\rm max}+\varpi_{\rm min})/2$ and $e = (\varpi_{\rm max}-\varpi_{\rm min})/(\varpi_{\rm max}+\varpi_{\rm
+  min})$. These parameters are intended to highlight the strong orbit-to-orbit variations in the motion of the planet.
+A more refined estimate of the orbital elements is unnecessary because the potential of the disk, even if were
+axisymmetric, will cause departures from Keplerian dynamics.
 
@@
 that are fully applicable to our radiatively cooled disk, using eq. (70) from \citet{tanaka2002} with parameters from
-the simulated disk, we estimate 6$\times10^{-3}$ AU yr$^{-1}$ as an analytic expectation. While the interaction with a
-GI-active disk does not seem to result in a drastically different overall migration rate, the direction and magnitude of
-migration in both simulations is quite variable and seems to depend strongly on interactions of the planet with discrete
-modes in the disk. In $t = 10$ planet run, the overall radial migration rate is similar but is even more erratic in
-direction than in the $t = 0$ planet run. The important departures from laminar disk theory are threefold: the more
-erratic nature of migration, the importance of interactions with discrete low-order modes, and stalling of migration
-near the inner edge of the strongest GI activity. The migration is neither a random walk nor monotonic but has a chaotic
-character due to nonlinear interactions. For all these reasons, it will be important in future studies to explore the
-dependence of the planet's behavior on its mass and on its placement relative to the site of GI-eruptions (for $t = 0$
-type runs) and relative to the CRs of low-order GI modes (for $t = 10$ type runs).
+the simulated disk, we estimate 6$\times10^{-3}$ AU yr$^{-1}$ as an analytic expectation (see figure \ref{fig:hr} for
+disk aspect ratios). While the interaction with a GI-active disk does not seem to result in a drastically different
+overall migration rate, the direction and magnitude of migration in both simulations is quite variable and seems to
+depend strongly on interactions of the planet with discrete modes in the disk. In $t = 10$ planet run, the overall
+radial migration rate is similar but is even more erratic in direction than in the $t = 0$ planet run. The important
+departures from laminar disk theory are threefold: the more erratic nature of migration, the importance of interactions
+with discrete low-order modes, and stalling of migration near the inner edge of the strongest GI activity. The migration
+is neither a random walk nor monotonic but has a chaotic character due to nonlinear interactions. For all these reasons,
+it will be important in future studies to explore the dependence of the planet's behavior on its mass and on its
+placement relative to the site of GI-eruptions (for $t = 0$ type runs) and relative to the CRs of low-order GI modes
+(for $t = 10$ type runs).
 
@@
 similarities to \citet{baruteau2011}, their simulations employ an idealized cooling rate governed by the local dynamical
-timescale while the simulations in this work use the more realistic radiative physics of \citet{boley2007b}. {\bf
-  citations on global vs. local needed} Another primary difference between this work and that of \citet{baruteau2011} is
-the simulations presented here are three dimensional. Many studies have shown that three dimensional simulations are
-necessary to accurately measure the effects of GIs {\bf need some more citations here}. Further in depth analysis of the
-GIs and the interaction between the GIs and the planet will be presented in a future work.
-
-{\bf Say something about possible physical scenarios for t=0 run}
+timescale while the simulations in this work use the more realistic radiative physics of
+\citet{boley2007b}. Three-dimensional simulations with radiative physics are necessary for understanding the evolution
+of spiral structure in vertically stratified disks because shocks and heating along spiral arms can change the disk's
+vertical structure \citep[e.g.][]{boley2006a}, which affects the migration rate of planets.
 
@@
 was conducted with the support of NASA grants from the Origins of Solar Systems Program (NNG05GN11G and
-NNX08AK36G). S.M. was supported during portions of this research by an NESSP Graduate Fellowship and by Indiana Space
-Grant Consortium Graduate Fellowships; A.C.B. is currently a Sagan Postdoctoral Fellow at U. Florida. This material is
-based upon work supported by the National Science Foundation under Grants No. ACI-0338618l, OCI-0451237, OCI-0535258,
-OCI-0504075, and CNS-0521433, and through TeraGrid resources provided by Indiana University, PSC, SDSC, and
-NCSA. TeraGrid systems are hosted by Indiana University, LONI, NCAR, NCSA, NICS, ORNL, PSC, Purdue University, SDSC,
-TACC, and UC/ANL. Any opinions, findings and conclusions or recommendations expressed in this material are those of the
-authors and do not necessarily reflect the views of the National Science Foundation.
+NNX08AK36G). A.C.B. is currently a Sagan Postdoctoral Fellow at U. Florida. This material is based upon work supported
+by the National Science Foundation under Grants No. ACI-0338618l, OCI-0451237, OCI-0535258, OCI-0504075, and
+CNS-0521433, and through TeraGrid resources provided by Indiana University, PSC, SDSC, and NCSA.
 
@@
 \newpage
-
-\begin{figure}[t]
-\center
-\includegraphics[width=12cm]{Figures/hr_migration.eps}
-\caption{(Color figure available in the online version) Plots of the disk aspect ratio $h/r$ as determined by a decrease
-  from midplane density by a factor of $e^2$ for the fiducial and $t=0$ runs (black) at $\approx 10$ ORP and the fiducial and
-  $t=10$ runs (red) at $\approx 20$ ORP.}
-\label{fig:hr}
-\end{figure}
 
@@
 
+\begin{figure}[t]
+\center
+\includegraphics[width=12cm]{Figures/hr_migration.eps}
+\caption{(Color figure available in the online version) Plots of the disk aspect ratio $h/r$ as determined by a decrease
+  from midplane density by a factor of $e^2$ for the fiducial and $t=0$ runs (black) at $10.5$ ORP and the fiducial and
+  $t=10$ runs (light gray) at $20.5$ ORP.}
+\label{fig:hr}
+\end{figure}
+
 %\begin{figure}[t]

### 2011-06-24 Text/migration_apjlsubmission.tex
Final, Final, Final edits. Added Dick's comment and submitted 06/07/2011
	modified:   Text/migration.tex
	modified:   Text/migration_apjlsubmission.tex

git-svn-id: http://jaga.astro.indiana.edu/paperrepo/ApJL/Migration2010@87 3b7f7548-5c66-0410-83ab-3997167635fc
Manuscript prepared for resubmission to ApJL
Added h/r plot and text to address several of the referee's concerns.
	modified:   ../README.md
	modified:   migration.tex
	modified:   migration_apjlsubmission.tex

@@
 
-The model disk, based on an equlibrium model from \citet{pickett2003}, orbits a 1 $M_{\odot}$ star and has a mass $M_d = 0.14 M_{\odot}$, inner and outer radii at 5 and 40 AU, and an initial surface density distribution $\Sigma \sim \varpi^{-1/2}$.
-The time unit of one ORP (= outer rotation period) is defined as the rotation period of the initial disk at $\varpi \approx$ 32 AU, or about 180 yr.
-The disk is initially located between radial grid zones 30 and 240 and is close to isentropic, which results in a Toomre-$Q$ distribution with a marginally unstable \citep[see ][]{durisen2007} minimum value of 1.38 at radial grid zone 161 (26.7 AU).
-The computational grid extends radially to 512 zones to accommodate expansion of the outer disk once GIs become nonlinear.
-An outflow boundary condition is enforced at the upper vertical grid boundary, the outer radial grid boundary, and an inner radial boundary at 2 AU.
-To seed nonaxisymmetry, the density distribution is given an initial 0.01 \% random cell-to-cell perturbation.
+The model disk, based on an equilibrium model from \citet{pickett2003}, orbits a 1 $M_{\odot}$ star and has a mass $M_d =
+0.14 M_{\odot}$, inner and outer radii at 5 and 40 AU, and an initial surface density distribution $\Sigma \sim
+\varpi^{-1/2}$. The time unit of one ORP (= outer rotation period) is defined as the rotation period of the initial disk
+at $\varpi \approx$ 32 AU, or about 180 yr. The disk is initially located between radial grid zones 30 and 240 and is
+close to isentropic, which results in a Toomre-$Q$ distribution with a marginally unstable \citep[see ][]{durisen2007}
+minimum value of 1.38 at radial grid zone 161 (26.7 AU). The computational grid extends radially to 512 zones to
+accommodate expansion of the outer disk once GIs become nonlinear. An outflow boundary condition is enforced at the
+upper vertical grid boundary, the outer radial grid boundary, and an inner radial boundary at 2 AU. To seed
+nonaxisymmetry, the density distribution is given an initial 0.01 \% random cell-to-cell perturbation.
 
-This Letter presents three simulations.
-The first, which we call the {\sl fiducial run},  is simulated from time $t = 0$ to 21 ORP without a planet.
-In the second simulation, which we call the {\sl t = 0 planet run}, a one Jupiter-mass ($M_J$) planet with a roughly circular midplane orbit is included in the initial $t = 0$ ORP equilibrium disk at 25 AU.
-In the third simulation, which we call the {\sl t = 10 planet run}, a 1 $M_J$ planet with a similar orbit is inserted into the fiducial run at $t = 10.5$ ORP.
-These three simulations are part of a larger suite, to be described elsewhere, in which the planet mass is varied.
-We estimate semi-major axes $a$ and eccentricities $e$ from the planet's $\varpi(t)$ by using the maximum and minimum radii, $\varpi_{\rm max}$ and $\varpi_{\rm min}$, over each complete orbit of the planet, where one orbit is $\Delta\phi=2\pi$.
-Specifically, $a = (\varpi_{\rm max}+\varpi_{\rm min})/2$ and $e = (\varpi_{\rm max}-\varpi_{\rm min})/(\varpi_{\rm max}+\varpi_{\rm min})$.
-These parameters are intended to highlight the strong orbit-to-orbit variations in the motion of the planet.
-A more refined estimate of the orbital elements is unnecessary because the potential of the disk, even if were axisymmetric, will cause departures from Keplerian dynamics.
+This Letter presents three simulations. The first, which we call the {\sl fiducial run}, is simulated from time $t = 0$
+to 21 ORP without a planet. In the second simulation, which we call the {\sl t = 0 planet run}, a one Jupiter-mass
+($M_J$) planet with a roughly circular midplane orbit at 25 AU is included in the initial $t = 0$ ORP equilibrium
+disk. This run investigates the effects of an existing planet on the growth of GIs, which could occur if disk the disk
+undergoes episodic or localized instabilities. In the third simulation, which we call the {\sl t = 10 planet run}, a 1
+$M_J$ planet with a similar orbit is inserted into the fiducial run at $t = 10.5$ ORP. We estimate semi-major axes $a$
+and eccentricities $e$ from the planet's $\varpi(t)$ by using the maximum and minimum radii, $\varpi_{\rm max}$ and
+$\varpi_{\rm min}$, over each complete orbit of the planet, where one orbit is $\Delta\phi=2\pi$. Specifically, $a =
+(\varpi_{\rm max}+\varpi_{\rm min})/2$ and $e = (\varpi_{\rm max}-\varpi_{\rm min})/(\varpi_{\rm max}+\varpi_{\rm
+  min})$. These parameters are intended to highlight the strong orbit-to-orbit variations in the motion of the planet.
+A more refined estimate of the orbital elements is unnecessary because the potential of the disk, even if were
+axisymmetric, will cause departures from Keplerian dynamics.
 
@@
 
-Even in non-self-gravitating disks, the presence of discrete radial structure is known to cause instabilities \citep{li2000,li2001}.
-\citet{meschiari2008} demonstrated that a gap, presumably opened by an embedded planet, can cause global modes to become unstable in self-gravitating disks that are otherwise stable against the development of GIs.
-Here we are dealing with a planet that is not quite massive enough to open a gap, but its gravitational interaction with the disk also has a strong effect on the growth and onset of GIs.
+Even in non-self-gravitating disks, the presence of discrete radial structure is known to cause instabilities
+\citep{li2000,li2001}. \citet{meschiari2008} demonstrated that a gap, presumably opened by an embedded planet, can cause
+global modes to become unstable in self-gravitating disks that are otherwise stable against the development of GIs. Here
+we are dealing with a planet that is not quite massive enough to open a gap, but its gravitational interaction with the
+disk also has a strong effect on the growth and onset of GIs.
 
-Figure \ref{fig:a} suggests that the overall radial drift during the first 10 ORP of the $t = 0$ planet run is not too different from what one would get by extrapolating the migration due solely to the planet's wake during the first 2 or 3 ORP, where the average radial migration is about 4$\times10^{-3}$ AU yr$^{-1}$.
-Although there are no simple formulae that are fully applicable to our radiatively cooled disk, using eq.
-(70) from \citet{tanaka2002} with parameters from the simulated disk, we estimate 6$\times10^{-3}$ AU yr$^{-1}$ as an analytic expectation.
-This is a reasonable estimate for comparison purposes, because our resolution is probably insufficient to compute the corotation torque \citep{paardekooper2010}.
-While the interaction with a GI-active disk does not seem to result in a drastically different overall migration rate, the direction and magnitude of migration in both simulations is quite variable and seems to depend strongly on interactions of the planet with discrete modes in the disk.
-In $t = 10$ planet run, the overall radial migration rate is similar but is even more erratic in direction than in the $t = 0$ planet run.
-The important departures from laminar disk theory are threefold: the more erratic nature of migration, the importance of interactions with discrete low-order modes, and stalling of migration near the inner edge of the strongest GI activity.
-The migration is neither a random walk nor monotonic but has a chaotic character due to nonlinear interactions.
-For all these reasons, it will be important in future studies to explore the dependence of the planet's behavior on its mass and on its placement relative to the site of GI-eruptions (for $t = 0$ type runs) and relative to the CRs of low-order GI modes (for $t = 10$ type runs).
+Figure \ref{fig:a} suggests that the overall radial drift during the first 10 ORP of the $t = 0$ planet run is not too
+different from what one would get by extrapolating the migration due solely to the planet's wake during the first 2 or 3
+ORP, where the average radial migration is about 4$\times10^{-3}$ AU yr$^{-1}$. Although there are no simple formulae
+that are fully applicable to our radiatively cooled disk, using eq. (70) from \citet{tanaka2002} with parameters from
+the simulated disk, we estimate 6$\times10^{-3}$ AU yr$^{-1}$ as an analytic expectation (see figure \ref{fig:hr} for
+disk aspect ratios). While the interaction with a GI-active disk does not seem to result in a drastically different
+overall migration rate, the direction and magnitude of migration in both simulations is quite variable and seems to
+depend strongly on interactions of the planet with discrete modes in the disk. In $t = 10$ planet run, the overall
+radial migration rate is similar but is even more erratic in direction than in the $t = 0$ planet run. The important
+departures from laminar disk theory are threefold: the more erratic nature of migration, the importance of interactions
+with discrete low-order modes, and stalling of migration near the inner edge of the strongest GI activity. The migration
+is neither a random walk nor monotonic but has a chaotic character due to nonlinear interactions. For all these reasons,
+it will be important in future studies to explore the dependence of the planet's behavior on its mass and on its
+placement relative to the site of GI-eruptions (for $t = 0$ type runs) and relative to the CRs of low-order GI modes
+(for $t = 10$ type runs).
 
-In both simulations, the inward migration appears to halt near 17 AU.
-This may simply be a coincidence, because the overall average migration rates are similar, the starting positions are essentially the same, and the calculations have a similar duration.
-We plan to extend simulations like these to much longer times in the future.
-However, the final radial positions of both planets happen to be the location of the inner Lindblad resonance (ILR) for a strong m = 2 mode present in both runs.
-This ILR is associated with a surface density enhancement in the disk.
-Near this enhancement, the surface density gradient may be falling steeply enough to halt migration, a mechanism that can work in a laminar disk \citep{paardekooper2009}.
-Alternatively, at 17 AU, the level of GI activity is quite different interior to the planet ($Q \gtrsim 2$) and exterior to the planet ($Q \lesssim 2$), and this may affect the balance of the Lindblad torques exerted by the disk on the planet.
-Whether the migration really does stall and what the mechanism may be are subjects for further research.
+In both simulations, the inward migration appears to halt near 17 AU. This may simply be a coincidence, because the
+overall average migration rates are similar, the starting positions are essentially the same, and the calculations have
+a similar duration. We plan to extend simulations like these to much longer times in the future. However, the final
+radial positions of both planets happen to be the location of the inner Lindblad resonance (ILR) for a strong m = 2 mode
+present in both runs. This ILR is associated with a surface density enhancement in the disk. Near this enhancement, the
+surface density gradient may be falling steeply enough to halt migration, a mechanism that can work in a laminar disk
+\citep{paardekooper2009}.  Alternatively, at 17 AU, the level of GI activity is quite different interior to the planet
+($Q \gtrsim 2$) and exterior to the planet ($Q \lesssim 2$), and this may affect the balance of the Lindblad torques
+exerted by the disk on the planet. Whether the migration really does stall and what the mechanism may be are subjects
+for further research.
 
-{\it Summary.} The simulations in this Letter show that a planet of only 1$M_J$ placed near the $Q$-minimum in a 0.14$M_{\odot}$ disk can hasten the onset of gravitational instabilities and change the nature of the dominant modes during the initial burst.
-Although the magnitude of radial migration (a few$\times10^{-3}$ AU yr$^{-1}$) is not too different from that computed using the \citet{tanaka2002}  isothermal disk formula based on the Lindblad torques, it can fluctuate in sign on time scales on the order of the orbit period and can be strongly affected by gravitational interactions of the planet with discrete GI-driven spiral modes, leading to phases of rapid radial motion.
-Planet orbit eccentricities fluctuate between about 0.02 and 0.10.
-The simulations also suggest that planet migration may stall for long periods near the ILRs of dominant global GI-modes.
+Due to the global nature of the GIs and their strong interaction with the embedded planet, it is important that the gas
+phase physics are as accurate as possible and detailed analysis of the interaction between the planet and global modes
+is required. Of particular importance is the treatment of cooling in the disk. Although these simulations show some
+similarities to \citet{baruteau2011}, their simulations employ an idealized cooling rate governed by the local dynamical
+timescale while the simulations in this work use the more realistic radiative physics of
+\citet{boley2007b}. Three-dimensional simulations with radiative physics are necessary for understanding the evolution
+of spiral structure in vertically stratified disks because shocks and heating along spiral arms can change the disk's
+vertical structure \citep[e.g.][]{boley2006a}, which affects the migration rate of planets.
+
+{\it Summary.} The simulations in this Letter show that a planet of only 1$M_J$ placed near the $Q$-minimum in a
+0.14$M_{\odot}$ disk can hasten the onset of gravitational instabilities and change the nature of the dominant modes
+during the initial burst. Although the magnitude of radial migration (a few$\times10^{-3}$ AU yr$^{-1}$) is not too
+different from that computed using the \citet{tanaka2002} isothermal disk formula based on the Lindblad torques, it can
+fluctuate in sign on time scales on the order of the orbit period and can be strongly affected by gravitational
+interactions of the planet with discrete GI-driven spiral modes, leading to phases of rapid radial motion. Planet orbit
+eccentricities fluctuate between about 0.02 and 0.10. The simulations also suggest that planet migration may stall for
+long periods near the ILRs of dominant global GI-modes.
 
@@
 
-The authors would like to thank F.C.
-Adams and J.C.B.
-Papaloizou for useful encouragement in this project.
-The research was conducted with the support of NASA grants from the Origins of Solar Systems  Program (NNG05GN11G and NNX08AK36G).
-S.M.
-was supported during portions of this research by an NESSP Graduate Fellowship and by Indiana Space Grant Consortium Graduate Fellowships; A.C.B.
-is currently a Sagan Postdoctoral Fellow at U.
-Florida.
-This material is based upon work supported by the National Science Foundation under Grants No.
-ACI-0338618l, OCI-0451237, OCI-0535258, OCI-0504075, and CNS-0521433, and through TeraGrid resources provided by Indiana University, PSC, SDSC, and NCSA.
-TeraGrid systems are hosted by Indiana University, LONI, NCAR, NCSA, NICS, ORNL, PSC, Purdue University, SDSC, TACC, and UC/ANL.
-Any opinions, findings and conclusions or recommendations expressed in this material are those of the authors and do not necessarily reflect the views of the National Science Foundation.
+The authors would like to thank F.C. Adams and J.C.B. Papaloizou for useful encouragement in this project. The research
+was conducted with the support of NASA grants from the Origins of Solar Systems Program (NNG05GN11G and
+NNX08AK36G). A.C.B. is currently a Sagan Postdoctoral Fellow at U. Florida. This material is based upon work supported
+by the National Science Foundation under Grants No. ACI-0338618l, OCI-0451237, OCI-0535258, OCI-0504075, and
+CNS-0521433, and through TeraGrid resources provided by Indiana University, PSC, SDSC, and NCSA.
 
@@
 \center
-\includegraphics[width=11cm]{amt.eps}
-\caption{(Color figure available in the online version) Global amplitudes of nonaxisymmetric density perturbations $A_m$ as a function of time for individual cos($m\phi$) perturbations.
-The formula for computing $A_m$ is equation (15) of \citep{boley2006}.
-Here the integrals are done only over the disk outside 15 AU to suppress contributions from edge effects in the inner disk.
-The top and bottom panels correspond to the fiducial run (no planet) and the $t = 0$ planet run, respectively.}
+\includegraphics[width=12cm]{Figures/amt.eps}
+\caption{(Color figure available in the online version) Global amplitudes of nonaxisymmetric density perturbations $A_m$
+  as a function of time for individual cos($m\phi$) perturbations. The formula for computing $A_m$ is equation (15) of
+  \citep{boley2006}. Here the integrals are done only over the disk outside 15 AU to suppress contributions from edge
+  effects in the inner disk. The top and bottom panels correspond to the fiducial run (no planet) and the $t = 0$ planet
+  run, respectively.}
 \label{fig:Am}
@@
 \center
-\includegraphics[width=12cm]{planeta.eps}
-\caption{Plot of the planet's radial position as a function of time $\varpi_p(t)$ for the two planet simulations.
-The dashed-diamond and solid-star lines (colored red and black in the online version) indicate the curves ($\varpi(t)$), points ($a$), and horizontal scales ($t$) that correspond to the $t = 0$ and $t = 10$ planet runs, respectively.
-The symbols are the approximate semi-major axes $a$ computed for each 0 to 2$\phi$ change in azimuth of the planet as explained in the text.}
+\includegraphics[width=12cm]{Figures/planeta.eps}
+\caption{Plot of the planet's radial position as a function of time $\varpi_p(t)$ for the two planet simulations. The
+  dashed-diamond and solid-star lines (colored red and black in the online version) indicate the curves ($\varpi(t)$),
+  points ($a$), and horizontal scales ($t$) that correspond to the $t = 0$ and $t = 10$ planet runs, respectively. The
+  symbols are the approximate semi-major axes $a$ computed for each 0 to 2$\phi$ change in azimuth of the planet as
+  explained in the text.}
 \label{fig:a}
@@
 \center
-\includegraphics[width=12cm]{planete.eps}
-\caption{Plot of the eccentricities $e$, computed for each 0 to 2$\phi$ change in azimuth of the planet as explained in the text, as a function of time.
-The diamonds and stars (colored red and black in the online version) indicate the points ($e$) and horizontal scales ($t$) that correspond to the $t = 0$ and $t = 10$ planet runs, respectively.}
+\includegraphics[width=12cm]{Figures/planete.eps}
+\caption{Plot of the eccentricities $e$, computed for each 0 to 2$\phi$ change in azimuth of the planet as explained in
+  the text, as a function of time. The diamonds and stars (colored red and black in the online version) indicate the
+  points ($e$) and horizontal scales ($t$) that correspond to the $t = 0$ and $t = 10$ planet runs, respectively.}
 \label{fig:e}
+\end{figure}
+
+\begin{figure}[t]
+\center
+\includegraphics[width=12cm]{Figures/hr_migration.eps}
+\caption{(Color figure available in the online version) Plots of the disk aspect ratio $h/r$ as determined by a decrease
+  from midplane density by a factor of $e^2$ for the fiducial and $t=0$ runs (black) at $10.5$ ORP and the fiducial and
+  $t=10$ runs (light gray) at $20.5$ ORP.}
+\label{fig:hr}
 \end{figure}

### 2011-06-26 Text/Response_to_referee1.txt
Added bw hr figure, changed README to reflect submission date and added response to ref.
	modified:   README.md
	new file:   Text/Figures/hr_migration_bw.eps
	new file:   Text/Response_to_referee1.txt

This document contains the authors responses to the initial referee's report for ApJL submission LET26384. Several
changes have been made to the manuscript. The most significant changes are an expanded discussion section and an
additional figure, figure 4, depicting the aspect ratio of the disks at different times. Detailed responses to the
referee's comments follow below.

1. "Even though there are some interesting results in the paper, most of the major results on migration in
gravitationally unstable disks were already reported in Baruteau et al. (2011). There is therefore no need for fast
publication as a Letter, especially since there appears to be very little explanation of the results in the current
manuscript. I would encourage the authors to resubmit a more detailed version of this work to the main journal, in
particular focusing on the differences of their approach compared to that of Baruteau et al. (2011)."

We believe it is important to note that the paper referenced here by the referee had been accepted to the MNRAS only one
week before our submission. It appears that the article is currently "in press" and is only publicly available on
astro-ph. As such we feel it should not block the publication of this article to the ApJL. In addition, as outlined in
our extended comments in the discussion section, Baruteau et al. use very different physics in their protoplanetary disk
simulations. The simulations presented in our work are fully three dimensional and use realistic radiative physics,
while the Baruteau et al. simulations are two dimensional and use an idealized cooling prescription. It has been shown
by many authors that both 3D and realistic cooling are required to accurately characterize and measure GIs.

2. "The critical dimensionless parameter governing the linearity of planet-disk interactions is q/h^3, where q is the
mass ratio of the planet - central star, and h is the aspect ratio of the disk. A plot of h(r) would therefore be
helpful."

Disk aspect ratios for the end of both planet runs and the corresponding time in the fiducial run have been provided in
figure 4.

3. "Gap formation is governed by above parameter together with viscosity (Crida et al. 2006). In the current set-up,
gravitational and Reynolds stresses probably strongly depend on radius. A radial plot of the stresses would give some
insight in gap formation at various locations in the disk."

Although we would like to include such figures and relevant discussion, and will in our future more detailed article,
the constraints of the Letter format prevent us from doing so.

4. "If the planet does not open up a gap, it is subject to Type I migration. The migration rate can then be very
sensitive to background temperature and density profiles (Paardekooper et al. 2010). Plots of azimuthally averaged
surface density and temperature profiles would help to determine whether one would expect inward or outward Type I
migration."

In fact, one of the main points of the article is that distinctions such as type II, type II, and type III migration,
which depend on the presence and character of a gap in a laminar disk, need not apply in a GI active disk. Although the
overall migration rate is similar to that predicted by type I migration theory, the same mechanism is not acting here.

5. "The authors state that they do not resolve the width w of the horseshoe region of the planet. Again, the critical
parameter is q/h^3. For h=0.1, w should already be larger than the scale height. Even if h=0.2, w should be close to
h/2. Not resolving w therefore implies not resolving h, which would be quite bad. Are the authors sure they can not
resolve w?"

We have removed the statement which led the referee to this conclusion. What we intended to convey was the fact that we
lack sufficient resolution to accurately measure corotation torques, not that we could not resolve the horseshoe region.

6. "In the introduction, please note that Baruteau & Masset (2008) were actually the first to include self-gravity in
migration simulations (but in laminar disks only). It is the gravitationally unstable nature of the disk that is new in
the present study and in Baruteau et al. (2011)."

As this article deals exclusively with GI active disks, and we did not indicate that another article was the first to
include self-gravity we declined to cite Baruteau & Masset (2008).

7. "I find the 't=0 planet run' difficult to understand. The authors seem to claim that the presence of the planet
affects the development of GI. How can they justify these initial conditions? If the planet was formed by GI, the disk
should already be turbulent. If it formed in a laminar disk (or in a previous GI burst), it should already have migrated
inward."

We have include the following text to illuminate some of the possible scenarios: "This run investigates the effects of
an existing planet on the growth of GIs, which could occur if disk the disk undergoes episodic or localized
instabilities."

8. "I am not sure what the role of the ILR of the dominant spiral mode is in stopping the migration of the
planet. Intuitively, one might expect an ILR to push the planet inward. What does the torque balance look like?"

Again, space limitations prevent us from presenting full torque diagnostics, however they will be presented in our
follow-up article. As we say in the article, this is an area for further research.

9. "The interaction of the planet with all the GI modes appears to be quite stochastic. It seems to me that very little
can be concluded from a single run."

The simulations presented here are by no means comprehensive, but are intended to give an indication of what possible
effects GIs might have on migration, and vice-versa. We provide as much detail for the given runs as allowed in the
Letter format, including additional planet runs (as will be done in the follow-up article) is not feasible in the
allowed space.

### 2011-07-12 Text/migration.tex
Responded to second referee report, should be the final submission.
	new file:   Response_to_referee2.txt
	modified:   general.bib
	modified:   migration.tex
	modified:   migration_apjlsubmission.tex

@@
 on their radial migration, which is a chaotic and messy affair in a fragmenting disk
-\citep[e.g.,][]{boley2009,boley2010,vorobyov2010,boley2010b}. The occurrence of gravitational instabilities (GIs) may be
+\citep[e.g.,][]{boley2009,boley2010,vorobyov2010,boley2010b}. The occurrence of gravitational instabilities may be
 episodic \citep[e.g.,][]{vorobyov2006,vorobyov2010b,zhu2010}. Clumps that survive and contract to the dimensions of
@@
 To improve our understanding of how planets migrate in GI-active disks that are not fragmenting, we have begun a
-systematic study, using numerical 3D radiative hydrodynamics, where we investigate the effects on both the disk and the
-planet of inserting a planet-mass object into disks susceptible to GIs. Using techniques developed in earlier research
-by our group \citep{pickett2003, mejia2005,cai2006,cai2008,boley2006,boley2007b,michael2010}, we can identify the
-dominant spiral waves in a simulation and analyze how the waves interact with the planet's motion. Our goal is to
-determine both the effect of giant planets on GIs and the effect of GIs on planet migration. Because GIs are sensitive
-to radiative physics, we use a well-tested radiative scheme \citep{boley2007b} and realistic opacities
-\citep{dalessio2001}.
+systematic study to investigate the effects on both the disk and the planet by inserting a planet-mass object into disks
+susceptible to GIs. Using techniques developed in earlier research by our group \citep{pickett2003,
+  mejia2005,cai2008,boley2006,boley2007b,michael2010}, we can identify the dominant spiral waves in a simulation and
+analyze how the waves interact with the planet's motion. Our goal is to determine both the effect of giant planets on
+GIs and the effect of GIs on planet migration. Because GIs are sensitive to radiative physics, we use a well-tested
+radiative scheme \citep{boley2007b} and realistic opacities \citep{dalessio2001}.
 
@@
 with discrete low-order modes, and stalling of migration near the inner edge of the strongest GI activity. The migration
-is neither a random walk nor monotonic but has a chaotic character due to nonlinear interactions. For all these reasons,
+is neither a random walk nor monotonic but has a chaotic character due to nonlinear interactions, similar to the erratic
+migration seen in planet/disk simulations with Rossby vortex instabilities \citep{yu2010}. For all these reasons,
 it will be important in future studies to explore the dependence of the planet's behavior on its mass and on its
@@
 NNX08AK36G). A.C.B. is currently a Sagan Postdoctoral Fellow at U. Florida. This material is based upon work supported
-by the National Science Foundation under Grants No. ACI-0338618l, OCI-0451237, OCI-0535258, OCI-0504075, and
-CNS-0521433, and through TeraGrid resources provided by Indiana University, PSC, SDSC, and NCSA.
+by the National Science Foundation under Grants No. OCI-0451237, OCI-0535258, OCI-0504075, and through TeraGrid resources provided by Indiana University, PSC, and NCSA.
 

### 2011-07-12 Text/migration_apjlsubmission.tex
Responded to second referee report, should be the final submission.
	new file:   Response_to_referee2.txt
	modified:   general.bib
	modified:   migration.tex
	modified:   migration_apjlsubmission.tex

@@
   {Mej{\'{\i}}a}, A.~C. 2008, ApJ, 673, 1138
-
-\bibitem[{{Cai} {et~al.}(2006){Cai}, {Durisen}, {Michael}, {Boley},
-  {Mej{\'{\i}}a}, {Pickett}, \& {D'Alessio}}]{cai2006}
-{Cai}, K., {Durisen}, R.~H., {Michael}, S., {Boley}, A.~C., {Mej{\'{\i}}a},
-  A.~C., {Pickett}, M.~K., \& {D'Alessio}, P. 2006, ApJL, 636, L149
 
@@
 
+\bibitem[{{Yu} {et~al.}(2010){Yu}, {Li}, {Li}, {Lubow}, \& {Lin}}]{yu2010}
+{Yu}, C., {Li}, H., {Li}, S., {Lubow}, S.~H., \& {Lin}, D.~N.~C. 2010, ApJ,
+  712, 198
+
 \bibitem[{{Zhu} {et~al.}(2010){Zhu}, {Hartmann}, \& {Gammie}}]{zhu2010}
@@
 \center
-\includegraphics[width=12cm]{Figures/amt.eps}
+\includegraphics[width=12cm]{amt.eps}
 \caption{(Color figure available in the online version) Global amplitudes of nonaxisymmetric density perturbations $A_m$
@@
 \center
-\includegraphics[width=12cm]{Figures/planeta.eps}
+\includegraphics[width=12cm]{planeta.eps}
 \caption{Plot of the planet's radial position as a function of time $\varpi_p(t)$ for the two planet simulations. The
@@
 \center
-\includegraphics[width=12cm]{Figures/planete.eps}
+\includegraphics[width=12cm]{planete.eps}
 \caption{Plot of the eccentricities $e$, computed for each 0 to 2$\phi$ change in azimuth of the planet as explained in
@@
 \center
-\includegraphics[width=12cm]{Figures/hr_migration.eps}
+\includegraphics[width=12cm]{hr_migration.eps}
 \caption{(Color figure available in the online version) Plots of the disk aspect ratio $h/r$ as determined by a decrease

### 2011-07-13 Text/migration.tex
Second submission to second referee report

@@
 at $\varpi \approx$ 32 AU, or about 180 yr. The disk is initially located between radial grid zones 30 and 240 and is
-close to isentropic, which results in a Toomre-$Q$ distribution with a marginally unstable \citep[see ][]{durisen2007}
+close to isentropic, which results in a Toomre-$Q$ distribution with a marginally unstable \citep[see][]{durisen2007}
 minimum value of 1.38 at radial grid zone 161 (26.7 AU). The computational grid extends radially to 512 zones to
@@
 upper vertical grid boundary, the outer radial grid boundary, and an inner radial boundary at 2 AU. To seed
-nonaxisymmetry, the density distribution is given an initial 0.01 \% random cell-to-cell perturbation.
+nonaxisymmetry, the density distribution is given an initial 0.01\% random cell-to-cell perturbation.
 
@@
 $M_J$ planet with a similar orbit is inserted into the fiducial run at $t = 10.5$ ORP. We estimate semi-major axes $a$
-and eccentricities $e$ from the planet's $\varpi(t)$ by using the maximum and minimum radii, $\varpi_{\rm max}$ and
+and eccentricities $e$ from the planet's $\varpi_p(t)$ by using the maximum and minimum radii, $\varpi_{\rm max}$ and
 $\varpi_{\rm min}$, over each complete orbit of the planet, where one orbit is $\Delta\phi=2\pi$. Specifically, $a =
@@
 somewhat more slowly, and lags the other modes in time. In the $t = 0$ planet run, the planet develops an organized wake
-within about an ORP in which $m = 3$ is a dominant component ($\sim 5$ \% global density perturbation). This nonlinear
+within about an ORP in which $m = 3$ is a dominant component ($\sim 5$\% global density perturbation). This nonlinear
 seeding causes $m = 3$ to dominate the GI burst, which also occurs $\sim$ 2 ORP earlier. Because of our initial
